### 2023-02-09 Schramm-Manuscript-2023/Schramm-Manuscript-2023.Rmd
results

---
title: Assessing linkages between watershed nutrient loading and estuary water quality in Lavaca Bay, Texas
author:
  - name: Michael Schramm
    affil: 1
    orcid: 0000-0003-1876-6592
affiliation:
  - num: 1
    address: |
      Texas A&M AgriLife Research - 
      Texas Water Resources Institute
      1001 Holleman Dr. E. 
      College Station, TX 77840-2118
    email: michael.schramm@ag.tamu.edu
# author citation list in chicago format
authorcitation: |
  Schramm, M.
correspondence: |
  michael.schramm@ag.tamu.edu
# specify the mdpi cls and location of file
# as of 2021 the template files are located in
# a subdirectory and should be retained there 
# since file locations are hard coded in the cls file.
cls: Definitions/mdpi
# class options
classoption:
  - water #first option is the journal name
  - article #second option is the type of manuscript
  - submit #do not change
  - oneauthor #change to oneauthor for single author papers
  #- pdftex #remove if compiling with XeLaTeX
# front matter
simplesummary: |
  A Simple summary goes here.
abstract: |
  A single paragraph of about 200 words maximum. For research articles, 
  abstracts should give a pertinent overview of the work. We strongly encourage
  authors to use the following style of structured abstracts, but without 
  headings: 1) Background: Place the question addressed in a broad context and
  highlight the purpose of the study; 2) Methods: Describe briefly the main
  methods or treatments applied; 3) Results: Summarize the article's main 
  findings; and 4) Conclusion: Indicate the main conclusions or interpretations. 
  The abstract should be an objective representation of the article, it must not 
  contain results which are not presented and substantiated in the main text and 
  should not exaggerate the main conclusions.
# back matter
keywords: |
  keyword 1; keyword 2; keyword 3 (list three to ten pertinent keywords specific 
  to the article, yet reasonably common within the subject discipline.).
acknowledgement: |
  All sources of funding of the study should be disclosed. Please clearly 
  indicate grants that you have received in support of your research work. 
  Clearly state if you received funds for covering the costs to publish in open 
  access.
funding: |
  This paper was funded by financial assistance provided
  by the Coastal Zone Management Act of 1972, as amended, administered by the
  National Oceanic and Atmospheric Administration (NOAA), Office for Coastal
  Management, pursuant to NOAA Award No. NA21NOS4190136. The views
  expressed herein are those of the author(s) and do not necessarily 
  reflect the views of NOAA, the U.S. Department of Commerce, or any of their 
  subagencies.
dataavailability: |
  Data and code are openly available in Zenodo at 
  https://doi.org/10.5281/zenodo.7330754.
conflictsofinterest: |
  The author declares no conflict of interest. The founding sponsors had no 
  role in the design of the study; in the collection, analyses, or 
  interpretation of data; in the writing of the manuscript, an in the decision 
  to publish the results.
bibliography: mybibfile.bib
endnotes: false
output: 
  rticles::mdpi_article:
    latex_engine: xelatex #pdflatex or XeLaTeX allowed
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
library(targets)
#library(unitar)
library(knitr)
library(tidyverse)
library(kableExtra)
library(patchwork)
library(ggrepel)
library(rcartocolor)
library(twriTemplates)
library(ggspatial)
library(ragg)
library(flextable)
library(modelsummary)
library(scico)
library(ggridges)
library(ggtext)
library(lubridate)
library(units)
library(gratia)

## this sets our default code chunk options
knitr::opts_chunk$set(dev = "cairo_pdf",
                      echo = FALSE,
                      ## figure chunk options
                      dpi = 300,
                      fig.pos = "H",
                      fig.align = "center",
                      out.width = "100%")

## custom hook for wide figures

hook_plot_wide <- function(x, options) {
  x_out <- knitr:::hook_plot_tex(x, options)
  if(isTRUE(options$wideplot)) {
    x_out <- gsub(x = x_out,
                  pattern='\\begin{figure}[H]', 
                  replacement='\\begin{figure}[H]\\begin{adjustwidth}{-\\extralength}{0cm}',
                  fixed=TRUE)

    x_out <- gsub(x = x_out,
                  pattern='\\caption',
                  replacement='\\end{adjustwidth}\\caption',
                  fixed = TRUE)
  }
  x_out
}
knitr::knit_hooks$set(plot=hook_plot_wide)

store <- "C:/Data-Analysis-Projects/lavaca-nutrients/_targets"
```

# Introduction

Like many estuaries globally, estuaries along the Texas Gulf coast are facing 
pressures from increasing population, increases in point source 
and non-point source pollution and alterations to freshwater inflows leading to
increases in the occurrences and risks of algal blooms and eutrophication [@bricker_effects_2008; @kennicuttWaterQualityGulf2017; 
@bugica_water_2020]. Recent studies indicate that estuary water quality
dynamics in both agriculturally dominated and urban watersheds within Texas are
displaying signals of conditions increasingly conducive to eutrophication 
[@wetzWaterQualityDynamics2016; @wetz_exceptionally_2017; 
@bugica_water_2020; @chinPhytoplanktonBiomassCommunity2022].

# Materials and Methods

## Study Area and Data

Lavaca Bay is a secondary bay in the Matagorda Bay system located on the 
Texas Gulf coast, roughly halfway between the cities of Houston and Corpus 
Christi (Figure  \ref{fig:fig1}). Lavaca Bay is 190 km^2^ with the majority
of freshwater inflow provided by the Lavaca and Navidad River systems. The 
Garcitas-Arenosa, Placedo Creek, and Cox Bay watersheds provide additional 
freshwater inflows. The entire watershed land area for Lavaca Bay is 8,149 km^2^. 
The Lavaca and Navidad River watersheds are a combined 5,966 km^2^, or 
approximately 73% of the entire Lavaca Bay watershed area. Discharge from the 
Navidad River is regulated by Lake Texana which has been in operation since 1980.
Lake Texana provides 170,000 acre-feet of water storage and discharges into the 
tidal section of the Navidad River which ultimately joins the tidal section of 
the Lavaca River 15 km upstream of the confluence with the Bay.

\begin{figure}[H]
\begin{adjustwidth}{-\extralength}{0cm}
\centering
\includegraphics[width=15.5cm]{Schramm-Manuscript-2023_files/figure-latex/fig1-1.png}
\end{adjustwidth}
\caption{Map of the Lavaca Bay watershed, location of USGS gages where nutrient loads were calculated, and location of estuary water quality sampling sites.\label{fig1}}
\end{figure}

Daily discharges for the Lavaca River (USGS-08164000) were obtained from the 
United States Geologic Survey (USGS) National Water Information System using the 
*dataRetrieval* R package [@deciccoDataRetrievalPackagesDiscovering2022]. Gaged
dailly discharges from Lake Texana (USGS-0816425) were provided by the Texas
Water Development Board (TWDB) (April 21, 2022 email from R. Neupane, TWDB).

Water quality sample data for both freshwater and estuary locations were o
btained from the Texas Commission on Environmental Quality (TCEQ) Surface Water 
Quality Monitoring Information System. Data submitted through the system 
are required to be collected under Quality Assurance Project Plans and lab 
method procedures outlined by the TCEQ's procedures manual. 
The QAPP and procedures manuals ensure the consistent collection and laboratory 
methods are applied between samples collected by different entities and under 
different projects. 
For freshwater locations, total phosphorus (TP) and nitrate-nitrogen (NO~3~) 
data were downloaded (Table \ref{tab:fwsummary}). Unfortunately, insufficient data was available for 
assessment of total nitrogen (TN) and total Kjeldahl nitrogen (TKN) loadings, 
so analysis was restricted to TP and NO~3~ loads. For estuary locations, we 
obtained data for TP, Nitrite+Nitrate (NO*~x~*), TKN, chlorophyll-*a*, and
dissolved oxygen (Table \ref{tab:estuarysummary}).

```{r fwsummary}
df <- tar_read(model_data, store = store)

df <- df |>  
  select(Date, site_no, Flow, original_NO3, original_TP) |>
  mutate(site_no = case_when(
    site_no == "lktexana_g" ~ "usgs08164525",
    site_no != "lktexana_g" ~ site_no
  )) |>
  filter(site_no == "usgs08164000" |
           site_no == "usgs08164525") |> 
  rename(`NO\\textsubscript{3} (mg/L)` = original_NO3,
         `TP (mg/L)` = original_TP,
         `Station ID` = site_no,
         `Mean Daily Streamflow (cfs)` = Flow)
df$`Station ID` <- str_replace(df$`Station ID`, "usgs", "USGS-")

datasummary(`Station ID` * (`TP (mg/L)` + `NO\\textsubscript{3} (mg/L)` + `Mean Daily Streamflow (cfs)`) ~
               Mean*Arguments(na.rm=TRUE) + SD*Arguments(na.rm=TRUE) + N,
            data = df, output = "kableExtra",
            booktabs = TRUE, table.envir = "table", 
            position = "H", escape = FALSE,
            caption = "Summary of gauged streamflow and freshwater water quality samples between January 1, 2000 and December 31, 2020.")
```

```{r estuarysummary}
df <- tar_read(estuary_model_data, store = store)

df <- df |> 
  filter(station_id == "13384" |
         station_id == "13383" |
         station_id == "13563") |>
  filter(parameter_code == "00630" |
           parameter_code == "00625" |
           parameter_code =="00665" |
           parameter_code == "70953" |
           parameter_code == "00300") |> 
  filter(end_date >= as.Date("2005-01-01")) |> 
  select(end_date, station_id, station_description, parameter_description, parameter_code, value) |> 
  mutate(parameter_description =
           case_when(
             parameter_code == "70953" ~ "Chlorophyll-\\emph{a} ($\\mu$g/L)",
             parameter_code == "00625" ~ "TKN (mg/L)",
             parameter_code == "00630" ~ "NO\\textsubscript{\\emph{x}} (mg/L)",
             parameter_code == "00665" ~ "TP (mg/L)",
             parameter_code == "00300" ~ "DO (mg/L)"
           )) |>
  mutate(station_id = paste0("TCEQ-", station_id)) |> 
  rename(`Station ID` = station_id) |> 
  pivot_wider(names_from = parameter_description,
              values_from = value)

 modelsummary::datasummary(`Station ID` * (`TP (mg/L)`+
                            `NO\\textsubscript{\\emph{x}} (mg/L)` +
                            `TKN (mg/L)` +
                            `Chlorophyll-\\emph{a} ($\\mu$g/L)` +
                            `DO (mg/L)`) ~
                            Mean*Arguments(na.rm=TRUE) + SD*Arguments(na.rm=TRUE) + N,
                          data = df, output = "dataframe") |> 
   mutate(`Station ID` = case_when(
     `Station ID` == "" ~ NA_character_,
     `Station ID` != "" ~ `Station ID`
     )) |>
   tidyr::fill(`Station ID`, .direction = "down") |> 
   kbl(format = "latex", booktabs = TRUE, table.envir = "table", 
       position = "H", escape = FALSE,
       caption = "Summary of estuary water quality samples collected between January 1, 2005 and December 31, 2020.") |> 
   collapse_rows(columns = 1, latex_hline = "major",
                 valign = "middle")

```

## Estimating Watershed Based Nutrient Loads

Estimates of nutrient loads were developed using Generalized Additive Models
(GAMs) relating nutrient concentration to river discharge, season, and time.
Separate models were fit at each station for each parameter and used to predict
nutrient concentrations for each day in the study period. GAMs can be specified
in a functionally similar manner to the commonly used LOADEST 
[@cohn_validity_1992] or WRTDS 
[@hirsch_weighted_2010] regression models and have been shown to
produce reliable estimates of nutrient and sediment loadings 
[@wangLoadEstimationUncertainties2011; @kroonRiverLoadsSuspended2012; 
@kuhnert_quantifying_2012; @robson_prediction_2015-1; 
@hagemannEstimatingNutrientOrganic2016; @mcdowell_implications_2021; 
@biagi_novel_2022]. GAMs are a semiparametric extension of generalized linear 
models where the linear predictor is represented as the sum of multiple unknown 
smooth functions and parametric linear predictors [@wood_fast_2011]. Although 
the underlying parameter estimation procedure of GAMs is substantially different 
than WRTDS, both the functional form and results are demonstrated to be similar [@beckNumericalQualitativeContrasts2017]. The use of GAMs over other 
regression-based approaches was (1) the ability to easily explore and 
incorporate different model terms, (2) the ability to incorporate non-linear
smooth function without explicit apriori knowledge of the expect shape, and (3)
the ability to specify a link function that relates the expected value of the
response to the linear predictors and allows use to avoid data transformations
as much as possible.

GAMs were fit using the *mgcv* package in R which makes available multiple 
types of smooth functions with automatic smoothness selection [@wood_fast_2011]. 
The general form of the model relating NO~3~ and TP concentration to streamflow,
season, and time was:

\begin{equation}\label{eq:1}
\begin{aligned}
g(\mu) &= \alpha + f_1(ddate) + f_2(yday) + f_3(log1p(Q)) + f_4(ma) + f_5(fa)  \\
y &\sim \mathcal{N}(\mu,\,\sigma^{2}),
\end{aligned}
\end{equation}

where $\mu$ is the conditional expected NO~3~-N or TP concentration, *g()* is the 
log-link, $\alpha$ is the intercept, *f~n~()* are smoothing functions. *y* is the 
response variable (NO~3~ or TP concentration) modeled as normally distributed 
with mean $\mu$ and standard deviation $\sigma$. *ddate* is the date converted to 
decimal notation, *yday* is numeric day of year (1-366), and *log1p(Q)* is the 
natural log of mean daily streamflow plus 1.

Moving average (*ma*) is an exponentially smoothed moving average that attempts 
to incorporate the influence of prior streamflow events on concentration at the 
current time period. @wangLoadEstimationUncertainties2011, 
@kuhnert_quantifying_2012 and @zhang_improving_2017 refer to this as averaged or 
smoothed discounted flow and demonstrated improvements in nutrient loading 
models by including the term. @kuhnert_quantifying_2012 expresses MA as:

\begin{equation}\label{eq:2}
ma(\delta) = d{\kappa_{i-1}}+(1-\delta)\hat{q}_{i-1}\quad\text{and}\quad \kappa_{i}=\sum_{m=1}^{i}\hat{Q}_m,
\end{equation}

where $\delta$ is the discount factor (here, set equal to 0.95), $\kappa_i$ 
is the cumulative flow (*Q*) up to the *i*th day.

Flow anomaly (*fa*) is a unitless term that represents how wet or dry the 
current time period is from a previous time period [@vecchia_trends_2009; 
@zhang_improving_2017]. Long-term flow anomaly (*ltfa*) is the streamflow over 
the previous year relative to the entire period and calculated as described 
by @zhang_improving_2017:

\begin{equation}\label{eq:3}
ltfa(t) = \bar{x}_{1\,year}(t) - \bar{x}_{entire\,period} 
\end{equation}

and the short-term flow anomaly (*stfa*) calculated as the current day flow 
compared to the preceding 1-month streamflow:

\begin{equation}\label{eq:4}
stfa(t) = x_{current\,day}(t) - \bar{x}_{1\,month}(t) 
\end{equation}

where *x* are the averages of log-transformed streamflow over the antecedent 
period (*1-year*, *1-month*, etc.) for time *t*.  We used *ltfa* in NO~3 models 
and *stfa* in TP models based on results from @zhang_improving_2017 
demonstrating major improvements in NO~x~ regression models that incorporated 
*ltfa* and moderate improvements in TP regression models that incorporated 
*stfa*.

The calculation of model terms for the Lake Texana site were slightly modified
because daily loads are not a function of natural stream flow processes alone,
but of dam releases and nutrient concentrations at the discharge point of the 
lake. *Q*, *ma*, and *fa* terms were calculated based on total gaged inflow from
the 4 major tributaries to the lake. Thin-plate regression splines were used 
for *ddate*, *log1p(Q)*, *fa*, and *ma*. A cyclic cubic regression spline was 
used for *yday* to ensure the ends of the spline match 
(day 1 and day 366 are expected to match). First order penalties were applied 
to the smooths of flow-based variables which penalize departures from a 
flat function to help constrain extrapolations for high flow measurements.

Left-censored data were not uncommon in this dataset. Several methods are
available to account for censored data. We transformed left-censored nutrient 
concentrations to one-half the detection limit. Although this simple approach 
can introduce bias [@hornungEstimationAverageConcentration1990],  we considered 
it acceptable because high concentrations and loadings are associated with 
high-flow events and low-flow/low-concentration events will account for a small 
proportion of total loadings [@mcdowell_implications_2021].

Daily loads were estimated as the predicted concentration multiplied by the 
daily streamflow. For the Lake Texana site, model terms were slightly modified
because daily loads are a function of dam releases and nutrient concentration,
but concentration will be a function of lake inflows and or other lake processes.
 *Q*, *ma*, and *fa* terms were calculated based on total gaged inflow from
the 4 major tributaries to the lake and laily loads at the dam were calculated 
from the discrete daily concentration at the discharge point of the lake and 
corresponding reported daily discharge from the dam.
Flow-normalized loads were estimated similar to WRTDS by setting flow-based 
covariates on each day of the year equal to each of the historical values for
that day of the year over the study period [@hirsch_weighted_2010]. The 
flow-normalized estimate was calculated as the mean of all the predictions for 
each day considering all possible flow values.
Standard deviations and credible intervals were obtained by 
drawing samples from the multivariate normal posterior distribution of the 
fitted GAM [@woodConfidenceIntervalsGeneralized2006; 
@marraCoveragePropertiesConfidence2012; @mcdowell_implications_2021].
Uncertainty in loads were reported as 90% credible intervals developed by 
drawing 1000 realizations of parameter estimates from the multivariate normal 
posterior distribution of the model parameters. GAM performance was evaluated 
using repeated 5-fold cross validation  [@burmanComparativeStudyOrdinary1989] 
and average Nash-Suttclifee Efficiency (NSE), r^2^ and percent bias (PBIAS) 
metrics across folds were calculated for each model.

## Linking Estuary Water Quality to Hydrology and Nutrient Loads

To test if changes in freshwater inflow and nutrient loading had explanatory
effect on changes in estuary water quality a series of GAM models were fit at
each site relating parameter concentration to temporal trends, inflow, and 
nutrient loads [@murphyNutrientImprovementsChesapeake2022]:

\begin{equation}\label{eq:5}
\begin{aligned}
g(\mu) &= \alpha + f_1(ddate) + f_2(yday) \\
y &\sim \Gamma(\mu,\lambda),
\end{aligned}
\end{equation}

\begin{equation}\label{eq:6}
\begin{aligned}
g(\mu) &= \alpha + f_1(ddate) + f_2(yday) + f_3(Q) \\
y &\sim \Gamma(\mu,\lambda),
\end{aligned}
\end{equation}

\begin{equation}\label{eq:7}
\begin{aligned}
g(\mu) &= \alpha + f_1(ddate) + f_2(yday) + f_3(Q) + f_4(Load) \\
y &\sim \Gamma(\mu,\lambda),
\end{aligned}
\end{equation}

where $\mu$ is the conditional expected response (nutrient concentration), *g()* 
is the log link, and response variable was modeled as Gamma distributed with 
mean $\mu$ and scale $\lambda$. *f~1~(ddate)* is decimal date smoothed with a 
thin-plate regression spline, *f~2~(yday)* is the numeric day of year smoothed 
with a cyclic cubic regression spline, *f~3~(Q)* is mean daily inflow (the 
combined measurements from Lavaca River and Lake Texana) and *f~4~(Load)* is 
the total NO~3~ or TP watershed load. The set of models specified for each water
quality response are in Table \ref{rab:estgammodels}.

```{r estgammodels, echo=FALSE}
df <- tibble(
  `Water Quality Response Parameter` = c(rep("TP", 3), rep("NO\\textsubscript{\\emph{x}}", 3),
               rep("Chlorophyll-\\emph{a}", 3), rep("Dissolved Oxygen", 3),
               rep("TKN", 2)),
  Model = c("Temporal", "Flow", "Flow+Load",
            "Temporal", "Flow", "Flow+Load",
            "Temporal", "Flow", "Flow+Load",
            "Temporal", "Flow", "Flow+Load",
            "Temporal", "Flow"),
  `Model Terms` = c("s(ddate) + s(yday)",
                    "s(ddate) + s(yday) + s(Q)",
                    "s(ddate) + s(yday) + s(Q) + s(TP Load)",
                    "s(ddate) + s(yday)",
                    "s(ddate) + s(yday) + s(Q)",
                    "s(ddate) + s(yday) + s(Q) + s(NO\\textsubscript{3} Load)",
                    "s(ddate) + s(yday)",
                    "s(ddate) + s(yday) + s(Q)",
                    "s(ddate) + s(yday) + s(Q) + s(TP Load) + s(NO\\textsubscript{3} Load)",
                    "s(ddate) + s(yday)",
                    "s(ddate) + s(yday) + s(Q)",
                    "s(ddate) + s(yday) + s(Q) + s(TP Load) + s(NO\\textsubscript{3}  Load)",
                    "s(ddate) + s(yday)",
                    "s(ddate) + s(yday) + s(Q)")
)

kbl(df, format = "latex",
    booktabs = TRUE,
    caption = "Set of GAM models specified for each water quality parameter response.",
    col.names = linebreak(c("Water Quality\nResponse Parameter",
                          "Model",
                          "Model Terms")),
    table.envir = "table", 
    position = "H", escape = FALSE) |> 
  collapse_rows(columns = 1, latex_hline = "major",
                valign = "middle")

```

Because streamflow and nutrient loads are tightly correlated, freshwater inflow
can mask signals from nutrient loads alone. Following the 
methodology implemented by @murphyNutrientImprovementsChesapeake2022,
both streamflow and nutrient loads were prepossessed to account for season
and flow. Instead of using raw freshwater inflow and nutrient loading values, 
these values were replaced by seasonally adjusted inflow and flow-adjusted 
nutrient loads by fitting a GAM relating season (day of year) to log transformed
daily freshwater inflow values:

\begin{equation}\label{eq:8}
g(\mu) = \alpha + f_1(yday),
\end{equation}

and a GAM relating log transformed NO~3~ or TP loads to log transformed daily
inflow:

\begin{equation}\label{eq:9}
g(\mu) = \alpha + f_1(log(Q)),
\end{equation}

where the response variables were modeled as normally distributed with an 
identity link function. Response residuals from the respective GAM models were 
used as *Q* and *Load* in Equation \ref{eq:6} and Equation \ref{eq:7}.

# Results

## Watershed Nutrient Loads

Based on criteria provided by @moriasiHydrologicWaterQuality2015, GAMs ranged
from "satisfactory" to "very good" based on median NSE, r^2^ and PBIAS 
calculated from repeated 5-fold cross validation on predicted and measured
nutrient loads. NO~3~ GAM models had median NSE values of 0.34 and 0.48 at
USGS-08164000 and USGS-08164390 respectively. Median r^2^ values were 0.70
(USGS-08164000) and 0.87 (USGS-08164525) and PBIAS values were 2.00
(USGS-08164000) and 10.90 (USGS-08164525) for NO~3~ loads. Median NSE values were
0.80 (USGS-08164000) and 0.91 (USGS-08164525) for TP loads. Median r^2^ values 
for TP loads were 0.93 (USGS-08164000) and 0.99 (USGS-08164525) and median PBIAS 
values were -7.20 (USGS-08164000) and -3.30 (USGS-08164525). Density plots of
metrics show similar distribution of values between sites for the 
same parameter, with the exception r^2^ values for NO~3~ loads where 
USGS-08164000 showed a much larger variance in values compared to USGS-08164525
(Figure \ref{fig:fig2}).In addition to higher average NSE and r^2^ values, 
GAMs had smaller variance in metric values for TP compared to NO~3~.

```{r fig2, message=FALSE, warning=FALSE, fig.height=3.5, fig.width=6.5, fig.pos="H", fig.cap="Density plots of goodness-of-fit metrics (NSE, r\\textsuperscript{2}, and PBIAS) from repeated 5-fold cross validation between predicted nutrient loads from GAM models and measured nutrient loads. Color indicates the tail probability calcualted from the empirical cumulative distribution of the goodness-of-fit metrics.", out.width="100%", wideplot=TRUE}

## some functions to overide scales in facets via dewey dunnington
#https://dewey.dunnington.ca/post/2018/modifying-facet-scales-in-ggplot2/
scale_override <- function(which, scale) {
  if(!is.numeric(which) || (length(which) != 1) || (which %% 1 != 0)) {
    stop("which must be an integer of length 1")
  }

  if(is.null(scale$aesthetics) || !any(c("x", "y") %in% scale$aesthetics)) {
    stop("scale must be an x or y position scale")
  }

  structure(list(which = which, scale = scale), class = "scale_override")
}

CustomFacetWrap <- ggproto(
  "CustomFacetWrap", FacetWrap,
  init_scales = function(self, layout, x_scale = NULL, y_scale = NULL, params) {
    # make the initial x, y scales list
    scales <- ggproto_parent(FacetWrap, self)$init_scales(layout, x_scale, y_scale, params)

    if(is.null(params$scale_overrides)) return(scales)

    max_scale_x <- length(scales$x)
    max_scale_y <- length(scales$y)

    # ... do some modification of the scales$x and scales$y here based on params$scale_overrides
    for(scale_override in params$scale_overrides) {
      which <- scale_override$which
      scale <- scale_override$scale

      if("x" %in% scale$aesthetics) {
        if(!is.null(scales$x)) {
          if(which < 0 || which > max_scale_x) stop("Invalid index of x scale: ", which)
          scales$x[[which]] <- scale$clone()
        }
      } else if("y" %in% scale$aesthetics) {
        if(!is.null(scales$y)) {
          if(which < 0 || which > max_scale_y) stop("Invalid index of y scale: ", which)
          scales$y[[which]] <- scale$clone()
        }
      } else {
        stop("Invalid scale")
      }
    }

    # return scales
    scales
  }
)

facet_wrap_custom <- function(..., scale_overrides = NULL) {
  # take advantage of the sanitizing that happens in facet_wrap
  facet_super <- facet_wrap(...)

  # sanitize scale overrides
  if(inherits(scale_overrides, "scale_override")) {
    scale_overrides <- list(scale_overrides)
  } else if(!is.list(scale_overrides) || 
            !all(vapply(scale_overrides, inherits, "scale_override", FUN.VALUE = logical(1)))) {
    stop("scale_overrides must be a scale_override object or a list of scale_override objects")
  }

  facet_super$params$scale_overrides <- scale_overrides

  ggproto(NULL, CustomFacetWrap,
          shrink = facet_super$shrink,
          params = facet_super$params
  )
}

df <- tar_read(cv_no3_08164000, store = store) |> 
  mutate(site = "USGS-08164000",
         parameter = "NO<sub>3</sub>") |> 
  bind_rows(
    tar_read(cv_tp_08164000, store = store) |> 
      mutate(site = "USGS-08164000",
             parameter = "TP")
  ) |> 
  bind_rows(
    tar_read(cv_no3_texana, store = store) |> 
      mutate(site = "USGS-08164525",
             parameter = "NO<sub>3</sub>")
  ) |> 
  bind_rows(
    tar_read(cv_tp_texana, store = store) |> 
      mutate(site = "USGS-08164525",
             parameter = "TP")
  )

df_long <- df |> 
  ungroup() |> 
  select(NSE, r2, pbias, site, parameter) |> 
  pivot_longer(cols = c(NSE, r2, pbias),
               names_to = "metric")

## plots the density estimates of the repeated 5-fold cross-validation goodness-of-fit metric results,
## color indicates the tail probability calculated from the empirical cumulitive distribution of the goodness-of-fit metric values 
p1 <- ggplot(df, aes(y = site, x = NSE,
               fill = 0.5 - abs(0.5 - after_stat(ecdf)))) +
  stat_density_ridges(geom = "density_ridges_gradient",
                      calc_ecdf = TRUE,
                      n = 10000) +
  scale_fill_scico("Tail Probability", palette = "hawaii", direction = -1,
                   limit = c(0, 0.5), breaks = c(0, 0.1, 0.2, 0.3, 0.4)) +
  guides(fill = guide_colorbar(barwidth = unit(100L, "pt"),
                               title.position = "top")) +
  facet_wrap_custom(~parameter, scales = "free_x",
                    ncol = 2,
                    scale_overrides = list(
                      scale_override(1, scale_x_continuous(limits = c(-1,1),
                                                           expand = c(0,0))),
                      scale_override(2, scale_x_continuous(limits = c(0,1),
                                                           expand = c(0,0)))
                    )) +
  labs(x = "NSE", y = "") +
  theme_TWRI_print() +
  theme(axis.text = element_text(size = 6),
        axis.title.x = element_markdown(size = 6),
        strip.text.x = element_markdown(size = 8),
        strip.background = element_rect(fill = "white"),
        panel.spacing.x = unit(20L, "pt"),
        legend.direction = "horizontal",
        legend.text = element_text(size = 8),
        legend.title = element_text(size = 8))

p2 <- ggplot(df, aes(y = site, x = r2,
               fill = 0.5 - abs(0.5 - after_stat(ecdf)))) +
  stat_density_ridges(geom = "density_ridges_gradient",
                      calc_ecdf = TRUE,
                      n = 10000) +
  scale_fill_scico("Tail Probability", palette = "hawaii", direction = -1,
                   limit = c(0, 0.5), breaks = c(0, 0.1, 0.2, 0.3, 0.4)) +
  guides(fill = guide_colorbar(barwidth = unit(100L, "pt"),
                               title.position = "top")) +
  facet_wrap_custom(~parameter, scales = "free_x",
                    ncol = 2,
                    scale_overrides = list(
                      scale_override(1, scale_x_continuous(limits = c(0,1),
                                                           expand = c(0,0))),
                      scale_override(2, scale_x_continuous(limits = c(0,1),
                                                           expand = c(0,0)))
                    )) +
  labs(x = "r<sup>2</sup>", y = "") +
  theme_TWRI_print() +
  theme(axis.text = element_text(size = 6),
        axis.title.x = element_markdown(size = 6),
        strip.text.x = element_markdown(size = 8),
        strip.background = element_rect(fill = "white"),
        panel.spacing.x = unit(20L, "pt"),
        legend.direction = "horizontal",
        legend.text = element_text(size = 8),
        legend.title = element_text(size = 8))

p3 <- ggplot(df, aes(y = site, x = pbias,
               fill = 0.5 - abs(0.5 - after_stat(ecdf)))) +
  stat_density_ridges(geom = "density_ridges_gradient",
                      calc_ecdf = TRUE,
                      n = 10000) +
  scale_fill_scico("Tail Probability", palette = "hawaii", direction = -1,
                   limit = c(0, 0.5), breaks = c(0, 0.1, 0.2, 0.3, 0.4)) +
  guides(fill = guide_colorbar(barwidth = unit(100L, "pt"),
                               title.position = "top")) +
  facet_wrap_custom(~parameter, scales = "free_x",
                    ncol = 2,
                    scale_overrides = list(
                      scale_override(1, scale_x_continuous(limits = c(-150,200),
                                                           expand = c(0,0))),
                      scale_override(2, scale_x_continuous(limits = c(-75,75),
                                                           expand = c(0,0)))
                    )) +
  labs(x = "PBIAS", y = "") +
  theme_TWRI_print() +
  theme(axis.text = element_text(size = 6),
        axis.title.x = element_markdown(size = 6),
        strip.text.x = element_markdown(size = 8),
        strip.background = element_rect(fill = "white"),
        panel.spacing.x = unit(20L, "pt"),
        legend.direction = "horizontal",
        legend.text = element_text(size = 8),
        legend.title = element_text(size = 8),
        axis.text.y = element_blank())

layout <- "
AAAA####
AAAA#DD#
AAAA####
BBBBCCCC
BBBBCCCC
BBBBCCCC
"

p1 / p2 / p3 / guide_area() + plot_layout(guides = 'collect',
                                          design = layout)

```

```{r echo=FALSE}
## calculates min and max load values and years
load_lav_no3 <- tar_read(daily_no3_08164000, store = store)
load_nav_no3 <- tar_read(daily_no3_texana, store = store)

load_lav_no3$annually |> 
  mutate(site = "Lavaca River") -> lavaca_no3_annually

load_nav_no3$annually |> 
  mutate(site = "Navidad River") -> navidad_no3_annually

mean_no3 <- scales::label_comma()(mean(lavaca_no3_annually$NO3_Estimate + navidad_no3_annually$NO3_Estimate ))
mean_no3_low <- scales::label_comma()(mean(lavaca_no3_annually$NO3_Lower + navidad_no3_annually$NO3_Lower ))
mean_no3_upper <- scales::label_comma()(mean(lavaca_no3_annually$NO3_Upper + navidad_no3_annually$NO3_Upper ))

no3 <- data.frame(lavaca = lavaca_no3_annually$NO3_Estimate,
                  navidad = navidad_no3_annually$NO3_Estimate,
                  year = lavaca_no3_annually$year)
no3 <- no3 |> 
  mutate(total = lavaca+navidad)

min_no3 <- min(no3$total)
min_no3_year <- no3[no3$total == min_no3,]$year
min_no3 <- scales::label_comma()(min_no3)

max_no3 <- max(no3$total)
max_no3_year <- no3[no3$total == max_no3,]$year
max_no3 <- scales::label_comma()(max_no3)

no3_prop <- no3 |> 
  mutate(prop = round(navidad/total, 2))

no3_mean_prop <- scales::label_percent()(mean(no3_prop$prop))

min_no3_prop <- scales::label_percent()(min(no3_prop$prop))

load_lav_tp <- tar_read(daily_tp_08164000, store = store)
load_nav_tp <- tar_read(daily_tp_texana, store = store)

load_lav_tp$annually |> 
  mutate(site = "Lavaca River") |> 
  filter(year >= 2005)-> lavaca_tp_annually

load_nav_tp$annually |> 
  mutate(site = "Navidad River") |> 
  filter(year >= 2005) -> navidad_tp_annually

mean_tp <- scales::label_comma()(mean(lavaca_tp_annually$TP_Estimate + navidad_tp_annually$TP_Estimate ))
mean_tp_low <- scales::label_comma()(mean(lavaca_tp_annually$TP_Lower + navidad_tp_annually$TP_Lower ))
mean_tp_upper <- scales::label_comma()(mean(lavaca_tp_annually$TP_Upper + navidad_tp_annually$TP_Upper ))

tp <- data.frame(lavaca = lavaca_tp_annually$TP_Estimate,
                  navidad = navidad_tp_annually$TP_Estimate,
                  year = lavaca_tp_annually$year)
tp <- tp |> 
  mutate(total = lavaca+navidad)

min_tp <- min(tp$total)
min_tp_year <- tp[tp$total == min_tp,]$year
min_tp <- scales::label_comma()(min_tp)

max_tp <- max(tp$total)
max_tp_year <- tp[tp$total == max_tp,]$year
max_tp <- scales::label_comma()(max_tp)

tp_prop <- tp |> 
  mutate(prop = round(navidad/total, 2))

tp_mean_prop <- scales::label_percent()(mean(tp_prop$prop))

min_tp_prop <- scales::label_percent()(min(tp_prop$prop))
```

Annual NO~3~ and TP loads show considerable variation, generally following
patterns in discharge (Figure \ref{fig:fig3}, Figure \ref{fig:fig4}). 
Flow-normalized TP loads at both sites and flow-normalized loads at 
USGS-08164000 indicate watershed based loads have not changed much over time
when accounting for changes in streamflow (Figure \ref{fig:fig3}). 
Flow-normalized loads at USGS-08164000 show small variation over time with 
some decreases in NO~3~ loads since 2013. Aggregated across both sites, the mean 
annual NO~3~ load 2005 through 2020 was `r mean_no3` kg (`r mean_no3_low` kg - 
`r mean_no3_upper` kg, 90% CI). Annual NO~3~ loads ranged from `r min_no3` kg 
in `r min_no3_year` to `r max_no3` kg in `r max_no3_year`.

Total annual TP loads ranged from `r min_tp` kg in `r min_tp_year` to 
`r max_tp` kg  in `r max_tp_year`. Mean annual TP loading from 2005 through 2020
was `r mean_tp` kg (`r mean_tp_low` kg - `r mean_tp_upper` kg, 90% CI).
On average, USGS-08164525 accounted for `r no3_mean_prop` of NO~3~ loads and 
`r tp_mean_prop` of TP loads from 2005 through 2020. However, during periods of extreme 
drought the Lavaca River (USGS-08164000) became the primary source of 
nutrient loading in the watershed with the Navidad River only accounting for 
`r min_no3_prop` and `r min_tp_prop` of NO~3~ and TP loads in 2011 (Figure 
\ref{fig:fig4}).

```{r fig3, error=FALSE, message=FALSE, fig.height=4, fig.width=5, fig.pos="H", fig.cap="Aggregated estimated annual and flow-normalized annual NO\textsubscript{3} and TP loads for USGS-08164000 and USGS-08164525.", out.width="100%"}
load <- tar_read(daily_no3_08164000, store = store)
fn_load <- tar_read(daily_no3_08164000_fn, store = store)

load_lav_tp <- tar_read(daily_tp_08164000, store = store)
load_nav_tp <- tar_read(daily_tp_texana, store = store)

a <- ggplot() +
  geom_point(data = load$annually, aes(year, NO3_Estimate,
                                   color = "Total Annual Load + 90% CI",
                                   shape = "Total Annual Load + 90% CI")) +
  geom_line(data = load$annually, aes(x = year, y = NO3_Estimate,
                                  color = "Total Annual Load + 90% CI",
                                  linetype = "Total Annual Load + 90% CI"),
            alpha = 0.5) +
  geom_linerange(data = load$annually, aes(x = year, ymin = NO3_Lower, ymax = NO3_Upper,
                                       color = "Total Annual Load + 90% CI")) +
  geom_point(data = fn_load$annually, aes(year, NO3_Estimate,
                                 color = "Flow-Normalized Annual Load + 90% CI",
                                 shape = "Flow-Normalized Annual Load + 90% CI")) +
  geom_line(data = fn_load$annually,
            aes(x = year, y = NO3_Estimate,
                color = "Flow-Normalized Annual Load + 90% CI",
                linetype = "Flow-Normalized Annual Load + 90% CI"),
            alpha = 0.5) +
  geom_linerange(data = fn_load$annually,
                 aes(x = year, ymin = NO3_Lower, ymax = NO3_Upper,
                     color = "Flow-Normalized Annual Load + 90% CI")) +
  labs(x = "", y = "Annual NO<sub>3</sub> Load [kg]", subtitle = "USGS-08164000") +
  scale_shape_manual(name = "values",
                     values = c(21, 19)) +
  scale_color_manual(name = "values",
                     values = c("#7E1900", "#1A3399")) +
  scale_linetype_manual(name = "values",
                        values = c(1, 2)) +
  scale_y_continuous(trans = "pseudo_log",
                     breaks = c(0,1E1, 1E2, 1E3, 1E4, 1E5, 1E6, 1E7),
                     labels = scales::label_log()) +
  coord_cartesian(ylim = c(1000, 1E7)) +
  theme_TWRI_print() +
  theme(axis.title.y = element_markdown(size = 8),
        plot.subtitle = element_text(size = 8),
        panel.grid.major.x = element_line(color = "#d9d9d9",
                                          linetype = "dotted"),
        legend.title = element_blank(),
        legend.text = element_text(size = 8))

load <- tar_read(daily_no3_texana, store = store)
fn_load <- tar_read(daily_no3_texana_fn, store = store)

b <- ggplot() +
  geom_point(data = load$annually, aes(year, NO3_Estimate,
                                   color = "Total Annual Load + 90% CI",
                                   shape = "Total Annual Load + 90% CI")) +
  geom_line(data = load$annually, aes(x = year, y = NO3_Estimate,
                                  color = "Total Annual Load + 90% CI",
                                  linetype = "Total Annual Load + 90% CI"),
            alpha = 0.5) +
  geom_linerange(data = load$annually, aes(x = year, ymin = NO3_Lower, ymax = NO3_Upper,
                                       color = "Total Annual Load + 90% CI")) +
  geom_point(data = fn_load$annually, aes(year, NO3_Estimate,
                                 color = "Flow-Normalized Annual Load + 90% CI",
                                 shape = "Flow-Normalized Annual Load + 90% CI")) +
  geom_line(data = fn_load$annually,
            aes(x = year, y = NO3_Estimate,
                color = "Flow-Normalized Annual Load + 90% CI",
                linetype = "Flow-Normalized Annual Load + 90% CI"),
            alpha = 0.5) +
  geom_linerange(data = fn_load$annually,
                 aes(x = year, ymin = NO3_Lower, ymax = NO3_Upper,
                     color = "Flow-Normalized Annual Load + 90% CI")) +
  labs(x = "", y = "Annual NO<sub>3</sub> Load [kg]", subtitle = "USGS-08164525") +
  scale_shape_manual(name = "values",
                     values = c(21, 19)) +
  scale_color_manual(name = "values",
                     values = c("#7E1900", "#1A3399")) +
  scale_linetype_manual(name = "values",
                        values = c(1, 2)) +
  scale_y_continuous(trans = "pseudo_log",
                     breaks = c(0,1E1, 1E2, 1E3, 1E4, 1E5, 1E6, 1E7),
                     labels = scales::label_log()) +
  coord_cartesian(ylim = c(1000, 1E7)) +
  theme_TWRI_print() +
  theme(axis.title.y = element_markdown(size = 8),
        plot.subtitle = element_text(size = 8),
        panel.grid.major.x = element_line(color = "#d9d9d9",
                                          linetype = "dotted"),
        legend.title = element_blank(),
        legend.text = element_text(size = 8))

load <- tar_read(daily_tp_08164000, store = store)
fn_load <- tar_read(daily_tp_08164000_fn, store = store)

load$annually <- load$annually |> 
  filter(year >= 2005)

fn_load$annually <- fn_load$annually |> 
  filter(year >= 2005)

c <- ggplot() +
  geom_point(data = load$annually, aes(year, TP_Estimate,
                                   color = "Total Annual Load + 90% CI",
                                   shape = "Total Annual Load + 90% CI")) +
  geom_line(data = load$annually, aes(x = year, y = TP_Estimate,
                                  color = "Total Annual Load + 90% CI",
                                  linetype = "Total Annual Load + 90% CI"),
            alpha = 0.5) +
  geom_linerange(data = load$annually, aes(x = year, ymin = TP_Lower, ymax = TP_Upper,
                                       color = "Total Annual Load + 90% CI")) +
  geom_point(data = fn_load$annually, aes(year, TP_Estimate,
                                 color = "Flow-Normalized Annual Load + 90% CI",
                                 shape = "Flow-Normalized Annual Load + 90% CI")) +
  geom_line(data = fn_load$annually,
            aes(x = year, y = TP_Estimate,
                color = "Flow-Normalized Annual Load + 90% CI",
                linetype = "Flow-Normalized Annual Load + 90% CI"),
            alpha = 0.5) +
  geom_linerange(data = fn_load$annually,
                 aes(x = year, ymin = TP_Lower, ymax = TP_Upper,
                     color = "Flow-Normalized Annual Load + 90% CI")) +
  labs(x = "", y = "Annual TP Load [kg]", subtitle = "USGS-08164000") +
  scale_shape_manual(name = "values",
                     values = c(21, 19)) +
  scale_color_manual(name = "values",
                     values = c("#7E1900", "#1A3399")) +
  scale_linetype_manual(name = "values",
                        values = c(1, 2)) +
  scale_y_continuous(trans = "pseudo_log",
                     breaks = c(0,1E1, 1E2, 1E3, 1E4, 1E5, 1E6, 1E7),
                     labels = scales::label_log()) +
  coord_cartesian(ylim = c(1000, 1E6)) +
  theme_TWRI_print() +
  theme(axis.title.y = element_markdown(size = 8),
        plot.subtitle = element_text(size = 8),
        panel.grid.major.x = element_line(color = "#d9d9d9",
                                          linetype = "dotted"),
        legend.title = element_blank(),
        legend.text = element_text(size = 8))

load <- tar_read(daily_tp_texana, store = store) 
fn_load <- tar_read(daily_tp_texana_fn, store = store)

load$annually <- load$annually |> 
  filter(year >= 2005)

fn_load$annually <- fn_load$annually |> 
  filter(year >= 2005)

d <- ggplot() +
  geom_point(data = load$annually, aes(year, TP_Estimate,
                                   color = "Total Annual Load + 90% CI",
                                   shape = "Total Annual Load + 90% CI")) +
  geom_line(data = load$annually, aes(x = year, y = TP_Estimate,
                                  color = "Total Annual Load + 90% CI",
                                  linetype = "Total Annual Load + 90% CI"),
            alpha = 0.5) +
  geom_linerange(data = load$annually, aes(x = year, ymin = TP_Lower, ymax = TP_Upper,
                                       color = "Total Annual Load + 90% CI")) +
  geom_point(data = fn_load$annually, aes(year, TP_Estimate,
                                 color = "Flow-Normalized Annual Load + 90% CI",
                                 shape = "Flow-Normalized Annual Load + 90% CI")) +
  geom_line(data = fn_load$annually,
            aes(x = year, y = TP_Estimate,
                color = "Flow-Normalized Annual Load + 90% CI",
                linetype = "Flow-Normalized Annual Load + 90% CI"),
            alpha = 0.5) +
  geom_linerange(data = fn_load$annually,
                 aes(x = year, ymin = TP_Lower, ymax = TP_Upper,
                     color = "Flow-Normalized Annual Load + 90% CI")) +
  labs(x = "", y = "Annual TP Load [kg]", subtitle = "USGS-08164525") +
  scale_shape_manual(name = "values",
                     values = c(21, 19)) +
  scale_color_manual(name = "values",
                     values = c("#7E1900", "#1A3399")) +
  scale_linetype_manual(name = "values",
                        values = c(1, 2)) +
  scale_y_continuous(trans = "pseudo_log",
                     breaks = c(0,1E1, 1E2, 1E3, 1E4, 1E5, 1E6, 1E7),
                     labels = scales::label_log()) +
  coord_cartesian(ylim = c(1000, 1E6)) +
  theme_TWRI_print() +
  theme(axis.title.y = element_markdown(size = 8),
        plot.subtitle = element_text(size = 8),
        panel.grid.major.x = element_line(color = "#d9d9d9",
                                          linetype = "dotted"),
        legend.title = element_blank(),
        legend.text = element_text(size = 8))

layout <- "
11112222
11112222
11112222
11112222
33334444
33334444
33334444
33334444
55555555
"

a + b + c + d + guide_area() + plot_layout(guides = 'collect', design = layout)
```

```{r fig4, error=FALSE, message=FALSE, fig.height=4, fig.width=5, fig.pos="H", fig.cap="Comparison of delivered annual loads at USGS-08164000 and USGS-08164525.", out.width="100%"}
load_lav_no3 <- tar_read(daily_no3_08164000, store = store)
load_nav_no3 <- tar_read(daily_no3_texana, store = store)

load_lav_no3$annually |> 
  mutate(site = "Lavaca River") -> lavaca_no3_annually

load_nav_no3$annually |> 
  mutate(site = "Navidad River") -> navidad_no3_annually

no3_annual <- bind_rows(lavaca_no3_annually, navidad_no3_annually) |> 
  mutate(x = as.Date(paste0(year, "-01-01"), "%Y-%m-%d"))

b <- ggplot(no3_annual) +
  geom_col(aes(year, NO3_Estimate, fill = site), width = 0.8) +
  scale_x_continuous(expand = expansion(mult = c(0.05, 0.05)), 
                     breaks = c(2005, 2010, 2015, 2020)) +
  scale_y_continuous(labels = scales::comma) +
  scale_fill_carto_d(palette = "Vivid", direction = -1) +
  labs(x = "", y = "Predicted Annual<br>NO<sub>3</sub>-N Load [kg]") +
  theme_TWRI_print() +
  theme(axis.title.y = element_markdown(size = 8),
        axis.text = element_text(size = 8),
        panel.grid.major.x = element_line(color = "#d9d9d9",
                                          linetype = "dotted"),
        legend.title = element_blank(),
        legend.text = element_text(size = 8))

prop <- no3_annual |> 
  mutate(proportion = NO3_Estimate/sum(NO3_Estimate))

c <- ggplot() +
  geom_col(data = prop, aes(year, proportion, fill  = site)) +
  scale_fill_carto_d(palette = "Vivid", direction = -1) +
  scale_x_continuous(breaks = c(2005, 2010, 2015, 2020),
                     expand = expansion(mult = c(0.05, 0.05))) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = "", y = "Proportion of Annual<br>NO<sub>3</sub> Load") +
  theme_TWRI_print() +
  theme(axis.title.y = element_markdown(size = 8),
        axis.text = element_text(size = 8),
        panel.grid.major.x = element_line(color = "#d9d9d9",
                                          linetype = "dotted"),
        legend.title = element_blank(),
        legend.text = element_text(size = 8))

d <- tar_read(qdata, store = store) |> 
  filter(site_no %in% c("8164000", "lktexana_g")) |>
  filter(Date >= as.Date("2005-01-01")) |> 
  mutate(site = case_when(
    site_no == "8164000" ~ "Lavaca River",
    site_no == "lktexana_g" ~ "Navidad River"
  )) |> 
  mutate(year = year(Date)) |>

  mutate(Flow = as_units(Flow, "ft^3/s")) |> 
  mutate(Flow = set_units(Flow, "ft^3/day")) |> 
  mutate(Flow = set_units(Flow, "1E6gallons/day")) |> 
  group_by(year, site) |> 
  summarise(Flow = sum(Flow, na.rm = TRUE)) |> 
  mutate(Flow = drop_units(Flow)) |> 
  ggplot() +
  geom_col(aes(year, Flow, fill = site)) +
  scale_fill_carto_d(palette = "Vivid", direction = -1) +
  scale_x_continuous(breaks = c(2005, 2010, 2015, 2020),
                     expand = expansion(mult = c(0.05, 0.05))) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = "", y = "Annual Discharge<br>[Millon Gallons]") +
  theme_TWRI_print() +
  theme(axis.title.y = element_markdown(size = 8),
        axis.text = element_text(size = 8),
        panel.grid.major.x = element_line(color = "#d9d9d9",
                                          linetype = "dotted"),
        legend.title = element_blank(),
        legend.text = element_text(size = 8))

load_lav_tp$annually |> 
  mutate(site = "Lavaca River") -> lavaca_tp_annually

load_nav_tp$annually |> 
  mutate(site = "Navidad River") -> navidad_tp_annually

tp_annual <- bind_rows(lavaca_tp_annually, navidad_tp_annually) |>
  filter(year >= 2005) |> 
  mutate(x = as.Date(paste0(year, "-01-01"), "%Y-%m-%d"))

e <- ggplot(tp_annual) +
  geom_col(aes(year, TP_Estimate, fill = site), width = 0.8) +
  scale_x_continuous(expand = expansion(mult = c(0.05, 0.05)), 
                     breaks = c(2005, 2010, 2015, 2020)) +
  scale_y_continuous(labels = scales::comma) +
  scale_fill_carto_d(palette = "Vivid", direction = -1) +
  labs(x = "", y = "Predicted Annual<br>TP Load [kg]") +
  theme_TWRI_print() +
  theme(axis.title.y = element_markdown(size = 8),
        axis.text = element_text(size = 8),
        panel.grid.major.x = element_line(color = "#d9d9d9",
                                          linetype = "dotted"),
        legend.title = element_blank(),
        legend.text = element_text(size = 8))

prop <- tp_annual |> 
  mutate(proportion = TP_Estimate/sum(TP_Estimate))

f <- ggplot() +
  geom_col(data = prop, aes(year, proportion, fill  = site)) +
  scale_fill_carto_d(palette = "Vivid", direction = -1) +
  scale_x_continuous(breaks = c(2005, 2010, 2015, 2020),
                     expand = expansion(mult = c(0.05, 0.05))) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = "", y = "Proportion of Annual<br>TP Load") +
  theme_TWRI_print() +
  theme(axis.title.y = element_markdown(size = 8),
        axis.text = element_text(size = 8),
        panel.grid.major.x = element_line(color = "#d9d9d9",
                                          linetype = "dotted"),
        legend.title = element_blank(),
        legend.text = element_text(size = 8))

layout <- "
BE
BE
BE
BE
CF
CF
CF
CF
DD
DD
DD
DD
GG
"

b + c + d + e + f + guide_area() + 
  plot_layout(guides = 'collect', design  = layout)
```

## Linkages Between Water Quality and Watershed Flows and Loads

```{r fig5, message=FALSE, warning=FALSE, fig.height=3.5, fig.pos="H", fig.width=6, fig.cap="Smoothed temporal trend component for water quality paramaters obtained from temporal estuary GAMs.", out.width="100%", wideplot=TRUE}
draw_smoothed_years <- function(model1, model2, model3,
                                ylab, subtitle) {

  `TCEQ-13563` <- model1
  `TCEQ-13383` <- model2
  `TCEQ-13384` <- model3

  comp <- gratia::compare_smooths(
    `TCEQ-13563`,
    `TCEQ-13383`,
    `TCEQ-13384`,
    smooths = "s(ddate)")

  crit <- gratia:::coverage_normal(0.90)

  comp |> 
    unnest(data) |> 
    mutate(lower_ci = est + (crit * se),
           upper_ci = est - (crit * se),
           label = if_else(ddate == max(ddate),
                           as.character(model),
                           NA_character_)) |> 
    ggplot(aes(x = ddate, y = est, group = model)) +
    geom_ribbon(aes(ymin = lower_ci,
                    ymax = upper_ci,
                    fill = model),
                alpha = 0.15) +
    geom_line(aes(color = model)) +
    geom_hline(yintercept = 0, linetype = 2, alpha = 0.25) +
    # geom_text_repel(aes(label = label,
    #                     color = model), 
    #                 nudge_x = 5,
    #                 hjust = "left",
    #                 size = 2) +
    scale_x_continuous(breaks = c(2005, 2010, 2015, 2020),
                       expand = expansion(mult = c(0.05,0.05))) +
    colorspace::scale_fill_discrete_qualitative(name = "") +
    colorspace::scale_color_discrete_qualitative(name = "") +
    labs(x = "Year", y = ylab,
         subtitle = subtitle) +
    theme_TWRI_print() +
    theme(axis.title.y = element_markdown(size = 8),
          axis.title.x = element_markdown(size = 8),
          plot.subtitle = element_markdown(size = 8),
          legend.direction = "vertical",
          legend.text = element_text(size = 8))

}

p_tp <- draw_smoothed_years(tar_read(tp_lavaca_13563_temporal, store = store),
                            tar_read(tp_lavaca_13383_temporal, store = store),
                            tar_read(tp_lavaca_13384_temporal, store = store),
                            ylab = "TP Smooth<br>Estimate [mg/L]", 
                            subtitle = "A: TP Trend")

p_no3 <- draw_smoothed_years(tar_read(no3_lavaca_13563_temporal, store = store),
                             tar_read(no3_lavaca_13383_temporal, store = store),
                             tar_read(no3_lavaca_13384_temporal, store = store),
                             ylab = "NO*<sub>x</sub>* Smooth<br>Estimate [mg/L]", 
                             subtitle = "B: NO*<sub>x</sub>* Trend")

p_tkn <- draw_smoothed_years(tar_read(tkn_lavaca_13563_temporal, store = store),
                             tar_read(tkn_lavaca_13383_temporal, store = store),
                             tar_read(tkn_lavaca_13384_temporal, store = store),
                             ylab = "TKN Smooth<br>Estimate [mg/L]", 
                             subtitle = "C: TKN Trend")

p_chla <- draw_smoothed_years(tar_read(chla_lavaca_13563_temporal, store = store),
                             tar_read(chla_lavaca_13383_temporal, store = store),
                             tar_read(chla_lavaca_13384_temporal, store = store),
                             ylab = "Chlorophyll-*a*\nSmooth<br>Estimate [µg/L]", 
                             subtitle = "D: Chlorophyll-*a* Trend")

p_do <- draw_smoothed_years(tar_read(do_lavaca_13563_temporal, store = store),
                             tar_read(do_lavaca_13383_temporal, store = store),
                             tar_read(do_lavaca_13384_temporal, store = store),
                             ylab = "DO\nSmooth<br>Estimate [mg/L]", 
                            subtitle = "E: DO Trend")

design <- "
112266
334455
"

#(p_tp + p_no3) / (p_chla + p_tkn) / (p_do)
p_tp + p_no3 + p_chla + p_tkn + p_do + guide_area() + plot_layout(guides = "collect",
                                                   design = design)
```

GAMs did not identify significant changes in TP or DO concentrations at any
of the Lavaca Bay sites from 2005 through 2020 (Figure \ref{fig:fig5}). The
upper-bay site, TCEQ-13563, had a linear increase in NO~x~ concentration and
and decrease in chlorophyll-*a* from 2005 through 2014. The mid-bay site,
TCEQ-13383, showed a periodic pattern in NO~x~ concentration that appeared
similar to precipitation/inflow patterns, as well as a post 2011 increase
in TKN concentrations. No significant long-term trends in concentrations 
were identified by GAMs for the lower-bay TCEQ-13384 site.

Freshwater inflow provided additional explanation for changes in TP and 
NO~*x*~ concentration at all of the Lavaca Bay sites according to AIC~c~ and 
model probability values (Table \ref{tab:tab4}).
TCEQ-13563, the site closest to the river outlet, was the only site that
had improvements in the explanations of DO and TKN concentration with the inclusion 
of inflow. Both TCEQ-13563 and TCEQ-13383, the mid-bay site, saw improvements 
in explanations for variations in chlorophyll-*a* with the inclusion of freshwater
inflow. The addition of nutrient loads (both TP and NO~3~) terms did not provide
additional explanation for changes in chlorophyll-*a* or DO concentrations.
Inclusion of TP loads provided additional explanation of TP concentrations at
the upper- and mid-bay sites, TCEQ-13563 and TCEQ-13383. Inclusion of NO~3~
loads only provided marginal improvements in the explanation of NO~*X*~ 
concentration at the lower-bay TCEQ-13384 site.

```{r tab4, echo=FALSE, message=FALSE, warning=FALSE}

dplyr::bind_rows(
    tar_read(tp_13563, store = store), 
    tar_read(tp_13383, store = store),
    tar_read(tp_13384, store = store),
    tar_read(no3_13563, store = store), 
    tar_read(no3_13383, store = store),
    tar_read(no3_13384, store = store),
    tar_read(chla_13563, store = store), 
    tar_read(chla_13383, store = store),
    tar_read(chla_13384, store = store),
    tar_read(do_13563, store = store), 
    tar_read(do_13383, store = store),
    tar_read(do_13384, store = store),
    tar_read(tkn_13563, store = store),
    tar_read(tkn_13383, store = store),
    tar_read(tkn_13384, store = store)
    ) |> 
  mutate(Parameter = case_when(
    Parameter == "Nitrite+Nitrate" ~ "NO\\textsubscript{\\emph{x}}",
    Parameter == "Chlorophyll-a" ~ "Chlorophyll-\\emph{a}",
    TRUE ~ as.character(Parameter)
  )) |> 
  mutate(Parameter = fct_relevel(Parameter, "TP", "NO\\textsubscript{\\emph{x}}", "Chlorophyll-\\emph{a}","TKN", "DO"),
         Model = fct_relevel(Model, "Temporal", "Flow", "Flow + Load")) |>
  tidyr::complete(Site, Model, nesting(Parameter)) |>
  mutate(model_prob = paste0(round(AICc,1), " (", round(weight,2), ")")) |> 
  mutate(model_prob = case_when(
    model_prob == "NA (NA)" ~ "-",
    model_prob != "NA (NA)" ~ model_prob
  )) |> 
  select(-c(AICc, weight)) |> 
  pivot_wider(names_from = Model, values_from = model_prob) |> 
  select(Parameter, Site, Temporal, Flow, `Flow + Load`) |> 
  arrange(Parameter, Site) |> 
  kbl(format = "latex",
    booktabs = TRUE,
    table.envir = "table", 
    position = "H", escape = FALSE,
    caption = "Estuary GAM AIC\\textsubscript{c} values and associated model probabilities (in parenthesis). Models with the highest probability for each site and water quality parameter combination are bolded and italicized for emphasis.") |> 
    collapse_rows(columns = 1,
                valign = "middle", latex_hline = "major") |> 
  column_spec(3, bold = c(rep(F, 7), T, F, T, T, F, T, T, F),
              italic = c(rep(F, 7), T, F, T, T, F, T, T, F)) |> 
  column_spec(4, bold = c(F,T,F,T,F,T,T,F,T,F,F,T,F,F,T),
              italic = c(F,T,F,T,F,T,T,F,T,F,F,T,F,F,T)) |> 
  column_spec(5, bold = c(T,F,T,F,T,F,F,F,F,F,F,F,F,F,F),
              italic =c(T,F,T,F,T,F,F,F,F,F,F,F,F,F,F))

```

```{r fig6, message=FALSE, warning=FALSE, fig.height=3.5, fig.pos="H", fig.cap="Estimated effects of mean daily inflow residuals on mean TP, NO\\textsubscript{\\emph{x}}, chlorophyll-\\emph{a}, TKN, and DO concentrations in Lavaca Bay obtained from flow estuary GAMs.", out.width="100%", wideplot=TRUE}
draw_smoothed_surface <- function(comp,
                                  x = flw_res,
                                  xlab,
                                  ylab, 
                                  subtitle) {

  crit <- gratia:::coverage_normal(0.90)

  comp |>
    unnest(data) |>
    mutate(lower_ci = est + (crit * se),
           upper_ci = est - (crit * se)) |>
    group_by(model) |> 
    mutate(label = if_else(
      {{ x }} == max( {{ x }} ),
      as.character(model),
      NA_character_)) |> 
    ggplot(aes(x = {{ x }}, y = est, group = model)) +
    geom_ribbon(aes(ymin = lower_ci,
                    ymax = upper_ci,
                    fill = model),
                alpha = 0.15) +
    geom_line(aes(color = model)) +
    geom_hline(yintercept = 0, linetype = 2, alpha = 0.5) +
    # geom_text_repel(aes(label = label,
    #                     color = model),
    #                 nudge_x = 5,
    #                 hjust = "left",
    #                 size = 2) +
    scale_x_continuous(expand = expansion(mult = c(0.05,0.05))) +
    colorspace::scale_fill_discrete_qualitative(name = "") +
    colorspace::scale_color_discrete_qualitative(name = "") +
    labs(x = xlab, y = ylab,
         subtitle = subtitle) +
    theme_TWRI_print() +
    theme(axis.title.y = element_markdown(size = 8),
          axis.title.x = element_markdown(size = 8),
          plot.subtitle = element_markdown(size = 8),
          legend.direction = "vertical",
          legend.text = element_text(size = 8))

}

`TCEQ-13383` <- tar_read(tp_lavaca_13383_flow, store = store)

comp <- compare_smooths(model = `TCEQ-13383`,
                      `TCEQ-13384` = tar_read(tp_lavaca_13384_flow, store = store),
                      `TCEQ-13563` = tar_read(tp_lavaca_13563_flow, store = store),
                      smooths = c("s(flw_res)"))
p_tp <- draw_smoothed_surface(comp,
                      x = flw_res,
                      xlab = "log(Inflow) Residuals [cfs]",
                      ylab =  "TP Smooth<br>Estimate [mg/L]",
                      subtitle = "A: TP")

`TCEQ-13383` <- tar_read(no3_lavaca_13383_flow, store = store)

comp <- compare_smooths(model = `TCEQ-13383`,
                      `TCEQ-13384` = tar_read(no3_lavaca_13384_flow, store = store),
                      `TCEQ-13563` = tar_read(no3_lavaca_13563_flow, store = store),
                      smooths = c("s(flw_res)"))
p_no3 <- draw_smoothed_surface(comp,
                      x = flw_res,
                      xlab = "log(Inflow) Residuals [cfs]",
                      ylab = "NO*<sub>x</sub>* Smooth<br>Estimate [mg/L]",
                      subtitle = "B: NO*<sub>x</sub>*")

`TCEQ-13383` <- tar_read(chla_lavaca_13383_flow, store = store)

comp <- compare_smooths(model = `TCEQ-13383`,
                      `TCEQ-13384` = tar_read(chla_lavaca_13384_flow, store = store),
                      `TCEQ-13563` = tar_read(chla_lavaca_13563_flow, store = store),
                      smooths = c("s(flw_res)"))
p_chla <- draw_smoothed_surface(comp,
                      x = flw_res,
                      xlab = "log(Inflow) Residuals [cfs]",
                      ylab = "Chlorophyll-*a* Smooth<br>Estimate [µg/L]",
                      subtitle = "C: Chlorophyll-*a*")

`TCEQ-13383` <- tar_read(tkn_lavaca_13383_flow, store = store)

comp <- compare_smooths(model = `TCEQ-13383`,
                      `TCEQ-13384` = tar_read(tkn_lavaca_13384_flow, store = store),
                      `TCEQ-13563` = tar_read(tkn_lavaca_13563_flow, store = store),
                      smooths = c("s(flw_res)"))
p_tkn <- draw_smoothed_surface(comp,
                      x = flw_res,
                      xlab = "log(Inflow) Residuals [cfs]",
                      ylab = "TKN Smooth<br>Estimate [mg/L]",
                      subtitle = "D: TKN")

`TCEQ-13383` <- tar_read(do_lavaca_13383_flow, store = store)

comp <- compare_smooths(model = `TCEQ-13383`,
                      `TCEQ-13384` = tar_read(do_lavaca_13384_flow, store = store),
                      `TCEQ-13563` = tar_read(do_lavaca_13563_flow, store = store),
                      smooths = c("s(flw_res)"))
p_do <- draw_smoothed_surface(comp,
                      x = flw_res,
                      xlab = "log(Inflow) Residuals [cfs]",
                      ylab = "DO Smooth<br>Estimate [mg/L]",
                      subtitle = "E: DO")

design <- "
112266
334455
"

p_tp + p_no3 + p_chla + p_tkn + p_do + guide_area() + plot_layout(guides = "collect",
                                                   design = design)

```

GAMs show that increases in freshwater inflow result in nearly linear increases 
in TP and NO~*x*~ at all three sites (Figure \ref{fig:6}). At the upper-bay
TCEQ-13563 site, GAMs showed increases in freshwater inflow initially 
increased chlorophyll-*a* and DO concentration but concentrations leveled and
potentially decreased at higher flows. The mid-bay TCEQ-13383 site showed a
nearly linear increased in chlorophyll-*a* concentration with increases in
freshwater inflow. Freshwater flow did not have significant
effects on chlorophyll-*a*, TKN, or DO at the lower-bay TCEQ-13384 site.

Increases in TP loads resulted in nearly linear increases of TP concentration at
the upper- and mid-bay sites, TCEQ-13563 and TCEQ-13383 respectively (Figure \ref{fig:fig7}).
The relative effect size appeared to much smaller than the effect of freshwater
inflow alone. Increases in NO~3~ loads only showed an effect at the 
lower-bay TCEQ-13384 site. The effect was quite small compared to streamflow and
provided only small improvements to the model (Table \ref{tab:tab4}). As noted
above, nutrient loadings did not provide any explanation in changes in the
remaining assessed water quality parameters.

```{r fig7, message=FALSE, warning=FALSE, fig.height=2.25, fig.width=5.5, fig.pos="H", fig.cap="Estimated effects of nutrient load residuals on TP and NO\\textsubscript{\\emph{x}} concentrations in Lavaca Bay obtained from flow+load estuary GAMs.", out.width="100%", wideplot=FALSE}

`TCEQ-13383` <- tar_read(tp_lavaca_13383_full, store = store)

comp <- compare_smooths(model = `TCEQ-13383`,
                      `TCEQ-13384` = tar_read(tp_lavaca_13384_full, store = store),
                      `TCEQ-13563` = tar_read(tp_lavaca_13563_full, store = store),
                      smooths = c("s(TP_resid)"))
p_tp <- draw_smoothed_surface(comp,
                      x = TP_resid,
                      xlab = "TP Load Residuals [kg]",
                      ylab = "TP [mg/L]",
                      subtitle = "A: TP")

`TCEQ-13383` <- tar_read(no3_lavaca_13383_full, store = store)

comp <- compare_smooths(model = `TCEQ-13383`,
                      `TCEQ-13384` = tar_read(no3_lavaca_13384_full, store = store),
                      `TCEQ-13563` = tar_read(no3_lavaca_13563_full, store = store),
                      smooths = c("s(NO3_resid)"))
p_no3 <- draw_smoothed_surface(comp,
                      x = NO3_resid,
                      xlab = "NO<sub>3</sub> Load Residuals [kg]",
                      ylab = "NO*<sub>x</sub>* [mg/L]",
                      subtitle = "B: NO*<sub>x</sub>*")

design <- "
11223
"

p_tp + p_no3 + guide_area() + plot_layout(guides = "collect",
                                                   design = design)

```

# Discussion

```{r tpcomp}

load_lav_tp <- tar_read(daily_tp_08164000, store = store)
load_nav_tp <- tar_read(daily_tp_texana, store = store)

dat <- load_lav_tp$annually |> 
  ungroup() |> 
  filter(year >= 2005) |> 
  summarise(annual = round(mean(TP_Estimate)/2116,1),
            upper = round(mean(TP_Upper)/2116,1),
            lower = round(mean(TP_Lower)/2116,1)) ## 2116 sq km

df <- tibble(
  Parameter = rep("TP", 5),
  `Reported Yield (kg$\\cdot$km\\textsuperscript{-2}$\\cdot$year\\textsuperscript{-1})` = 
    c(paste0(dat[[1,1]]," (", dat[[1,3]], ", ", dat[[1,2]], ")"), 
      "45.2", "42", "20.81-91.58", "28.9"),
  Approach = c("GAM", "SPARROW", "SWAT", "SPARROW", "LOADEST"),
  `Time Period` = c("2005-2020", "2000-2014", "1977-2005", "1980-2002", "1972-1993"),
  Reference = c("This work", "\\citet{wise_spatially_2019}",
                "\\citet{omaniEstimationSedimentNutrient2014}",
                "\\citet{rebich_sources_2011}",
                "\\citet{dunnTrendsNutrientInflows1996}")
)

df[[1,2]] <- paste0(df[[1,2]], footnote_marker_symbol(1))
df[[4,2]] <- paste0(df[[4,2]], footnote_marker_symbol(2))

kbl(df, format = "latex",
    booktabs = TRUE,
    table.envir = "table",
    position = "H", escape = FALSE,
    col.names = linebreak(c("Parameter",
                            "Reported Yield\n(kg$\\cdot$km\\textsuperscript{-2}$\\cdot$year\\textsuperscript{-1})",
                          "Approach",
                          "Time Period",
                          "Reference")),
    caption = "Mean estimates of annual TP yield in the Lavaca River watershed in published studies.") |>
  add_footnote(c("Values represent the mean of annual point estimates, lower and upper 95% credible intervals.",
                 "A single point estimate was not reported, these value represent the range depicted on the choropleth map provided in the report."),
               notation = "symbol",
               threeparttable = TRUE,
               escape = TRUE)

```

Authors should discuss the results and how they can be interpreted in 
perspective of previous studies and of the working hypotheses. The findings and 
their implications should be discussed in the broadest context possible. Future 
research directions may also be highlighted.

# Conclusion

This section is not mandatory, but can be added to the manuscript if the
discussion is unusually long or complex.