### 2023-02-20 Schramm_2023_02_PLOS/Schramm_2023_02_PLOS.qmd
format plos paper

---
title: "Assessing linkages between watershed nutrient loading and water quality in a subtropical estuary with semiparametric models."
author: 
  - name: Michael Schramm
    email: michael.schramm@ag.tamu.edu
    corresponding: true
    affiliations:
      - ref: agrilife
affiliations:
  - id: agrilife
    name: Texas A&M AgriLife Research, Texas Water Resources Institute
    city: College Station
    state: Texas
    country: United States of America
format: 
  docx:
    number-sections: false
    reference-doc: format.docx
crossref:
  fig-prefix: Fig   # (default is "Figure")
  tbl-prefix: Table    # (default is "Table")
  title-delim: .     # (default is ":")
  ref-hyperlink: false # (default is true)
bibliography: mybibfile.bib
csl: https://www.zotero.org/styles/plos?source=1
filters: 
  - authors-affiliations.lua
keep-md: true
---

<!--
TODO:
- find right csl format
- fix cross refs
- fix equations
- bold captions(?)
- fix symbols, subscripts, superscripts in tables
 List corresponding author’s initials in 
parentheses after the email address.
make intial field? for corresponding author?
- use imagemagick to convert png to tiff?
-->

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
library(targets)
library(knitr)
library(tidyverse)
library(kableExtra)
library(patchwork)
library(ggrepel)
library(rcartocolor)
library(twriTemplates)
library(ggspatial)
library(ragg)
library(flextable)
library(modelsummary)
library(scico)
library(ggridges)
library(ggtext)
library(lubridate)
library(units)
library(gratia)
library(gt)

ragg_tiff <- function(..., res = 300) {
  ragg::agg_tiff(..., res = res, units = "in")
}

## this sets our default code chunk options
knitr::opts_chunk$set(echo = FALSE,
                      ## figure chunk options
                      dev = "CairoTIFF",
                      dpi = 300,
                      fig.align = "center",
                      out.width = "100%")

store <- "C:/Data-Analysis-Projects/lavaca-nutrients/_targets"
```

# Abstract
Lavaca Bay is a small secondary embayment on the Texas coast that is
displaying early signals of water quality degradation. This
study applied a semiparametric approach to assess both watershed nutrient loads 
and the responses in estuary water quality to nutrient loading and streamflow. 
Cross-validation indicated that, despite data constraints, semiparametric models 
performed well at nutrient load prediction. Based on these models, delivered
annual nutrient loads varied substantially from year to year. In contrast,
minimal changes in calculated flow-normalized loads indicate that changes in 
nutrient loading were driven by natural variation in precipitation and runoff as
opposed to changes in discharge or nonpoint source management efforts.
Models did not identify significant long-term changes within Lavaca Bay for dissolved 
oxygen or chlorophyll-*a*. However, site specific long-term increases in both
organic and inorganic nitrogen are concerning. Further analysis found freshwater
inflow was a strong driver of nutrient and chlorophyll-*a* concentrations within
Lavaca Bay but there was no evidence that changes in nutrient loading explained
variation in dissolved oxygen or chlorophyll-*a* concentrations. In addition to
providing a baseline assessment of watershed nutrient loading and water quality
responses in the Lavaca Bay watershed, this study provides methodological
support for the use of semiparametric methods in load regression models and 
estuary assessments.

# Introduction

Like many coastal areas globally, the coastal watersheds along the Texas Gulf coast 
are facing pressures from increasing population, increases in point source 
and non-point source pollution and alterations to freshwater flows that degrade
water quality in downstream estuaries [@bricker_effects_2008; @kennicuttWaterQualityGulf2017; @bugica_water_2020]. Despite these increasing pressures, national scale 
assessments have classified coastal estuaries in Texas as moderate or lower for
exhibiting eutrophic conditions [@bricker_effects_2008]. However, a suite of 
recent studies indicate that estuary water quality dynamics in both 
agriculturally dominated and urban watersheds within Texas are in fact 
expressing conditions that are increasingly conducive to algal blooms and 
eutrophication [@wetzWaterQualityDynamics2016; @wetz_exceptionally_2017; 
@bugica_water_2020; @chinPhytoplanktonBiomassCommunity2022]. With identification
of localized areas of estuary water quality concern along the Texas coast 
[@bugica_water_2020], localized studies are being prioritized to better 
inform management actions.

This project provides an assessment of nutrient loading and water quality
responses in Lavaca Bay, Texas. Lavaca Bay is a secondary bay in the larger
Matagorda Bay system located roughly halfway between Houston, Texas and Corpus
Christi, Texas. Lavaca Bay faces substantial challenges associated with
legacy contamination but general water quality parameters such as dissolved oxygen 
(DO), nutrients, and biological parameters have been well within state
water quality standards. However, long-term declines of benthic 
fauna abundance, biomass, and diversity in Lavaca Bay primarily
linked to reductions in freshwater inflows and changes in estuary salinity 
[@beserespollackLongtermTrendsResponse2011; @palmerImpactsDroughtsLow2015;
@montagnaAssessmentRelationshipFreshwater2020] are a concern to local 
stakeholders. Recent water quality assessments identified monotonic increases in total 
phosphorus (TP), orthophosphate, total Kjeldahl nitrogen (TKN), and
chlorophyll-*a* at sites within Lavaca Bay [@bugica_water_2020]. Although 
long-term changes in DO concentrations were not identified, the trends in 
nutrient concentrations are concerning due to the role of nitrogen as a 
limiting factor for primary production in many Texas estuaries 
[@gardnerNitrogenFixationDissimilatory2006; @houTransformationFateNitrate2012; @doradoUnderstandingInteractionsFreshwater2015; @paudelRelationshipSuspendedSolids2019; @wetz_exceptionally_2017] and the 
ramifications that changes in nitrogen loadings could have for 
productivity and eutrophication in Lavaca Bay.

There are ongoing efforts between local, state, and federal agencies to 
address water quality impairments in the freshwater portions of the 
Lavaca Bay watershed [@jainTechnicalSupportDocument2021; @schramm_lavaca_2018; @bertholdDirectMailingEducation2021]. However, at a statewide scale, these
approaches have shown limited success and emphasize a need for improved efforts
at assessing and linking management actions with downstream water quality to
better identify  and replicate effective management actions across the state
[@schrammTotalMaximumDaily2022]. The identification and communication of 
changes and trends in water quality is complicated by the fact that trends are 
often non-linear and confounded by precipitation and runoff that 
hinder traditional analysis [@wazniakLinkingWaterQuality2007; 
@lloydMethodsDetectingChange2014]. To provide
actionable information for resource managers, water quality conditions must be 
evaluated relative to changes in natural environmental drivers to 
better understand  and manage potential anthropogenic effects.
This study utilizes semiparemetric methods to develop estimates of delivered 
and flow-normalized nutrient loads and assess changes in loads delivered to 
Lavaca Bay. The study also assesses the response of water quality parameters in
Lavaca Bay over time and in response to freshwater inflow controlled
for seasonality and to watershed nutrient loads that are controlled for
environmentally driven variation.

# Materials and Methods

## Study Area and Data

Lavaca Bay is 190 km^2^ with the majority of freshwater inflow provided by the 
Lavaca and Navidad River systems (@fig-fig1). The Garcitas-Arenosa, 
Placedo Creek, and Cox Bay watersheds provide additional freshwater inflows. 
The entire watershed land area is 8,149 km^2^ and primarily rural. Watershed 
land cover and land use is 50% grazed pasture and rangeland, 20% cultivated 
cropland (primarily rows crops such as corn, cotton, and sorghum), and 5% 
suburban/urban. Pasture and rangeland is concentrated in the Lavaca River 
watershed, while cultivated crops are generally located along the eastern 
tributaries of the Navidad river. The Lavaca and Navidad River watersheds are a 
combined 5,966 km^2^, or approximately 73% of the entire Lavaca Bay watershed 
area. Discharge from the Navidad River is regulated by Lake Texana which has 
been in operation since 1980. Lake Texana provides 0.210 km^3^ of water 
storage and discharges into the tidal section of the Navidad River which 
ultimately joins the tidal section of the Lavaca River 15 km upstream of the 
confluence with the Lavaca Bay.

```{r}
#| label: fig-fig1
#| fig-width: 5.2
#| fig-height: 6.75
#| out-width: "5.2in"
#| fig-cap: "Map of Lavaca Bay and the contribution watershed. The freshwater sites are the most downstream freshwater stream locations with water quality and streamflow data used for nutrient load models. Water quality concentration data at the three Lavaca Bay sites were used to assess relationships between freshwater flows, loads and estuary water quality."

knitr::include_graphics("Figures/fig1-1.tiff")

```

Daily discharges for the Lavaca River (USGS-08164000, @fig-fig1) 
were obtained from the United States Geologic Survey (USGS) National Water 
Information System using the *dataRetrieval* R package 
[@deciccoDataRetrievalPackagesDiscovering2022]. Gaged daily discharges from 
the outlet of Lake Texana on the Navidad River (USGS-0816425) were provided by 
the Texas Water Development Board (TWDB) 
(April 21, 2022 email from R. Neupane, TWDB).

Water quality sample data for the two freshwater and three estuary locations were 
obtained from the Texas Commission on Environmental Quality (TCEQ) Surface Water 
Quality Monitoring Information System. Data submitted through the system 
are required to be collected under Quality Assurance Project Plans and lab 
method procedures outlined by the TCEQ's procedures manual. 
The QAPP and procedures manuals ensure the consistent collection and laboratory 
methods are applied between samples collected by different entities and under 
different projects. 
All sites had varying lengths of and availability of data.
For freshwater locations, TP from January 2000 through December 2020 and 
nitrate-nitrogen (NO~3~) data from January 2005 through December 2020 were 
downloaded (@tbl-table1). Less than 5-years of total
nitrogen and TKN concentration data were available at the freshwater sites and
deemed insufficient to develop load estimation models 
[@horowitzEvaluationSedimentRating2003; @snelderEstimationCatchmentNutrient2017]. 
The three estuary sites included an upper Lavaca Bay site near the outlet of the
Lavaca River system (TCEQ-13563), a mid-Lavaca Bay site (TCEQ-13383), and the 
lower Lavaca Bay site near the mouth of the Bay (TCEQ-13384). 
For estuary locations, we obtained data for TP, Nitrite+Nitrate (NO*~x~*), TKN, 
chlorophyll-*a*, and DO concentrations from January 2005 through December 2020 
(@tbl-table2).

```{r}
#| label: tbl-table1
#| tbl-cap: "Summary of gauged streamflow and freshwater water quality samples between January 1, 2000 and December 31, 2020."
#| warning: false
df <- tar_read(model_data, store = store)

df <- df |>  
  select(Date, site_no, Flow, original_NO3, original_TP) |>
  mutate(site_no = case_when(
    site_no == "lktexana_g" ~ "usgs08164525",
    site_no != "lktexana_g" ~ site_no
  )) |>
  filter(site_no == "usgs08164000" |
           site_no == "usgs08164525") |> 
  rename(`NO3 (mg/L)` = original_NO3,
         `TP (mg/L)` = original_TP,
         `Station ID` = site_no,
         `Mean Daily Streamflow (cfs)` = Flow)
df$`Station ID` <- str_replace(df$`Station ID`, "usgs", "USGS-")

datasummary(`Station ID` * (`TP (mg/L)` + `NO3 (mg/L)` + `Mean Daily Streamflow (cfs)`) ~
               Mean*Arguments(na.rm=TRUE) + SD*Arguments(na.rm=TRUE) + N,
            data = df, output = "kableExtra", format = "pipe") 
```

```{r}
#| label: tbl-table2
#| warning: false
#| tbl-cap: "Summary of estuary water quality samples collected between January 1, 2005 and December 31, 2020."
df <- tar_read(estuary_model_data, store = store)

df <- df |> 
  filter(station_id == "13384" |
         station_id == "13383" |
         station_id == "13563") |>
  filter(parameter_code == "00630" |
           parameter_code == "00625" |
           parameter_code =="00665" |
           parameter_code == "70953" |
           parameter_code == "00300") |> 
  filter(end_date >= as.Date("2005-01-01")) |> 
  select(end_date, station_id, station_description, parameter_description, parameter_code, value) |> 
  mutate(parameter_description =
           case_when(
             parameter_code == "70953" ~ "Chlorophyll-a (ug/L)",
             parameter_code == "00625" ~ "TKN (mg/L)",
             parameter_code == "00630" ~ "NOx (mg/L)",
             parameter_code == "00665" ~ "TP (mg/L)",
             parameter_code == "00300" ~ "DO (mg/L)"
           )) |>
  mutate(station_id = paste0("TCEQ-", station_id)) |> 
  rename(`Station ID` = station_id) |> 
  pivot_wider(names_from = parameter_description,
              values_from = value)

 modelsummary::datasummary(`Station ID` * (`TP (mg/L)`+
                            `NOx (mg/L)` +
                            `TKN (mg/L)` +
                            `Chlorophyll-a (ug/L)` +
                            `DO (mg/L)`) ~
                            Mean*Arguments(na.rm=TRUE) + SD*Arguments(na.rm=TRUE) + N,
                          data = df, output = "kableExtra", format = "pipe") #|>
  #  flextable::autofit()
   # mutate(`Station ID` = case_when(
   #   `Station ID` == "" ~ NA_character_,
   #   `Station ID` != "" ~ `Station ID`
   #   )) |>
   # tidyr::fill(`Station ID`, .direction = "down") |> 
   # gt::gt() |> 
   # gt::as_word()

```

## Estimating Watershed Based Nutrient Loads

Estimates of NO~3~ and TP loads at the Lavaca River (USGS-08164000) and the 
outlet of Lake Texana on the Navidad River (USGS-08164525) were developed using 
Generalized Additive Models (GAMs) relating nutrient concentration to river 
discharge, season, and time.
Separate models were fit at each station for each parameter and used to predict
nutrient concentrations for each day in the study period. GAMs can be specified
in a functionally similar manner to the commonly used LOADEST 
[@cohn_validity_1992] or WRTDS 
[@hirsch_weighted_2010] regression models and have been shown to
produce reliable estimates of nutrient and sediment loadings 
[@wangLoadEstimationUncertainties2011; @kroonRiverLoadsSuspended2012; 
@kuhnert_quantifying_2012; @robson_prediction_2015-1; 
@hagemannEstimatingNutrientOrganic2016; @mcdowell_implications_2021; 
@biagi_novel_2022]. GAMs are a semiparametric extension of generalized linear 
models where the linear predictor is represented as the sum of multiple unknown 
smooth functions and parametric linear predictors [@wood_fast_2011]. Although 
the underlying parameter estimation procedure of GAMs is substantially different 
than WRTDS, both the functional form and results are demonstrated to be similar [@beckNumericalQualitativeContrasts2017]. GAMs were chosen over other regression-based
approached for use in this study due to; (1) the ability to easily explore and 
incorporate different model terms; (2) the incorporation of non-liner smooth
functions that do not require explicit a priori knowledge of the expected
shape; and (3) inclusion of a link function that related the expected value of
the response to linear predictors thus avoiding unneeded data transformations
and bias corrections.

GAMs were fit using the *mgcv* package in R which makes available multiple 
types of smooth functions with automatic smoothness selection [@wood_fast_2011]. 
The general form of the model related NO~3~ or TP concentration to a long term
tend, season, streamflow, and two different antecedent discharge terms:

$$
g(\mu) = \alpha + f_1(ddate) + f_2(yday) + f_3(log1p(Q)) + f_4(ma) + f_5(fa)  
y \sim \mathcal{N}(\mu,\,\sigma^{2})
$${#eq-eq1}

where *μ* is the conditional expected NO~3~ or TP concentration, *g()* is the 
log-link, *α* is the intercept, *f~n~()* are smoothing functions. *y* is the 
response variable (NO~3~ or TP concentration) modeled as normally distributed 
with mean *μ* and standard deviation *σ*. *ddate* is the date converted to 
decimal notation, *yday* is numeric day of year (1-366), and *log1p(Q)* is the 
natural log of mean daily streamflow plus 1.

Moving average (*ma*) is an exponentially smoothed moving average that attempts 
to incorporate the influence of prior streamflow events on concentration at the 
current time period [@wangLoadEstimationUncertainties2011;  
@kuhnert_quantifying_2012; @zhang_improving_2017]:

$$
ma(\delta) = d{\kappa_{i-1}}+(1-\delta)\hat{q}_{i-1}\quad\text{and}\quad \kappa_{i}=\sum_{m=1}^{i}\hat{Q}_m
$${#eq-eq2}

where *δ* is the discount factor (here, set equal to 0.95), κ~i~
is the cumulative flow (*Q*) up to the *i*th day.

Flow anomaly (*fa*) is a unitless term that represents how wet or dry the 
current time period is from a previous time period [@vecchia_trends_2009; 
@zhang_improving_2017]. Long-term flow anomaly (*ltfa*) is the streamflow over 
the previous year relative to the entire period and calculated as described as
[@zhang_improving_2017]:

$$
ltfa(t) = \bar{x}_{1\,year}(t) - \bar{x}_{entire\,period}
$${#eq-eq3}

and the short-term flow anomaly (*stfa*) calculated as the current day flow 
compared to the preceding 1-month streamflow:

$$
stfa(t) = x_{current\,day}(t) - \bar{x}_{1\,month}(t)
$${#eq-eq4}

where *x* are the averages of log-transformed streamflow over the antecedent 
period (*1-year*, *1-month*, etc.) for time *t*.  We used *ltfa* in NO~3~ models 
and *stfa* in TP models based on previous work  
demonstrating major improvements in NO~x~ regression models that incorporated 
*ltfa* and moderate improvements in TP regression models that incorporated 
*stfa* [@zhang_improving_2017].

The calculation of model terms for the Lake Texana site were slightly modified
because daily loads are not a function of natural stream flow processes alone,
but of dam releases and nutrient concentrations at the discharge point of the 
lake. *Q*, *ma*, and *fa* terms were calculated based on total gaged inflow from
the 4 major tributaries to the lake. Thin-plate regression splines were used 
for *ddate*, *log1p(Q)*, *fa*, and *ma*. A cyclic cubic regression spline was 
used for *yday* to ensure the ends of the spline match 
(day 1 and day 366 are expected to match). First order penalties were applied 
to the smooths of flow-based variables which penalize departures from a 
flat function to help constrain extrapolations for high flow measurements.

Left-censored data were not uncommon in this dataset. Several methods are
available to account for censored data. We transformed left-censored nutrient 
concentrations to one-half the detection limit. Although this simple approach 
can introduce bias [@hornungEstimationAverageConcentration1990],  we considered 
it acceptable because high concentrations and loadings are associated with 
high-flow events and low-flow/low-concentration events will account for a small 
proportion of total loadings [@mcdowell_implications_2021].

Daily loads were estimated as the predicted concentration multiplied by the 
daily streamflow. For the Navidad River (USGS-08164525) site, daily loads at the 
dam were calculated 
from the discrete daily concentration at the discharge point of the lake and 
corresponding reported daily discharge from the dam.
Flow-normalized loads were estimated similar to WRTDS by setting flow-based 
covariates on each day of the year equal to each of the historical values for
that day of the year over the study period [@hirsch_weighted_2010]. The 
flow-normalized estimate was calculated as the mean of all the predictions for 
each day considering all possible flow values.
Standard deviations and credible intervals were obtained by 
drawing samples from the multivariate normal posterior distribution of the 
fitted GAM [@woodConfidenceIntervalsGeneralized2006; 
@marraCoveragePropertiesConfidence2012; @mcdowell_implications_2021].
Uncertainty in loads were calculated as 90% credible intervals estimated by 
drawing 1000 realizations of parameter estimates from the multivariate normal 
posterior distribution of the model parameters. GAM performance was evaluated 
using repeated 5-fold cross validation  [@burmanComparativeStudyOrdinary1989] 
and average Nash-Sutcliffe Efficiency (NSE), r^2^ and percent bias (PBIAS) 
metrics across folds were calculated for each model.

## Linking Estuary Water Quality to Hydrology and Nutrient Loads

To test if changes in freshwater inflow and nutrient loading had explanatory
effect on changes in estuary water quality a series of GAM models were fit at
each site relating parameter concentration to temporal trends, inflow, and 
nutrient loads [@murphyNutrientImprovementsChesapeake2022]:

$$
g(\mu) = \alpha + f_1(ddate) + f_2(yday) y \sim \Gamma(\mu,\lambda)
$${#eq-eq5}

$$
g(\mu) = \alpha + f_1(ddate) + f_2(yday) + f_3(Q) y \sim \Gamma(\mu,\lambda)
$${#eq-eq6}

$$
g(\mu) = \alpha + f_1(ddate) + f_2(yday) + f_3(Q) + f_4(Load) y \sim \Gamma(\mu,\lambda)
$${#eq-eq7}

where *μ* is the conditional expected response (nutrient concentration), *g()* 
is the log link, and response variable was modeled as Gamma distributed with 
mean *μ* and scale *λ*. *f~1~(ddate)* is decimal date smoothed with a 
thin-plate regression spline, *f~2~(yday)* is the numeric day of year smoothed 
with a cyclic cubic regression spline, *f~3~(Q)* is mean daily inflow (the 
combined measurements from Lavaca River and Navidad River) and *f~4~(Load)* is 
the total NO~3~ or TP watershed load. The set of models specified for each water
quality response are in @tbl-table3.

```{r}
#| label: tbl-table3
#| tbl-cap: "Set of GAM models specified for each water quality parameter response."
#| warning: false
df <- tibble(
  `Water Quality Response Parameter` = c(rep("TP", 3), rep("NOx", 3),
               rep("Chlorophyll-a", 3), rep("Dissolved Oxygen", 3),
               rep("TKN", 2)),
  Model = c("Temporal", "Flow", "Flow+Load",
            "Temporal", "Flow", "Flow+Load",
            "Temporal", "Flow", "Flow+Load",
            "Temporal", "Flow", "Flow+Load",
            "Temporal", "Flow"),
  `Model Terms` = c("s(ddate) + s(yday)",
                    "s(ddate) + s(yday) + s(Q)",
                    "s(ddate) + s(yday) + s(Q) + s(TP Load)",
                    "s(ddate) + s(yday)",
                    "s(ddate) + s(yday) + s(Q)",
                    "s(ddate) + s(yday) + s(Q) + s(NO3 Load)",
                    "s(ddate) + s(yday)",
                    "s(ddate) + s(yday) + s(Q)",
                    "s(ddate) + s(yday) + s(Q) + s(TP Load) + s(NO3 Load)",
                    "s(ddate) + s(yday)",
                    "s(ddate) + s(yday) + s(Q)",
                    "s(ddate) + s(yday) + s(Q) + s(TP Load) + s(NO3  Load)",
                    "s(ddate) + s(yday)",
                    "s(ddate) + s(yday) + s(Q)")
)

kbl(df, format = "pipe")

```

Because streamflow and nutrient loads are tightly correlated, freshwater inflow
can mask signals from nutrient loads alone. Following the 
methodology implemented by @murphyNutrientImprovementsChesapeake2022,
both streamflow and nutrient loads were prepossessed to account for season
and flow. Freshwater inflow and nutrient loads were replaced by 
seasonally adjusted log transformed inflow and flow-adjusted 
log transformed nutrient loads obtained by fitting a GAM relating season 
(day of year) to log transformed daily freshwater inflow values:

$$
g(\mu) = \alpha + f_1(yday),
$${#eq-eq8}

and a GAM relating log transformed NO~3~ or TP loads to log transformed daily
inflow:

$$
g(\mu) = \alpha + f_1(log(Q)),
$${#eq-eq9}

where the response variables were modeled as normally distributed with an 
identity link function. Response residuals from the respective GAM models were 
used as *Q* and *Load* in @eq-eq6 and @eq-eq7.

This study used an information theoretic approach to evaluate 
if nutrient loads and/or freshwater inflows provided evidence of effects on water quality concentrations in Lavaca Bay. Model probabilities were calculated and compared using the AIC~c~ scores between each group of temporal, flow, and flow+load models [@burnhamAICModelSelection2011]. Improvements in model probabilities provide evidence that the terms explain additional variation in the response variable. If model probabilities were tied, there wasn't evidence the more complicated model explains additional variation in water quality.

# Results

## Watershed Nutrient Loads

Using evaluation criteria from @moriasiHydrologicWaterQuality2015, predictive 
performance of nutrient loading GAMs ranged from "satisfactory" to
"very good" based on median NSE, r^2^, and PBIAS metrics calculated using 5-fold 
cross validation. Median NSE values were 0.34 (Lavaca River) and 
0.48 (Navidad River) for NO~3~ loads and 0.81 (Lavaca River) and 0.91 (Navidad 
River) for TP loads. Median r^2^ values were 0.70
(Lavaca River) and 0.87 (Navidad River) for NO~3~ loads and 0.93 (Lavaca River) 
and 0.99 (Navidad River) for TP loads. Median PBIAS values were 2.00
(Lavaca River) and 10.90 (Navidad River) for NO~3~ loads and -7.20 
(Lavaca River) and -3.30 (Navidad River) for TP loads. Density plots of
metrics show similar distribution of values between sites for the 
same parameter, with the exception r^2^ values for NO~3~ loads where 
Lavaca River had a much larger variance in values compared to the Navidad River
(@fig-fig2). TP GAMS had higher average NSE and r^2^ values and 
lower variance in metric values compared to NO~3~.

```{r}
#| label: fig-fig2
#| warning: false
#| message: false
#| fig-width: 5.2
#| fig-asp: 0.65
#| out-width: "5.2in"
#| fig-cap: "Density plots of goodness-of-fit metrics (NSE, r^2^, and PBIAS) from repeated 5-fold cross validation between predicted nutrient loads from GAM models and measured nutrient loads. Color indicates the tail probability calcualted from the empirical cumulative distribution of the goodness-of-fit metrics."

## some functions to overide scales in facets via dewey dunnington
#https://dewey.dunnington.ca/post/2018/modifying-facet-scales-in-ggplot2/
scale_override <- function(which, scale) {
  if(!is.numeric(which) || (length(which) != 1) || (which %% 1 != 0)) {
    stop("which must be an integer of length 1")
  }

  if(is.null(scale$aesthetics) || !any(c("x", "y") %in% scale$aesthetics)) {
    stop("scale must be an x or y position scale")
  }

  structure(list(which = which, scale = scale), class = "scale_override")
}

CustomFacetWrap <- ggproto(
  "CustomFacetWrap", FacetWrap,
  init_scales = function(self, layout, x_scale = NULL, y_scale = NULL, params) {
    # make the initial x, y scales list
    scales <- ggproto_parent(FacetWrap, self)$init_scales(layout, x_scale, y_scale, params)

    if(is.null(params$scale_overrides)) return(scales)

    max_scale_x <- length(scales$x)
    max_scale_y <- length(scales$y)

    # ... do some modification of the scales$x and scales$y here based on params$scale_overrides
    for(scale_override in params$scale_overrides) {
      which <- scale_override$which
      scale <- scale_override$scale

      if("x" %in% scale$aesthetics) {
        if(!is.null(scales$x)) {
          if(which < 0 || which > max_scale_x) stop("Invalid index of x scale: ", which)
          scales$x[[which]] <- scale$clone()
        }
      } else if("y" %in% scale$aesthetics) {
        if(!is.null(scales$y)) {
          if(which < 0 || which > max_scale_y) stop("Invalid index of y scale: ", which)
          scales$y[[which]] <- scale$clone()
        }
      } else {
        stop("Invalid scale")
      }
    }

    # return scales
    scales
  }
)

facet_wrap_custom <- function(..., scale_overrides = NULL) {
  # take advantage of the sanitizing that happens in facet_wrap
  facet_super <- facet_wrap(...)

  # sanitize scale overrides
  if(inherits(scale_overrides, "scale_override")) {
    scale_overrides <- list(scale_overrides)
  } else if(!is.list(scale_overrides) || 
            !all(vapply(scale_overrides, inherits, "scale_override", FUN.VALUE = logical(1)))) {
    stop("scale_overrides must be a scale_override object or a list of scale_override objects")
  }

  facet_super$params$scale_overrides <- scale_overrides

  ggproto(NULL, CustomFacetWrap,
          shrink = facet_super$shrink,
          params = facet_super$params
  )
}

df <- tar_read(cv_no3_08164000, store = store) |> 
  mutate(site = "USGS-08164000",
         parameter = "NO <sub>3</sub>") |> 
  bind_rows(
    tar_read(cv_tp_08164000, store = store) |> 
      mutate(site = "USGS-08164000",
             parameter = "TP")
  ) |> 
  bind_rows(
    tar_read(cv_no3_texana, store = store) |> 
      mutate(site = "USGS-08164525",
             parameter = "NO <sub>3</sub>")
  ) |> 
  bind_rows(
    tar_read(cv_tp_texana, store = store) |> 
      mutate(site = "USGS-08164525",
             parameter = "TP")
  ) |> 
  mutate(site = case_when(
    site == "USGS-08164525" ~ "Navidad",
    site == "USGS-08164000" ~ "Lavaca"
  ))

df_long <- df |> 
  ungroup() |> 
  select(NSE, r2, pbias, site, parameter) |> 
  pivot_longer(cols = c(NSE, r2, pbias),
               names_to = "metric")

## plots the density estimates of the repeated 5-fold cross-validation goodness-of-fit metric results,
## color indicates the tail probability calculated from the empirical cumulitive distribution of the goodness-of-fit metric values 
p1 <- ggplot(df, aes(y = site, x = NSE,
               fill = 0.5 - abs(0.5 - after_stat(ecdf)))) +
  stat_density_ridges(geom = "density_ridges_gradient",
                      calc_ecdf = TRUE,
                      n = 10000) +
  scale_fill_scico("Tail Probability", palette = "hawaii", direction = -1,
                   limit = c(0, 0.5), breaks = c(0, 0.1, 0.2, 0.3, 0.4)) +
  guides(fill = guide_colorbar(barwidth = unit(100L, "pt"),
                               title.position = "top")) +
  facet_wrap_custom(~parameter, scales = "free_x",
                    ncol = 2,
                    scale_overrides = list(
                      scale_override(1, scale_x_continuous(limits = c(-1,1),
                                                           expand = c(0,0),
                                                           breaks = c(-1,0,1))),
                      scale_override(2, scale_x_continuous(limits = c(0,1),
                                                           expand = c(0,0),
                                                           breaks = c(0,0.5,1)))
                    )) +
  labs(x = "NSE", y = "") +
  theme_TWRI_print(base_family = "Arial") +
  theme(axis.text = element_text(size = 8),
        axis.title.x = element_markdown(size = 8),
        strip.text.x = element_markdown(size = 8),
        strip.background = element_rect(fill = "white"),
        panel.spacing.x = unit(15L, "pt"),
        legend.direction = "horizontal",
        legend.text = element_text(size = 8),
        legend.title = element_text(size = 8),
        plot.margin = unit(c(1L,0L,1L,0L), "pt"))

p2 <- ggplot(df, aes(y = site, x = r2,
               fill = 0.5 - abs(0.5 - after_stat(ecdf)))) +
  stat_density_ridges(geom = "density_ridges_gradient",
                      calc_ecdf = TRUE,
                      n = 10000) +
  scale_fill_scico("Tail Probability", palette = "hawaii", direction = -1,
                   limit = c(0, 0.5), breaks = c(0, 0.1, 0.2, 0.3, 0.4)) +
  guides(fill = guide_colorbar(barwidth = unit(100L, "pt"),
                               title.position = "top")) +
  facet_wrap_custom(~parameter, scales = "free_x",
                    ncol = 2,
                    scale_overrides = list(
                      scale_override(1, scale_x_continuous(limits = c(0,1),
                                                           expand = c(0,0),
                                                           breaks = c(0,0.5,1))),
                      scale_override(2, scale_x_continuous(limits = c(0,1),
                                                           expand = c(0,0),
                                                           breaks = c(0,0.5,1)))
                    )) +
  labs(x = "r<sup>2</sup>", y = "") +
  theme_TWRI_print(base_family = "Arial") +
  theme(axis.text = element_text(size = 8),
        axis.title.x = element_markdown(size = 8),
        strip.text.x = element_markdown(size = 8),
        strip.background = element_rect(fill = "white"),
        panel.spacing.x = unit(15L, "pt"),
        legend.direction = "horizontal",
        legend.text = element_text(size = 8),
        legend.title = element_text(size = 8),
        plot.margin = unit(c(0L,0L,1L,0L), "pt"))

p3 <- ggplot(df, aes(y = site, x = pbias,
               fill = 0.5 - abs(0.5 - after_stat(ecdf)))) +
  stat_density_ridges(geom = "density_ridges_gradient",
                      calc_ecdf = TRUE,
                      n = 10000) +
  scale_fill_scico("Tail Probability", palette = "hawaii", direction = -1,
                   limit = c(0, 0.5), breaks = c(0, 0.1, 0.2, 0.3, 0.4)) +
  guides(fill = guide_colorbar(barwidth = unit(100L, "pt"),
                               title.position = "top")) +
  facet_wrap_custom(~parameter, scales = "free_x",
                    ncol = 2,
                    scale_overrides = list(
                      scale_override(1, scale_x_continuous(limits = c(-150,200),
                                                           expand = c(0,0))),
                      scale_override(2, scale_x_continuous(limits = c(-75,75),
                                                           expand = c(0,0)))
                    )) +
  labs(x = "PBIAS", y = "") +
  theme_TWRI_print(base_family = "Arial") +
  theme(axis.text = element_text(size = 8),
        axis.title.x = element_markdown(size = 8),
        strip.text.x = element_markdown(size = 8),
        strip.background = element_rect(fill = "white"),
        panel.spacing.x = unit(15L, "pt"),
        legend.direction = "horizontal",
        legend.text = element_text(size = 8),
        legend.title = element_text(size = 8),
        axis.text.y = element_blank(),
        plot.margin = unit(c(0L,0.25,1L,0L), "pt"))

layout <- "
AAAA####
AAAA#DD#
AAAA####
BBBBCCCC
BBBBCCCC
BBBBCCCC
"

p1 / p2 / p3 / guide_area() + plot_layout(guides = 'collect',
                                          design = layout)

```

```{r echo=FALSE}
## calculates min and max load values and years
load_lav_no3 <- tar_read(daily_no3_08164000, store = store)
load_nav_no3 <- tar_read(daily_no3_texana, store = store)

load_lav_no3$annually |> 
  mutate(site = "Lavaca River") -> lavaca_no3_annually

load_nav_no3$annually |> 
  mutate(site = "Navidad River") -> navidad_no3_annually

mean_no3 <- scales::label_comma()(mean(lavaca_no3_annually$NO3_Estimate + navidad_no3_annually$NO3_Estimate ))
mean_no3_low <- scales::label_comma()(mean(lavaca_no3_annually$NO3_Lower + navidad_no3_annually$NO3_Lower ))
mean_no3_upper <- scales::label_comma()(mean(lavaca_no3_annually$NO3_Upper + navidad_no3_annually$NO3_Upper ))

no3 <- data.frame(lavaca = lavaca_no3_annually$NO3_Estimate,
                  navidad = navidad_no3_annually$NO3_Estimate,
                  year = lavaca_no3_annually$year)
no3 <- no3 |> 
  mutate(total = lavaca+navidad)

min_no3 <- min(no3$total)
min_no3_year <- no3[no3$total == min_no3,]$year
min_no3 <- scales::label_comma()(min_no3)

max_no3 <- max(no3$total)
max_no3_year <- no3[no3$total == max_no3,]$year
max_no3 <- scales::label_comma()(max_no3)

no3_prop <- no3 |> 
  mutate(prop = round(navidad/total, 2))

no3_mean_prop <- scales::label_percent()(mean(no3_prop$prop))

min_no3_prop <- scales::label_percent()(min(no3_prop$prop))

load_lav_tp <- tar_read(daily_tp_08164000, store = store)
load_nav_tp <- tar_read(daily_tp_texana, store = store)

load_lav_tp$annually |> 
  mutate(site = "Lavaca River") |> 
  filter(year >= 2005)-> lavaca_tp_annually

load_nav_tp$annually |> 
  mutate(site = "Navidad River") |> 
  filter(year >= 2005) -> navidad_tp_annually

mean_tp <- scales::label_comma()(mean(lavaca_tp_annually$TP_Estimate + navidad_tp_annually$TP_Estimate ))
mean_tp_low <- scales::label_comma()(mean(lavaca_tp_annually$TP_Lower + navidad_tp_annually$TP_Lower ))
mean_tp_upper <- scales::label_comma()(mean(lavaca_tp_annually$TP_Upper + navidad_tp_annually$TP_Upper ))

tp <- data.frame(lavaca = lavaca_tp_annually$TP_Estimate,
                  navidad = navidad_tp_annually$TP_Estimate,
                  year = lavaca_tp_annually$year)
tp <- tp |> 
  mutate(total = lavaca+navidad)

min_tp <- min(tp$total)
min_tp_year <- tp[tp$total == min_tp,]$year
min_tp <- scales::label_comma()(min_tp)

max_tp <- max(tp$total)
max_tp_year <- tp[tp$total == max_tp,]$year
max_tp <- scales::label_comma()(max_tp)

tp_prop <- tp |> 
  mutate(prop = round(navidad/total, 2))

tp_mean_prop <- scales::label_percent()(mean(tp_prop$prop))

min_tp_prop <- scales::label_percent()(min(tp_prop$prop))
```

Predicted annual NO~3~ and TP loads show considerable variation, generally following
patterns in discharge (@fig-fig3; @fig-fig4). 
Flow-normalized TP loads at both sites and flow-normalized NO~3~ loads in the Lavaca River
indicated watershed based loads did not change much over time
when accounting for variation driven by streamflow (@fig-fig3). 
Flow-normalized loads in the Lavaca River showed small variation over time with 
some decreases in NO~3~ loads since 2013.

```{r}
#| label: fig-fig3
#| error: false
#| warning: false
#| message: false
#| fig-width: 5.2
#| fig-asp: 0.7
#| out-width: "5.2in"
#| fig-cap: "Aggregated estimated annual and flow-normalized annual NO~3~ and TP loads for USGS-08164000 and USGS-08164525."
load <- tar_read(daily_no3_08164000, store = store)
fn_load <- tar_read(daily_no3_08164000_fn, store = store)

load_lav_tp <- tar_read(daily_tp_08164000, store = store)
load_nav_tp <- tar_read(daily_tp_texana, store = store)

a <- ggplot() +
  geom_point(data = load$annually, aes(year, NO3_Estimate,
                                   color = "Total Annual Load + 90% CI",
                                   shape = "Total Annual Load + 90% CI")) +
  geom_line(data = load$annually, aes(x = year, y = NO3_Estimate,
                                  color = "Total Annual Load + 90% CI",
                                  linetype = "Total Annual Load + 90% CI"),
            alpha = 0.5) +
  geom_linerange(data = load$annually, aes(x = year, ymin = NO3_Lower, ymax = NO3_Upper,
                                       color = "Total Annual Load + 90% CI")) +
  geom_point(data = fn_load$annually, aes(year, NO3_Estimate,
                                 color = "Flow-Normalized Annual Load + 90% CI",
                                 shape = "Flow-Normalized Annual Load + 90% CI")) +
  geom_line(data = fn_load$annually,
            aes(x = year, y = NO3_Estimate,
                color = "Flow-Normalized Annual Load + 90% CI",
                linetype = "Flow-Normalized Annual Load + 90% CI"),
            alpha = 0.5) +
  geom_linerange(data = fn_load$annually,
                 aes(x = year, ymin = NO3_Lower, ymax = NO3_Upper,
                     color = "Flow-Normalized Annual Load + 90% CI")) +
  labs(x = "", 
       y = expression(Annual~NO[3]~Load~"[kg]"), 
       subtitle = "Lavaca River") +
  scale_shape_manual(name = "values",
                     values = c(21, 19)) +
  scale_color_manual(name = "values",
                     values = c("#7E1900", "#1A3399")) +
  scale_linetype_manual(name = "values",
                        values = c(1, 2)) +
  scale_y_continuous(trans = "pseudo_log",
                     breaks = c(0,1E1, 1E2, 1E3, 1E4, 1E5, 1E6, 1E7),
                     labels = scales::label_log()) +
  coord_cartesian(ylim = c(1000, 1E7)) +
  theme_TWRI_print(base_family = "sans") +
  theme(axis.title.y = element_text(size = 8),
        axis.text.x = element_blank(),
        plot.subtitle = element_text(size = 8, hjust = 0.5, face = "bold"),
        panel.grid.major.x = element_line(color = "#d9d9d9",
                                          linetype = "dotted"),
        legend.title = element_blank(),
        legend.text = element_text(size = 8),
        plot.margin = unit(c(0L,3L,0L,0L), "pt"))

load <- tar_read(daily_no3_texana, store = store)
fn_load <- tar_read(daily_no3_texana_fn, store = store)

b <- ggplot() +
  geom_point(data = load$annually, aes(year, NO3_Estimate,
                                   color = "Total Annual Load + 90% CI",
                                   shape = "Total Annual Load + 90% CI")) +
  geom_line(data = load$annually, aes(x = year, y = NO3_Estimate,
                                  color = "Total Annual Load + 90% CI",
                                  linetype = "Total Annual Load + 90% CI"),
            alpha = 0.5) +
  geom_linerange(data = load$annually, aes(x = year, ymin = NO3_Lower, ymax = NO3_Upper,
                                       color = "Total Annual Load + 90% CI")) +
  geom_point(data = fn_load$annually, aes(year, NO3_Estimate,
                                 color = "Flow-Normalized Annual Load + 90% CI",
                                 shape = "Flow-Normalized Annual Load + 90% CI")) +
  geom_line(data = fn_load$annually,
            aes(x = year, y = NO3_Estimate,
                color = "Flow-Normalized Annual Load + 90% CI",
                linetype = "Flow-Normalized Annual Load + 90% CI"),
            alpha = 0.5) +
  geom_linerange(data = fn_load$annually,
                 aes(x = year, ymin = NO3_Lower, ymax = NO3_Upper,
                     color = "Flow-Normalized Annual Load + 90% CI")) +
  labs(x = "", 
       y = expression(Annual~NO[3]~Load~"[kg]"), 
       subtitle = "Navidad River") +
  scale_shape_manual(name = "values",
                     values = c(21, 19)) +
  scale_color_manual(name = "values",
                     values = c("#7E1900", "#1A3399")) +
  scale_linetype_manual(name = "values",
                        values = c(1, 2)) +
  scale_y_continuous(trans = "pseudo_log",
                     breaks = c(0,1E1, 1E2, 1E3, 1E4, 1E5, 1E6, 1E7),
                     labels = scales::label_log()) +
  coord_cartesian(ylim = c(1000, 1E7)) +
  theme_TWRI_print(base_family = "sans") +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.text.x = element_blank(),
        plot.subtitle = element_text(size = 8, hjust = 0.5, face = "bold"),
        panel.grid.major.x = element_line(color = "#d9d9d9",
                                          linetype = "dotted"),
        legend.title = element_blank(),
        legend.text = element_text(size = 8),
        plot.margin = unit(c(0L,0L,0L,3L), "pt"))

load <- tar_read(daily_tp_08164000, store = store)
fn_load <- tar_read(daily_tp_08164000_fn, store = store)

load$annually <- load$annually |> 
  filter(year >= 2005)

fn_load$annually <- fn_load$annually |> 
  filter(year >= 2005)

c <- ggplot() +
  geom_point(data = load$annually, aes(year, TP_Estimate,
                                   color = "Total Annual Load + 90% CI",
                                   shape = "Total Annual Load + 90% CI")) +
  geom_line(data = load$annually, aes(x = year, y = TP_Estimate,
                                  color = "Total Annual Load + 90% CI",
                                  linetype = "Total Annual Load + 90% CI"),
            alpha = 0.5) +
  geom_linerange(data = load$annually, aes(x = year, ymin = TP_Lower, ymax = TP_Upper,
                                       color = "Total Annual Load + 90% CI")) +
  geom_point(data = fn_load$annually, aes(year, TP_Estimate,
                                 color = "Flow-Normalized Annual Load + 90% CI",
                                 shape = "Flow-Normalized Annual Load + 90% CI")) +
  geom_line(data = fn_load$annually,
            aes(x = year, y = TP_Estimate,
                color = "Flow-Normalized Annual Load + 90% CI",
                linetype = "Flow-Normalized Annual Load + 90% CI"),
            alpha = 0.5) +
  geom_linerange(data = fn_load$annually,
                 aes(x = year, ymin = TP_Lower, ymax = TP_Upper,
                     color = "Flow-Normalized Annual Load + 90% CI")) +
  labs(x = "", y = "Annual TP Load [kg]", subtitle = "Lavaca River") +
  scale_shape_manual(name = "values",
                     values = c(21, 19)) +
  scale_color_manual(name = "values",
                     values = c("#7E1900", "#1A3399")) +
  scale_linetype_manual(name = "values",
                        values = c(1, 2)) +
  scale_y_continuous(trans = "pseudo_log",
                     breaks = c(0,1E1, 1E2, 1E3, 1E4, 1E5, 1E6, 1E7),
                     labels = scales::label_log()) +
  coord_cartesian(ylim = c(1000, 1E6)) +
  theme_TWRI_print(base_family = "sans") +
  theme(axis.title.y = element_text(size = 8),
        panel.grid.major.x = element_line(color = "#d9d9d9",
                                          linetype = "dotted"),
        legend.title = element_blank(),
        legend.text = element_text(size = 8),
        plot.margin = unit(c(0L,3L,0L,0L), "pt"),
        plot.subtitle = element_blank())

load <- tar_read(daily_tp_texana, store = store) 
fn_load <- tar_read(daily_tp_texana_fn, store = store)

load$annually <- load$annually |> 
  filter(year >= 2005)

fn_load$annually <- fn_load$annually |> 
  filter(year >= 2005)

d <- ggplot() +
  geom_point(data = load$annually, aes(year, TP_Estimate,
                                   color = "Total Annual Load + 90% CI",
                                   shape = "Total Annual Load + 90% CI")) +
  geom_line(data = load$annually, aes(x = year, y = TP_Estimate,
                                  color = "Total Annual Load + 90% CI",
                                  linetype = "Total Annual Load + 90% CI"),
            alpha = 0.5) +
  geom_linerange(data = load$annually, aes(x = year, ymin = TP_Lower, ymax = TP_Upper,
                                       color = "Total Annual Load + 90% CI")) +
  geom_point(data = fn_load$annually, aes(year, TP_Estimate,
                                 color = "Flow-Normalized Annual Load + 90% CI",
                                 shape = "Flow-Normalized Annual Load + 90% CI")) +
  geom_line(data = fn_load$annually,
            aes(x = year, y = TP_Estimate,
                color = "Flow-Normalized Annual Load + 90% CI",
                linetype = "Flow-Normalized Annual Load + 90% CI"),
            alpha = 0.5) +
  geom_linerange(data = fn_load$annually,
                 aes(x = year, ymin = TP_Lower, ymax = TP_Upper,
                     color = "Flow-Normalized Annual Load + 90% CI")) +
  labs(x = "", y = "Annual TP Load [kg]", subtitle = "Navidad River") +
  scale_shape_manual(name = "values",
                     values = c(21, 19)) +
  scale_color_manual(name = "values",
                     values = c("#7E1900", "#1A3399")) +
  scale_linetype_manual(name = "values",
                        values = c(1, 2)) +
  scale_y_continuous(trans = "pseudo_log",
                     breaks = c(0,1E1, 1E2, 1E3, 1E4, 1E5, 1E6, 1E7),
                     labels = scales::label_log()) +
  coord_cartesian(ylim = c(1000, 1E6)) +
  theme_TWRI_print(base_family = "sans") +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        panel.grid.major.x = element_line(color = "#d9d9d9",
                                          linetype = "dotted"),
        legend.title = element_blank(),
        legend.text = element_text(size = 8),
        plot.margin = unit(c(0L,1L,0L,3L), "pt"),
        plot.subtitle = element_blank())

layout <- "
11112222
11112222
11112222
11112222
33334444
33334444
33334444
33334444
55555555
"

a + b + c + d + guide_area() + plot_layout(guides = 'collect', design = layout)
```

Aggregated across both sites, the mean 
annual NO~3~ load 2005 through 2020 was `r mean_no3` kg (`r mean_no3_low` kg - 
`r mean_no3_upper` kg, 90% CI). Annual NO~3~ loads ranged from `r min_no3` kg 
in `r min_no3_year` to `r max_no3` kg in `r max_no3_year`. Total annual TP loads 
ranged from `r min_tp` kg in `r min_tp_year` to 
`r max_tp` kg  in `r max_tp_year`. Mean annual TP loading from 2005 through 2020
was `r mean_tp` kg (`r mean_tp_low` kg - `r mean_tp_upper` kg, 90% CI).
On average, the Navidad River accounted for `r no3_mean_prop` of NO~3~ loads and 
`r tp_mean_prop` of TP loads from 2005 through 2020. However, during periods of extreme 
drought the Lavaca River became the primary source of 
nutrient loading in the watershed with the Navidad River only accounting for 
`r min_no3_prop` and `r min_tp_prop` of NO~3~ and TP loads in 2011 (@fig-fig4).

```{r}
#| label: fig-fig4
#| error: false
#| message: false
#| warning: false
#| fig-width: 3
#| fig-asp: 2
#| fig-cap: "Comparison of delivered annual loads and annual discharge at the Lavaca (USGS-08164000) and Navidad (USGS-08164525) Rivers."
#| out-width: "3in"
load_lav_no3 <- tar_read(daily_no3_08164000, store = store)
load_nav_no3 <- tar_read(daily_no3_texana, store = store)

load_lav_no3$annually |> 
  mutate(site = "Lavaca River") -> lavaca_no3_annually

load_nav_no3$annually |> 
  mutate(site = "Navidad River") -> navidad_no3_annually

no3_annual <- bind_rows(lavaca_no3_annually, navidad_no3_annually) |> 
  mutate(x = as.Date(paste0(year, "-01-01"), "%Y-%m-%d"))

# b <- ggplot(no3_annual) +
#   geom_col(aes(year, NO3_Estimate, fill = site), width = 0.8) +
#   scale_x_continuous(expand = expansion(mult = c(0.05, 0.05)), 
#                      breaks = c(2005, 2010, 2015, 2020)) +
#   scale_y_continuous(labels = scales::comma) +
#   scale_fill_carto_d(palette = "Vivid", direction = -1) +
#   labs(x = "", 
#        y = "Predicted Annual<br>NO <sub>3</sub> Load [kg]") +
#   theme_TWRI_print(base_family = "Arial") +
#   theme(axis.title.y = element_text(size = 8),
#         axis.text = element_text(size = 8),
#         panel.grid.major.x = element_line(color = "#d9d9d9",
#                                           linetype = "dotted"),
#         legend.title = element_blank(),
#         legend.text = element_text(size = 8))

prop <- no3_annual |> 
  mutate(proportion = NO3_Estimate/sum(NO3_Estimate))

ylab1 <- expression(Annual~NO[3]~Load)
ylab2 <- expression("[Proportion]")
c <- ggplot() +
  geom_col(data = prop, aes(year, proportion, fill  = site)) +
  scale_fill_carto_d(palette = "Vivid", direction = -1) +
  scale_x_continuous(breaks = c(2005, 2010, 2015, 2020),
                     expand = expansion(mult = c(0.05, 0.05))) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = "",
       y = "")  +
  coord_cartesian(clip = "off") +
  annotation_custom(grid::textGrob(ylab1,
                                   x = unit(-0.23, "npc"),
                                   rot = 90,
                                   gp = grid::gpar(fontfamily = "sans",
                                             fontsize = 8))) +
  annotation_custom(grid::textGrob(ylab2,
                                   x = unit(-0.18, "npc"),
                                   rot = 90,
                                   gp = grid::gpar(fontfamily = "sans",
                                                   fontsize = 8))) +
  theme_TWRI_print(base_family = "sans") +
  theme(axis.title.y = element_text(size = 8),
        axis.text = element_text(size = 8),
        panel.grid.major.x = element_line(color = "#d9d9d9",
                                          linetype = "dotted"),
        legend.title = element_blank(),
        legend.text = element_text(size = 8))

ylab1 <- expression("Annual Discharge")
ylab2 <- expression("[Million Gallons]")
d <- tar_read(qdata, store = store) |>
  filter(site_no %in% c("8164000", "lktexana_g")) |>
  filter(Date >= as.Date("2005-01-01")) |>
  mutate(site = case_when(
    site_no == "8164000" ~ "Lavaca River",
    site_no == "lktexana_g" ~ "Navidad River"
  )) |>
  mutate(year = year(Date)) |>

  mutate(Flow = as_units(Flow, "ft^3/s")) |>
  mutate(Flow = set_units(Flow, "ft^3/day")) |>
  mutate(Flow = set_units(Flow, "1E6gallons/day")) |>
  group_by(year, site) |>
  summarise(Flow = sum(Flow, na.rm = TRUE)) |>
  mutate(Flow = drop_units(Flow)) |>
  ggplot() +
  geom_col(aes(year, Flow, fill = site)) +
  scale_fill_carto_d(palette = "Vivid", direction = -1) +
  scale_x_continuous(breaks = c(2005, 2010, 2015, 2020),
                     expand = expansion(mult = c(0.05, 0.05))) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = "",
       y = "") +
  coord_cartesian(clip = "off") +
  annotation_custom(grid::textGrob(ylab1,
                                   x = unit(-0.25, "npc"),
                                   rot = 90,
                                   gp = grid::gpar(fontfamily = "sans",
                                             fontsize = 8))) +
  annotation_custom(grid::textGrob(ylab2,
                                   x = unit(-0.20, "npc"),
                                   rot = 90,
                                   gp = grid::gpar(fontfamily = "sans",
                                                   fontsize = 8))) +
  theme_TWRI_print(base_family = "Arial") +
  theme(axis.title.y = element_text(size = 8),
        axis.text = element_text(size = 8),
        panel.grid.major.x = element_line(color = "#d9d9d9",
                                          linetype = "dotted"),
        legend.title = element_blank(),
        legend.text = element_text(size = 8),
        legend.position = "none",
        plot.margin = unit(c(1L,1L,1L,2L), "pt"))

load_lav_tp$annually |>
  mutate(site = "Lavaca River") -> lavaca_tp_annually

load_nav_tp$annually |>
  mutate(site = "Navidad River") -> navidad_tp_annually

tp_annual <- bind_rows(lavaca_tp_annually, navidad_tp_annually) |>
  filter(year >= 2005) |>
  mutate(x = as.Date(paste0(year, "-01-01"), "%Y-%m-%d"))
# 
# e <- ggplot(tp_annual) +
#   geom_col(aes(year, TP_Estimate, fill = site), width = 0.8) +
#   scale_x_continuous(expand = expansion(mult = c(0.05, 0.05)), 
#                      breaks = c(2005, 2010, 2015, 2020)) +
#   scale_y_continuous(labels = scales::comma) +
#   scale_fill_carto_d(palette = "Vivid", direction = -1) +
#   labs(x = "", y = "Predicted Annual<br>TP Load [kg]") +
#   theme_TWRI_print(base_family = "Arial") +
#   theme(axis.title.y = element_text(size = 8),
#         axis.text = element_text(size = 8),
#         panel.grid.major.x = element_line(color = "#d9d9d9",
#                                           linetype = "dotted"),
#         legend.title = element_blank(),
#         legend.text = element_text(size = 8))
# 
# 
prop <- tp_annual |>
  mutate(proportion = TP_Estimate/sum(TP_Estimate))

ylab1 <- expression("Annual TP Load")
ylab2 <- expression("[Proportion]")
f <- ggplot() +
  geom_col(data = prop, aes(year, proportion, fill  = site)) +
  scale_fill_carto_d(palette = "Vivid", direction = -1) +
  scale_x_continuous(breaks = c(2005, 2010, 2015, 2020),
                     expand = expansion(mult = c(0.05, 0.05))) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = "",
       y = "") +
  coord_cartesian(clip = "off") +
  annotation_custom(grid::textGrob(ylab1,
                                   x = unit(-0.23, "npc"),
                                   rot = 90,
                                   gp = grid::gpar(fontfamily = "sans",
                                             fontsize = 8))) +
  annotation_custom(grid::textGrob(ylab2,
                                   x = unit(-0.18, "npc"),
                                   rot = 90,
                                   gp = grid::gpar(fontfamily = "sans",
                                                   fontsize = 8))) +
  theme_TWRI_print(base_family = "Arial") +
  theme(axis.title.y = element_text(size = 8),
        axis.text = element_text(size = 8),
        panel.grid.major.x = element_line(color = "#d9d9d9",
                                          linetype = "dotted"),
        legend.title = element_blank(),
        legend.text = element_text(size = 8),
        legend.position = "none")

layout <- "
CC
CC
CC
CC
FF
FF
FF
FF
DD
DD
DD
DD
GG
"

c + d + f + guide_area() +
  plot_layout(guides = 'collect', design  = layout)

```

## Linkages Between Water Quality and Watershed Flows and Loads

GAMs did not identify significant changes in TP or DO concentrations at any
of the Lavaca Bay sites from 2005 through 2020 (@fig-fig5). The
upper-bay site, TCEQ-13563, had a linear increase in NO~x~ concentration and
and decrease in chlorophyll-*a* from 2005 through 2014. The mid-bay site,
TCEQ-13383, showed a periodic pattern in NO~x~ concentration that appeared
similar to precipitation/inflow patterns, as well as a post 2011 increase
in TKN concentrations. No significant long-term trends in concentrations 
were identified by GAMs for the lower-bay TCEQ-13384 site.

```{r}
#| label: fig-fig5
#| message: false
#| warning: false
#| fig-width: 5.2
#| fig-asp: 0.7
#| fig-cap: "Smoothed temporal trend component for water quality paramaters obtained from temporal estuary GAMs."
#| out-width: "5.2in"
draw_smoothed_years <- function(model1, model2, model3,
                                ylab, subtitle) {

  `TCEQ-13563` <- model1
  `TCEQ-13383` <- model2
  `TCEQ-13384` <- model3

  newdata <- data_slice(model1, ddate = evenly(2005:2020, n = 200))

  comp <- gratia::compare_smooths(
    `TCEQ-13563`,
    `TCEQ-13383`,
    `TCEQ-13384`,
    smooths = "s(ddate)",
    data = newdata)

  crit <- gratia:::coverage_normal(0.90)

  comp |> 
    unnest(data) |> 
    mutate(lower_ci = est + (crit * se),
           upper_ci = est - (crit * se),
           label = if_else(ddate == max(ddate),
                           as.character(model),
                           NA_character_)) |> 
    ggplot(aes(x = ddate, y = est, group = model)) +
    geom_ribbon(aes(ymin = lower_ci,
                    ymax = upper_ci,
                    fill = model),
                alpha = 0.15) +
    geom_line(aes(color = model)) +
    geom_hline(yintercept = 0, linetype = 2, alpha = 0.25) +
    # geom_text_repel(aes(label = label,
    #                     color = model), 
    #                 nudge_x = 5,
    #                 hjust = "left",
    #                 size = 2) +
    scale_x_continuous(breaks = c(2005, 2010, 2015, 2020),
                       expand = expansion(mult = c(0.05,0.05))) +
    colorspace::scale_fill_discrete_qualitative(name = "") +
    colorspace::scale_color_discrete_qualitative(name = "") +
    labs(x = "Year", y = ylab,
         subtitle = subtitle) +
    theme_TWRI_print(base_family = "sans") +
    theme(axis.title.y = element_text(size = 8),
          axis.title.x = element_text(size = 8),
          axis.text.x = element_text(size = 8),
          axis.text.y = element_text(size = 8),
          plot.subtitle = element_text(size = 8),
          legend.direction = "vertical",
          legend.text = element_text(size = 8))

}

p_tp <- draw_smoothed_years(tar_read(tp_lavaca_13563_temporal, store = store),
                            tar_read(tp_lavaca_13383_temporal, store = store),
                            tar_read(tp_lavaca_13384_temporal, store = store),
                            ylab = "TP [mg/L]", 
                            subtitle = "A")

p_no3 <- draw_smoothed_years(tar_read(no3_lavaca_13563_temporal, store = store),
                             tar_read(no3_lavaca_13383_temporal, store = store),
                             tar_read(no3_lavaca_13384_temporal, store = store),
                             ylab = "NOx [mg/L]", 
                             subtitle = "B")

p_tkn <- draw_smoothed_years(tar_read(tkn_lavaca_13563_temporal, store = store),
                             tar_read(tkn_lavaca_13383_temporal, store = store),
                             tar_read(tkn_lavaca_13384_temporal, store = store),
                             ylab = "TKN [mg/L]", 
                             subtitle = "C")

p_chla <- draw_smoothed_years(tar_read(chla_lavaca_13563_temporal, store = store),
                             tar_read(chla_lavaca_13383_temporal, store = store),
                             tar_read(chla_lavaca_13384_temporal, store = store),
                             ylab = expression("Chlorophyll-"~italic(a)~"[μg/L]"), 
                             subtitle = "D")

p_do <- draw_smoothed_years(tar_read(do_lavaca_13563_temporal, store = store),
                             tar_read(do_lavaca_13383_temporal, store = store),
                             tar_read(do_lavaca_13384_temporal, store = store),
                             ylab = "DO [mg/L]", 
                            subtitle = "E")

design <- "
112266
334455
"

#(p_tp + p_no3) / (p_chla + p_tkn) / (p_do)
p_tp + p_no3 + p_chla + p_tkn + p_do + guide_area() + plot_layout(guides = "collect",
                                                   design = design)
```

Freshwater inflow provided additional explanation for changes in TP and 
NO~*x*~ concentration at all of the Lavaca Bay sites according to AIC~c~ and 
model probability values (@tbl-table4).
TCEQ-13563, the site closest to the river outlet, was the only site that
had improvements in the explanations of DO and TKN concentration with the inclusion 
of inflow. Both TCEQ-13563 and TCEQ-13383, the mid-bay site, saw improvements 
in explanations for variations in chlorophyll-*a* with the inclusion of freshwater
inflow. The addition of nutrient loads (both TP and NO~3~) terms did not provide
additional explanation for changes in chlorophyll-*a* or DO concentrations.
Inclusion of TP loads provided additional explanation of TP concentrations at
the upper- and mid-bay sites, TCEQ-13563 and TCEQ-13383. Inclusion of NO~3~
loads only provided marginal improvements in the explanation of NO~*X*~ 
concentration at the lower-bay TCEQ-13384 site.

```{r}
#| label: tbl-table4
#| message: false
#| warning: false
#| tbl-cap: "Estuary GAM AIC~c~ values and associated model probabilities (in parenthesis). Models with the highest probability for each site and water quality parameter combination are bolded and italicized for emphasis."

dplyr::bind_rows(
    tar_read(tp_13563, store = store), 
    tar_read(tp_13383, store = store),
    tar_read(tp_13384, store = store),
    tar_read(no3_13563, store = store), 
    tar_read(no3_13383, store = store),
    tar_read(no3_13384, store = store),
    tar_read(chla_13563, store = store), 
    tar_read(chla_13383, store = store),
    tar_read(chla_13384, store = store),
    tar_read(do_13563, store = store), 
    tar_read(do_13383, store = store),
    tar_read(do_13384, store = store),
    tar_read(tkn_13563, store = store),
    tar_read(tkn_13383, store = store),
    tar_read(tkn_13384, store = store)
    ) |> 
  mutate(Parameter = case_when(
    Parameter == "Nitrite+Nitrate" ~ "NO~*x*~",
    Parameter == "Chlorophyll-a" ~ "Chlorophyll-*a*",
    TRUE ~ as.character(Parameter)
  )) |> 
  mutate(Parameter = fct_relevel(Parameter, "TP", "NO~*x*~", "Chlorophyll-*a*","TKN", "DO"),
         Model = fct_relevel(Model, "Temporal", "Flow", "Flow + Load")) |>
  tidyr::complete(Site, Model, nesting(Parameter)) |>
  mutate(model_prob = paste0(round(AICc,1), " (", round(weight,2), ")")) |> 
  mutate(model_prob = case_when(
    model_prob == "NA (NA)" ~ "-",
    model_prob != "NA (NA)" ~ model_prob
  )) |> 
  select(-c(AICc, weight)) |> 
  pivot_wider(names_from = Model, values_from = model_prob) |> 
  select(Parameter, Site, Temporal, Flow, `Flow + Load`) |> 
  arrange(Parameter, Site) |> 
  kbl(format = "pipe")

```

GAMs showed increases in freshwater inflow resulted in nearly linear increases 
in TP and NO~*x*~ concentration at all three sites (@fig-fig6). At the upper-bay
TCEQ-13563 site, GAMs showed increases in freshwater inflow initially 
increased chlorophyll-*a* and DO concentration but concentrations leveled and
potentially decreased at higher flows. The mid-bay TCEQ-13383 site showed a
nearly linear increased in chlorophyll-*a* concentration in response to increases
freshwater inflow. Freshwater flow did not have significant
effects on chlorophyll-*a*, TKN, or DO at the lower-bay TCEQ-13384 site.

```{r}
#| label: fig-fig6
#| message: false
#| warning: false
#| fig-width: 5.2
#| fig-asp: 0.7
#| out-width: "5.2in"
#| fig-cap: "Estimated effects of mean daily inflow residuals on mean TP, NO~*x*~, chlorophyll-*a*, TKN, and DO concentrations in Lavaca Bay obtained from flow estuary GAMs."
draw_smoothed_surface <- function(comp,
                                  x = flw_res,
                                  xlab,
                                  ylab, 
                                  subtitle) {

  crit <- gratia:::coverage_normal(0.90)

  comp |>
    unnest(data) |>
    mutate(lower_ci = est + (crit * se),
           upper_ci = est - (crit * se)) |>
    group_by(model) |> 
    mutate(label = if_else(
      {{ x }} == max( {{ x }} ),
      as.character(model),
      NA_character_)) |> 
    ggplot(aes(x = {{ x }}, y = est, group = model)) +
    geom_ribbon(aes(ymin = lower_ci,
                    ymax = upper_ci,
                    fill = model),
                alpha = 0.15) +
    geom_line(aes(color = model)) +
    geom_hline(yintercept = 0, linetype = 2, alpha = 0.5) +
    # geom_text_repel(aes(label = label,
    #                     color = model),
    #                 nudge_x = 5,
    #                 hjust = "left",
    #                 size = 2) +
    scale_x_continuous(expand = expansion(mult = c(0.05,0.05))) +
    colorspace::scale_fill_discrete_qualitative(name = "") +
    colorspace::scale_color_discrete_qualitative(name = "") +
    labs(x = xlab, y = ylab,
         subtitle = subtitle) +
    theme_TWRI_print(base_family = "sans") +
    theme(axis.title.y = element_markdown(size = 8),
          axis.title.x = element_markdown(size = 8),
          plot.subtitle = element_markdown(size = 8),
          legend.direction = "vertical",
          legend.text = element_text(size = 8))

}

`TCEQ-13383` <- tar_read(tp_lavaca_13383_flow, store = store)

comp <- compare_smooths(model = `TCEQ-13383`,
                      `TCEQ-13384` = tar_read(tp_lavaca_13384_flow, store = store),
                      `TCEQ-13563` = tar_read(tp_lavaca_13563_flow, store = store),
                      smooths = c("s(flw_res)"))
p_tp <- draw_smoothed_surface(comp,
                      x = flw_res,
                      xlab = "Residuals log(Inflow) [cfs]",
                      ylab =  "TP Smooth<br>Estimate [mg/L]",
                      subtitle = "A: TP")

`TCEQ-13383` <- tar_read(no3_lavaca_13383_flow, store = store)

comp <- compare_smooths(model = `TCEQ-13383`,
                      `TCEQ-13384` = tar_read(no3_lavaca_13384_flow, store = store),
                      `TCEQ-13563` = tar_read(no3_lavaca_13563_flow, store = store),
                      smooths = c("s(flw_res)"))
p_no3 <- draw_smoothed_surface(comp,
                      x = flw_res,
                      xlab = "Residuals log(Inflow) [cfs]",
                      ylab = "NO*<sub>x</sub>* Smooth<br>Estimate [mg/L]",
                      subtitle = "B: NO*<sub>x</sub>*")

`TCEQ-13383` <- tar_read(chla_lavaca_13383_flow, store = store)

comp <- compare_smooths(model = `TCEQ-13383`,
                      `TCEQ-13384` = tar_read(chla_lavaca_13384_flow, store = store),
                      `TCEQ-13563` = tar_read(chla_lavaca_13563_flow, store = store),
                      smooths = c("s(flw_res)"))
p_chla <- draw_smoothed_surface(comp,
                      x = flw_res,
                      xlab = "Residuals log(Inflow) [cfs]",
                      ylab = "Chlorophyll-*a* Smooth<br>Estimate [µg/L]",
                      subtitle = "C: Chlorophyll-*a*")

`TCEQ-13383` <- tar_read(tkn_lavaca_13383_flow, store = store)

comp <- compare_smooths(model = `TCEQ-13383`,
                      `TCEQ-13384` = tar_read(tkn_lavaca_13384_flow, store = store),
                      `TCEQ-13563` = tar_read(tkn_lavaca_13563_flow, store = store),
                      smooths = c("s(flw_res)"))
p_tkn <- draw_smoothed_surface(comp,
                      x = flw_res,
                      xlab = "Residuals log(Inflow) [cfs]",
                      ylab = "TKN Smooth<br>Estimate [mg/L]",
                      subtitle = "D: TKN")

`TCEQ-13383` <- tar_read(do_lavaca_13383_flow, store = store)

comp <- compare_smooths(model = `TCEQ-13383`,
                      `TCEQ-13384` = tar_read(do_lavaca_13384_flow, store = store),
                      `TCEQ-13563` = tar_read(do_lavaca_13563_flow, store = store),
                      smooths = c("s(flw_res)"))
p_do <- draw_smoothed_surface(comp,
                      x = flw_res,
                      xlab = "Residuals log(Inflow) [cfs]",
                      ylab = "DO Smooth<br>Estimate [mg/L]",
                      subtitle = "E: DO")

design <- "
112266
334455
"

p_tp + p_no3 + p_chla + p_tkn + p_do + guide_area() + plot_layout(guides = "collect",
                                                   design = design)

```

Increased TP loads resulted in nearly linear increases of TP concentration at
the upper- and mid-bay sites, TCEQ-13563 and TCEQ-13383 respectively (@fig-fig7).
The relative effect size appeared to much smaller than the effect of freshwater
inflow alone. Increased NO~3~ loads only showed an effect at the 
lower-bay TCEQ-13384 site. The effect was quite small compared to streamflow and
provided only small improvements to the model (@tbl-table4). As noted
above, nutrient loadings did not provide any explanation in changes in the
remaining assessed water quality parameters.

```{r}
#| label: fig-fig7
#| message: false
#| warning: false
#| fig-width: 5.2
#| fig-asp: 0.45
#| out-width: "5.2in"
#| fig-cap: "Estimated effects of nutrient load residuals on TP and NO~*x*~ concentrations in Lavaca Bay obtained from flow+load estuary GAMs."

`TCEQ-13383` <- tar_read(tp_lavaca_13383_full, store = store)

comp <- compare_smooths(model = `TCEQ-13383`,
                      `TCEQ-13384` = tar_read(tp_lavaca_13384_full, store = store),
                      `TCEQ-13563` = tar_read(tp_lavaca_13563_full, store = store),
                      smooths = c("s(TP_resid)"))
p_tp <- draw_smoothed_surface(comp,
                      x = TP_resid,
                      xlab = "Residuals log(TP) Load [kg]",
                      ylab = "TP [mg/L]",
                      subtitle = "A: TP")

`TCEQ-13383` <- tar_read(no3_lavaca_13383_full, store = store)

comp <- compare_smooths(model = `TCEQ-13383`,
                      `TCEQ-13384` = tar_read(no3_lavaca_13384_full, store = store),
                      `TCEQ-13563` = tar_read(no3_lavaca_13563_full, store = store),
                      smooths = c("s(NO3_resid)"))
p_no3 <- draw_smoothed_surface(comp,
                      x = NO3_resid,
                      xlab = "Residuals log(NO<sub>3</sub>) Load [kg]",
                      ylab = "NO*<sub>x</sub>* [mg/L]",
                      subtitle = "B: NO*<sub>x</sub>*")

design <- "
11223
"

p_tp + p_no3 + guide_area() + plot_layout(guides = "collect",
                                                   design = design)

```

# Discussion

TP and NO~3~ loadings from the Lavaca Bay watershed showed high inter-annual
variability tied with changes in discharge. There is little evidence for changes
in flow-normalized TP loads in either rivers. There is some evidence of recent
decreases in flow-normalized NO~3~ loads in the Lavaca River. Although there is
no work directly correlating water quality planning and implementation efforts
in the watershed to water quality outcomes, efforts to increase agricultural 
producer participation in the watershed have been ongoing since 2016
[@schramm_lavaca_2018; @bertholdDirectMailingEducation2021]. The decrease in 
flow-normalized NO~3~ loads could be a reflection of those collective efforts 
but further data collection and research is required to support that statement.

Converted to average annual yield, the estimates of annual TP loads for the
Lavaca River are within the ranges in previous published studies
(@tbl-table5; [@dunnTrendsNutrientInflows1996; @rebich_sources_2011; @omaniEstimationSedimentNutrient2014; @wise_spatially_2019]). It isn't obvious
why TP estimates in @dunnTrendsNutrientInflows1996 were notably lower. Given that none
of the studies identify substantially sized trends in TP, it is is possible that 
the period used in @dunnTrendsNutrientInflows1996 was drier on average than the
other studies. The SPARROW models used in @rebich_sources_2011 and 
@wise_spatially_2019 utilize a version of LOADEST in the underlying load
estimation procedure, so a difference due to methodology alone is unlikely.

```{r}
#| label: tbl-table5
#| warning: false
#| message: false
#| tbl-cap: "Comparisons of previously published estimates of mean annual TP yield at the Lavaca River site."
#| results: asis

load_lav_tp <- tar_read(daily_tp_08164000, store = store)
load_nav_tp <- tar_read(daily_tp_texana, store = store)

dat <- load_lav_tp$annually |> 
  ungroup() |> 
  filter(year >= 2005) |> 
  summarise(annual = round(mean(TP_Estimate)/2116,1),
            upper = round(mean(TP_Upper)/2116,1),
            lower = round(mean(TP_Lower)/2116,1)) ## 2116 sq km

df <- tibble(
  Parameter = rep("TP", 5),
  "Reported Yield (kg ⋅ km^-2^ ⋅ year^-1^)" = 
    c(paste0(dat[[1,1]]," (", dat[[1,3]], ", ", dat[[1,2]], ")^a^"), 
      "45.2", "42", "20.81-91.58^b^", "28.9"),
  Approach = c("GAM", "SPARROW", "SWAT", "SPARROW", "LOADEST"),
  `Time Period` = c("2005-2020", "2000-2014", "1977-2005", "1980-2002", "1972-1993"),
  Reference = c("This work", "@wise_spatially_2019",
                "@omaniEstimationSedimentNutrient2014",
                "@rebich_sources_2011",
                "@dunnTrendsNutrientInflows1996")
)

x <- kable(df, format = "pipe")

cat(x, 
    "^a^Mean of the annual point estimates and the lower and upper 95% credible intervals.", 
    "^b^Represents a binned value range from a choropleth map.",
    sep = "\n")

```

Cross-validation of the GAM loading models indicated that GAMs performed well on
average at predicting daily nutrient loading values. The variance in scores was
very high indicating subsets of values were problematic at characterizing functional
relationships between nutrients and predictors. Because all of the water quality
data for these two locations in the TCEQ databases were ambient water quality
data, collected to be representative of typical flow conditions, there were few
data at the highest portions of the flow-duration curve. It was beyond the scope
of the current study to evaluate the subsets of cross-validation data and scores.
However, the cross-validation procedure is indicative that more robust sampling
would be beneficial for reducing prediction variance. Supplementary flow-biased 
monitoring targeting storm- or high-flow conditions is recommended here to
improve the precision of GAM predictions [@horowitzEvaluationSedimentRating2003; @snelderEstimationCatchmentNutrient2017].

The non-linear temporal water quality trends identified using GAMs differed
slightly from previously identified trends [@bugica_water_2020]. This is
not unexpected due to the different time periods, different methodology, and
generally small slopes identified for most of the significant water quality
parameters in prior work. The trend in DO and 
cholorophyll-*a* concentrations are stable in comparison to other Texas
estuaries that are facing larger demands for freshwater diversions, higher
population growth, and more intense agricultural production 
[@wetzWaterQualityDynamics2016; @bugica_water_2020]. The trend of 
increasing NO~*x*~ concentration at the upper-bay TCEQ-13563 site and recent 
increases in TKN concentration at the mid-bay TCEQ-13383
site are concerning due to the nitrogen limitation identified in many Texas 
estuaries [@gardnerNitrogenFixationDissimilatory2006; 
@houTransformationFateNitrate2012; @doradoUnderstandingInteractionsFreshwater2015; 
@paudelRelationshipSuspendedSolids2019; @wetz_exceptionally_2017] 
and the relatively low ambient concentrations observed in Lavaca Bay.

The strong positive effect of freshwater inflow on NO~*x*~, TKN, and TP are
suggestive of nonpoint watershed sources, consistent with watershed uses and
with other studies relating freshwater inflow with nutrient 
concentrations in Lavaca Bay and other estuaries [@russell_effect_2006; 
@caffreyHighNutrientPulses2007; @peierlsNonmonotonicResponsesPhytoplankton2012; 
@palmerImpactsDroughtsLow2015; @ciraPhytoplanktonDynamicsLowInflow2021]. Inflow 
had a non-linear relationship with TKN at the two upstream sites, with TKN 
increasing as freshwater inflow transitioned from low to moderate levels. 
At higher freshwater inflows, the effect was attenuated, possibly indicating a 
flushing effect at higher freshwater inflow. No relationship between TKN and 
freshwater inflow were observed at TCEQ-13384 located in the lower reach of 
Lavaca Bay. Tidal flushing from Matagorda Bay could be responsible for 
diluting TKN and acting as a control on the effects of freshwater inflow in 
lower reaches of Lavaca Bay. Previous work suggests the processing of 
organic loads in the upper portions of Lavaca Bay mgiht reduce the transport of 
nutrients into the lower reaches of the Bay [@russell_effect_2006].

Freshwater inflow had a strong positive effect on chlorophyll-*a* at 
the upper- and mid-bay sites. The upper-bay site, TCEQ-13563, showed decreases
in chlorophyll-*a* at the highest freshwater inflow volumes. Freshwater flushing
or increases in turbidity are associated with decreases in chlorophyll-*a* in
other estuaries [@peierlsNonmonotonicResponsesPhytoplankton2012; 
@cloernPhytoplanktonPrimaryProduction2014]. No relationships between inorganic
nitrogen or TP loadings with chlorophyll-*a* were observed. Due to the lack
of TKN loading information, no assessment between organic nitrogen loads and
chlorophyll-*a* were possible.

Although other studies have identified complex relationships between estuary 
nutrient concentrations, nutrient loading and chlorophyll-*a* concentrations in 
Texas estuaries [@ornolfsdottirNutrientPulsingRegulator2004; 
@doradoUnderstandingInteractionsFreshwater2015; @ciraPhytoplanktonDynamicsLowInflow2021; @tominackVariabilityPhytoplanktonBiomass2022], 
this study specifically used flow-adjusted freshwater derived nutrient loads to 
parse out contributions from changes in nutrient loadings while accounting for 
variations in load due to flow. Loading GAMs indicated no evidence of changes in 
flow-normalized TP loads in either river, and no changes in flow-normalized NO~3~
loads in the Navidad River. The small changes in flow-normalized NO~3~ loads in
the Lavaca River are probably masked under most conditions by discharge from the
Navidad River. Given the relatively small variation in flow-normalized loads,
it can be expected that they would contribute little to the variance in downstream
water quality.

GAMs did not identify responses in DO concentration to inflows or nutrient loads.
The seasonality term in the temporal GAM models explained a substantial amount of DO
variation at all of the sites. Responses of estuary metabolic processes and 
resulting DO concentrations can be quite complicated and often locally specific [@caffreyFactorsControllingNet2004]. While the lack of total nitrogen or TKN 
loading data hinders interpretation, the large seasonal effect on DO suggests 
physical factors play an important role and should be included in future models.
Prior work suggests that Lavaca Bay may not be limited by nutrients alone, 
with high turbidity or nutrient processing in upper portions of the Bay or 
intertidal river limiting production [@russell_effect_2006]. Finally, it
is reasonable to assume that fluctuations in DO may not occur immediately in 
response to nutrient pulses or freshwater inflow. Work has has shown that various
water quality parameters may have lagged effects lasting days or even months following
storms and large discharge events [@mooney_watershed_2012; @wetzExtremeFutureEstuaries2013; @bukaveckas_influence_2020; @walkerTimescalesMagnitudeWater2021]. 
However, our work only evaluates responses to
loading and inflows occuring the day of water quality observations.

# Conclusion

GAM models appear to provide reliable estimates of nutrient loads in the Lavaca
Bay watershed. However, additional flow-biased data collection efforts would
decrease the prediction variance and improve accuracy at critical 
high flow events. Ongoing projects will fill data gaps for total nitrogen and 
TKN loading. This study, consistent with others along the Texas coast, found
strong effects of freshwater flow on nutrient and chlorophyll-*a* 
concentrations. DO concentrations, dominated by seasonal effects, did not show
strong direct responses to freshwater flow. Small variance in flow-adjusted
nutrient loads indicates that (1) there have been limited changes in non-point
sources of nutrients and (2) that there isn't strong evidence that those small 
changes have had effects on chlorophyll-*a* or dissolved oxygen in Lavaca Bay. 
Although the study did not identify strong responses to changes in nutrient 
loading, this does provide a baseline assessment for future water quality 
management activities in the watershed.

# Acknowledgements

The author extends thanks to Dr. Mike Wetz (Harte Research Institute, 
Texas A&M Corpus Christi), Chad Kinsfather and Patrick Brzozowski 
(Lavaca-Navidad River Authority), Brian Koch (Texas State Soil and Water 
Conservation Board), Bill Balboa (Matagorda Bay Foundation), Jason Pinchbeck 
(Texas General Land Office) and the Lavaca Bay Foundation for supporting 
development of this project and providing valuable project feedback.

# References