### 2023-05-23 Schramm-2023-05-AS/Schramm-2023-05-AS.Rmd
lots of edits

---
# Supported options: 
#   sn-nature:       Style for submissions to Nature Portfolio journals
#   sn-basic:        Basic Springer Nature Reference Style/Chemistry Reference Style
#   sn-mathphys:     Math and Physical Sciences Reference Style
#   sn-aps:          American Physical Society (APS) Reference Style
#   sn-vancouver:    Vancouver Reference Style
#   sn-apa:          APA Reference Style 
#   sn-chicago:      Chicago-based Humanities Reference Style
#   default:         Default

classoptions: 
  - sn-basic      
  # - Numbered      # Optional: Use numbered references instead of namedate references (only for sn-nature, sn-basic, sn-vancouver, sn-chicago, sn-mathphys or sn-nature)
  - referee       # Optional: Use double line spacing 
  - lineno        # Optional: Add line numbers
  # - iicol         # Optional: Double column layour

title: Assessing linkages between watershed nutrient loading and water quality in a subtropical estuary with semiparametric models.
titlerunning: Watershed nutrient loading and water quality in a subtropical estuary

authors: 
  - firstname: Michael
    lastname: Schramm
    email: michael.schramm@ag.tamu.edu
    affiliation: [1]
    corresponding: TRUE

affiliations:
  - number: 1
    corresponding: TRUE
    info:
      orgdiv: Texas Water Resources Institute
      orgname: Texas A&M AgriLife Research
    address:
        street: 1001 Holleman Dr. E.
        city: College Station
        postcode: 77840
        state: TX
        country: United States

keywords:
  - key
  - dictionary
  - word

# Sample for unstructured abstract 
abstract: |
  Lavaca Bay is a secondary embayment on the Texas coast displaying early signals of water quality degradation. This study applied a semiparametric approach to assess both watershed nutrient loads and responses in estuary water quality to nutrient loading and streamflow. Cross-validation indicated that, despite data constraints, semiparametric models performed well at nutrient load prediction. Based on these models, delivered annual nutrient loads varied substantially from year to year. In contrast, minimal changes in flow-normalized loads indicate that nutrient loadings were driven by natural variation in precipitation and runoff as opposed to changes in management of nonpoint sources. Models indicated there was no evidence of long-term changes in dissolved oxygen or chlorophyll-*a* within Lavaca Bay. However, site specific long-term increases in both organic and inorganic nitrogen are concerning. Further analysis found freshwater inflow was a strong driver of nutrient and chlorophyll-*a* concentrations within Lavaca Bay but there was no evidence that changes in watershed nutrient loading explained variation in dissolved oxygen or chlorophyll-*a* concentrations. In addition to providing a baseline assessment of watershed nutrient loading and water quality responses in the Lavaca Bay watershed, this study provides methodological support for the use of semiparametric methods in load regression models and estuary assessments.

bibliography: bibliography.bib
output: rticles::springer_article
---

```{r setup, include=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      out.extra = "")  # This forces knitr to label all figures.

library(targets)
library(knitr)
library(tidyverse)
library(kableExtra)
library(patchwork)
library(ggrepel)
library(rcartocolor)
library(twriTemplates)
library(ggspatial)
library(ragg)
library(modelsummary)
library(scico)
library(ggridges)
library(ggtext)
library(lubridate)
library(units)
library(gratia)
library(gt)

## this sets our default code chunk options
knitr::opts_chunk$set(echo = FALSE,
                      ## figure chunk options
                      dev = "ragg_png",
                      dpi = 300,
                      fig.align = "center",
                      out.width = "100%")

store <- "C:/Data-Analysis-Projects/lavaca-nutrients/_targets"
```

*Remaining Items*

- keywords
- discussion
- formalize language (reduce first person active?)
- final language review
- springer doesn't want figures in subdirectory, need to knit to same directory as tex file.

# Introduction {#sec1}

Similar to many coastal areas globally, the coastal watersheds along the Texas Gulf coast 
are facing pressures from increasing population, increases in point source 
and non-point source pollution and alterations to freshwater flows that degrade
water quality in downstream estuaries [@bricker_effects_2008; @kennicuttWaterQualityGulf2017; @bugica_water_2020]. 
Despite these escalating pressures, national scale assessments have classified 
coastal estuaries in Texas as moderate or low risk for eutrophic conditions 
[@bricker_effects_2008]. However, a suite of recent studies indicates that 
estuary water quality dynamics in both agriculturally dominated and urban 
watersheds within Texas are expressing conditions that are increasingly 
conducive to algal blooms and eutrophication [@wetzWaterQualityDynamics2016; 
@wetz_exceptionally_2017; @bugica_water_2020; @chinPhytoplanktonBiomassCommunity2022]. 
With identification of several localized areas of estuary water quality concern along 
the Texas coast [@bugica_water_2020], localized studies are being prioritized 
to better inform management actions.

This project aims to provide an assessment of watershed nutrient loading and resulting water quality
responses in Lavaca Bay, Texas. Lavaca Bay is a secondary bay in the larger
Matagorda Bay system located roughly halfway between Houston, Texas and Corpus
Christi, Texas. Lavaca Bay faces substantial challenges associated with
legacy contamination but general water quality parameters such as dissolved oxygen 
(DO), nutrients, and biological parameters have been well within state
water quality standards. Despite largely meeting statewide water quality standards,
there have been concerning recently identified declines in abundance, biomass, and diversity of 
benthic fauna in Lavaca Bay [@beserespollackLongtermTrendsResponse2011]. These declines
are partially attributed to reductions in freshwater inflow and changes in estuary salinity
and are indicative of an already stressed system 
[@beserespollackLongtermTrendsResponse2011; @palmerImpactsDroughtsLow2015;
@montagnaAssessmentRelationshipFreshwater2020]. More recently, significant
linear increases in total phosphorus (TP), orthophosphate, 
total Kjeldahl nitrogen (TKN), and chlorophyll-*a* concentrations were identified
at monitoring sites within Lavaca Bay [@bugica_water_2020]. Although 
long-term changes in DO concentrations have not been identified, the trends in 
nutrient concentrations are concerning due to the role of nitrogen as a 
limiting factor for primary production in many Texas estuaries 
[@gardnerNitrogenFixationDissimilatory2006; @houTransformationFateNitrate2012;
@doradoUnderstandingInteractionsFreshwater2015; 
@paudelRelationshipSuspendedSolids2019; @wetz_exceptionally_2017] and the 
ramifications that changes in nitrogen loadings could have for 
productivity and eutrophication in Lavaca Bay.

There are ongoing efforts between local, state, and federal agencies to 
address water quality impairments in the freshwater portions of the 
Lavaca Bay watershed [@jainTechnicalSupportDocument2021; 
@schrammLavacaRiverWatershed2018; @bertholdDirectMailingEducation2021]. 
However, at a statewide scale, these approaches have shown limited success and 
emphasize a need for improved efforts at assessing and linking management 
actions with downstream water quality to identify and replicate effective 
management actions across the state [@schrammTotalMaximumDaily2022]. The 
identification and communication of changes and trends in water quality is 
complicated by the fact that trends are often non-linear and confounded by 
precipitation and runoff that hinder traditional analysis 
[@wazniakLinkingWaterQuality2007; @lloydMethodsDetectingChange2014]. 
The development and application of flexible statistical methods such as Weighted
Regressions on Time, Discharge and Season [WRTDS, @hirschWeightedRegressionsTime2010] 
and Generalized Additive Models [GAMs, @woodFastStableRestricted2011] have 
provided effective tools for researchers to quantify and
communicate non-linear changes in river and estuary pollutant loadings.

WRTDS calculates a time series of in-stream concentrations or loads 
(daily, monthly, or annually) and flow-normalized estimates of concentrations and loads
using locally weighted regresssion for unique combinations of time, discharge, and season. 
WRTDS has been widely used to assess and identify trends in riverine nutrients 
[@oelsner_recent_2019; @stackpooleLongTermMississippi2021],
chlorides [@stetsIncreasingChlorideRivers2018], and other pollutants of concern 
[@shodaWaterqualityTrendsRivers2019]. WRTDS has also been succesfully adapted to 
assess trends in estuarine water quality concentrations [@beckFourDecadesWater2018].

While WRTDS is a statistical approach developed specifically for water quality applications, 
GAMs are a broadly applicable statistical method. GAMs are a semiparametric 
extension of generalized linear models where the linear predictor is represented
as the sum of multiple unknown smooth functions and parametric linear predictors
[@woodFastStableRestricted2011]. Although the underlying parameter estimation 
procedure of GAMs is substantially different than WRTDS, both the functional 
form and results have been demonstrated to be similar when assessing nutrient
concentration trends [@beckNumericalQualitativeContrasts2017]. 
Water quality applications of GAMs have included include river and catchment 
nutrient concentration and load estimation
[@wangLoadEstimationUncertainties2011;@kroonRiverLoadsSuspended2012;
@kuhnertQuantifyingTotalSuspended2012;@robson_prediction_2015-1;
@hagemannEstimatingNutrientOrganic2016;@mcdowell_implications_2021;
@biagi_novel_2022], assessment of temporal trends of nutrients
[@beckNumericalQualitativeContrasts2017; @murphyGeneralizedAdditiveModel2019],
phytoplankton [@bergbuschUnexpectedShiftPhytoplankton2021], 
and cyanobacteria [@hayesEffectsLakeWarming2020]. 
Recently GAMs have also been used to link water quality responses in 
receiving water bodies to changes in nonpoint source nutrient inputs 
[@murphyNutrientImprovementsChesapeake2022]. @beckNumericalQualitativeContrasts2017
provides a substantial discussion on the differences (and similarities) between 
GAMs and WRTDS for water quality applications.

To provide actionable information for resource managers in Lavaca Bay, 
water quality conditions must be evaluated relative to changes in natural 
environmental drivers to better understand and manage potential human impacts.
This study utilizes GAMs to develop estimates of delivered 
and flow-normalized nutrient loads and assess changes in loads delivered to 
Lavaca Bay. GAMs were chosen over other regression-based
approached for use in this study due to; (1) the ability to easily explore and 
incorporate different model terms; (2) the incorporation of non-liner smooth
functions that do not require explicit a priori knowledge of the expected
shape; and (3) inclusion of a link function that related the expected value of
the response to linear predictors thus avoiding unneeded data transformations
and bias corrections. The exploratory study also assesses the response of water quality parameters in
Lavaca Bay over time and in response to freshwater inflow controlled
for seasonality and to watershed nutrient loads that are controlled for
environmentally driven variation.

# Methods {#sec2}

## Location and Data
Lavaca Bay is 190 km^2^ with the majority of freshwater inflow provided by the 
Lavaca and Navidad River systems (Figure\ \ref{fig:fig1}). The Garcitas-Arenosa, 
Placedo Creek, and Cox Bay watersheds provide additional freshwater inflows. 
The entire watershed land area is 8,149 km^2^ and primarily rural. Watershed 
land cover and land use is 50% grazed pasture and rangeland, 20% cultivated 
cropland (primarily rows crops such as corn, cotton, and sorghum), and 5% 
suburban/urban. Pasture and rangeland is concentrated in the Lavaca River 
watershed, while cultivated crops are generally located along the eastern 
tributaries of the Navidad river. The Lavaca and Navidad River watersheds are a 
combined 5,966 km^2^, or approximately 73% of the entire Lavaca Bay watershed 
area. Discharge from the Navidad River is regulated by Lake Texana which has 
been in operation since 1980. Lake Texana provides 0.210 km^3^ of water 
storage and discharges into the tidal section of the Navidad River which 
ultimately joins the tidal section of the Lavaca River 15 km upstream of the 
confluence with the Lavaca Bay.

```{r fig1, fig.width=6.75, fig.width=5.2, out.width="5.2in", fig.cap="Map of Lavaca Bay and the contribution watershed. The freshwater sites are the most downstream freshwater stream locations with water quality and streamflow data used for nutrient load models. Water quality concentration data at the three Lavaca Bay sites were used to assess relationships between freshwater flows, loads and estuary water quality."}

knitr::include_graphics("../fig1.png")

```

Daily discharges for the Lavaca River (USGS-08164000, Figure\ \ref{fig:fig1}) 
were obtained from the United States Geologic Survey (USGS) National Water 
Information System using the *dataRetrieval* R package 
[@deciccoDataRetrievalPackagesDiscovering2022]. Gaged daily discharges from 
the outlet of Lake Texana on the Navidad River (USGS-0816425) were provided by 
the Texas Water Development Board (TWDB) 
(April 21, 2022 email from R. Neupane, TWDB).

Water quality sample data for the two freshwater and three estuary locations were 
obtained from the Texas Commission on Environmental Quality (TCEQ) Surface Water 
Quality Monitoring Information System. Data submitted through the system 
are required to be collected under Quality Assurance Project Plans and lab 
method procedures outlined by the TCEQ's procedures manual to ensure
consistent collection and laboratory methods are applied between samples 
collected by different entities and under different projects. 
All sites had varying lengths of and availability of data.
For freshwater locations, TP from January 2000 through December 2020 and 
nitrate-nitrogen (NO~3~) data from January 2005 through December 2020 were 
downloaded (Table\ \ref{tab:table1}). Less than 5-years of total
nitrogen and TKN concentration data were available at the freshwater sites and
deemed insufficient to develop load estimation models 
[@horowitzEvaluationSedimentRating2003; @snelderEstimationCatchmentNutrient2017]. 
The three estuary sites included an upper Lavaca Bay site near the outlet of the
Lavaca River system (TCEQ-13563), a mid-Lavaca Bay site (TCEQ-13383), and the 
lower Lavaca Bay site near the mouth of the Bay (TCEQ-13384). 
For estuary locations, we obtained data for TP, Nitrite+Nitrate (NO*~x~*), TKN, 
chlorophyll-*a*, and DO concentrations from January 2005 through December 2020 
(Table\ \ref{tab:table2}).

```{r table1}

df <- tar_read(model_data, store = store)

df <- df |>  
  select(Date, site_no, Flow, original_NO3, original_TP) |>
  mutate(site_no = case_when(
    site_no == "lktexana_g" ~ "usgs08164525",
    site_no != "lktexana_g" ~ site_no
  )) |>
  filter(site_no == "usgs08164000" |
           site_no == "usgs08164525") |> 
  rename(`NO\\textsubscript{3} (mg/L)` = original_NO3,
         `TP (mg/L)` = original_TP,
         `Station ID` = site_no,
         `Mean Daily Streamflow (cfs)` = Flow)
df$`Station ID` <- str_replace(df$`Station ID`, "usgs", "USGS-")

datasummary(`Station ID` * (`TP (mg/L)` + `NO\\textsubscript{3} (mg/L)` + `Mean Daily Streamflow (cfs)`) ~
               Mean*Arguments(na.rm=TRUE) + SD*Arguments(na.rm=TRUE) + N,
            data = df, output = "data.frame") |> 
  mutate(`Station ID` = c("USGS-08164000", "USGS-08164000", "USGS-08164000",
                          "USGS-08164525", "USGS-08164525", "USGS-08164525")) |> 
  rename("Parameter" = " ") |> 
  kbl(format = "latex", booktabs = TRUE, escape = FALSE,
      caption = "Summary of gauged streamflow and freshwater water quality samples between January 1, 2000 and December 31, 2020.") |>
  collapse_rows(columns = 1, latex_hline = "major")

```

```{r table2}
#| warning: false

df <- tar_read(estuary_model_data, store = store)

df <- df |> 
  filter(station_id == "13384" |
         station_id == "13383" |
         station_id == "13563") |>
  filter(parameter_code == "00630" |
           parameter_code == "00625" |
           parameter_code =="00665" |
           parameter_code == "70953" |
           parameter_code == "00300") |> 
  filter(end_date >= as.Date("2005-01-01")) |> 
  select(end_date, station_id, station_description, parameter_description, parameter_code, value) |> 
  mutate(parameter_description =
           case_when(
             parameter_code == "70953" ~ "Chlorophyll-\\textit{a} (ug/L)",
             parameter_code == "00625" ~ "TKN (mg/L)",
             parameter_code == "00630" ~ "NO\\textsubscript{x} (mg/L)",
             parameter_code == "00665" ~ "TP (mg/L)",
             parameter_code == "00300" ~ "DO (mg/L)"
           )) |>
  mutate(station_id = paste0("TCEQ-", station_id)) |> 
  rename(`Station ID` = station_id) |> 
  pivot_wider(names_from = parameter_description,
              values_from = value)

modelsummary::datasummary(`Station ID` * (`TP (mg/L)`+
                           `NO\\textsubscript{x} (mg/L)` +
                           `TKN (mg/L)` +
                           `Chlorophyll-\\textit{a} (ug/L)` +
                           `DO (mg/L)`) ~
                           Mean*Arguments(na.rm=TRUE) + SD*Arguments(na.rm=TRUE) + N,
                         data = df, output = "data.frame") |> 
  mutate(`Station ID` = c(rep("TCEQ-13383", 5), rep("TCEQ-13384", 5),
                          rep("TCEQ-13563", 5))) |> 
  rename("Parameter" = " ") |> 
  kbl(format = "latex", booktabs = TRUE, escape = FALSE,
      caption = "Summary of estuary water quality samples collected between January 1, 2005 and December 31, 2020.") |> 
  collapse_rows(columns = 1, latex_hline = "major")

# modelsummary::datasummary(`Station ID` * (`TP (mg/L)`+
#                            `NO\\textsubscript{x} (mg/L)` +
#                            `TKN (mg/L)` +
#                            `Chlorophyll-\\textit{a} (ug/L)` +
#                            `DO (mg/L)`) ~
#                            Mean*Arguments(na.rm=TRUE) + SD*Arguments(na.rm=TRUE) + N,
#                          data = df, output = "kableExtra", format = "latex",
#                          escape = FALSE,
#                          caption = "Summary of estuary water quality samples collected between January 1, 2005 and December 31, 2020.")
```

## Estimating Watershed Based Nutrient Loads

Estimates of NO~3~ and TP loads at the Lavaca River (USGS-08164000) and the 
outlet of Lake Texana on the Navidad River (USGS-08164525) were developed using 
GAMs relating nutrient concentration to river discharge, season, and time.
Separate models were fit at each station for each parameter and used to predict
nutrient concentrations for each day in the study period. GAMs were fit using 
the *mgcv* package in R which makes available multiple 
types of smooth functions with automatic smoothness selection [@woodFastStableRestricted2011]. 
The general form of the model related NO~3~ or TP concentration to a long term
tend, season, streamflow, and two different antecedent discharge terms, 
shown in Eq. \ref{eq1}:

\begin{align}
g(\mu) &= \alpha + f_1(ddate) + f_2(yday) + f_3(log1p(Q)) + f_4(ma) + f_5(fa), \nonumber \\
y &\sim \mathcal{N}(\mu,\,\sigma^{2})
\label{eq1}
\end{align}

where $\mu$ is the conditional expected NO~3~ or TP concentration, $g()$ is the 
log-link, $\alpha$ is the intercept, $f_n()$ are smoothing functions. $y$ is the 
response variable (NO~3~ or TP concentration) modeled as normally distributed 
with mean $\mu$ and standard deviation $\sigma$. *ddate* is the date converted to 
decimal notation, *yday* is numeric day of year (1-366), and *log1p(Q)* is the 
natural log of mean daily streamflow plus 1.

Moving average (*ma*) is an exponentially smoothed moving average that attempts 
to incorporate the influence of prior streamflow events on concentration at the 
current time period [@wangLoadEstimationUncertainties2011;  
@kuhnertQuantifyingTotalSuspended2012; @zhang_improving_2017], using Eq. \ref{eq2}:

\begin{align}
ma(\delta) &= d{\kappa_{i-1}}+(1-\delta)\hat{q}_{i-1}, \nonumber \\
\kappa_{i} &=\sum_{m=1}^{i}\hat{Q}_m
\label{eq2}
\end{align}

where $\delta$ is the discount factor (here, set equal to 0.95), $\kappa_i$
is the cumulative flow (*Q*) up to the *i*th day.

Flow anomaly (*fa*) is a unitless term that represents how wet or dry the 
current time period is from a previous time period [@vecchiaTrendsConcentrationsUse2009; 
@zhang_improving_2017]. Long-term flow anomaly (*ltfa*) is the streamflow over 
the previous year relative to the entire period [@zhang_improving_2017] and 
calculated using Eq. \ref{eq3}:

\begin{equation}
ltfa(t) = \bar{x}_{1\,year}(t) - \bar{x}_{entire\,period}
\label{eq3}
\end{equation}

and the short-term flow anomaly (*stfa*) calculated as the current day flow 
compared to the preceding 1-month streamflow using \ref{eq4}:

\begin{equation}
stfa(t) = x_{current\,day}(t) - \bar{x}_{1\,month}(t)
\label{eq4}
\end{equation}

where $x$ are the averages of log-transformed streamflow over the antecedent 
period (*1-year*, *1-month*, etc.) for time *t*.  We used *ltfa* in NO~3~ models 
and *stfa* in TP models based on previous work demonstrating major improvements 
in NO~x~ regression models that incorporated 
*ltfa* and moderate improvements in TP regression models that incorporated 
*stfa* [@zhang_improving_2017]. Moving averages and flow anomalies were calculated
with the *adc* R package [@schrammAdcCalculateAntecedant2023].

The calculation of model terms for the Lake Texana site were modified
because daily loads are not a function of natural stream flow processes alone,
but of dam releases and nutrient concentrations at the discharge point of the 
lake. *Q*, *ma*, and *fa* terms were calculated based on total gaged inflow from
the 4 major tributaries to the lake. Thin-plate regression splines were used 
for *ddate*, *log1p(Q)*, *fa*, and *ma*. A cyclic cubic regression spline was 
used for *yday* to ensure the ends of the spline match 
(day 1 and day 366 are expected to match). First order penalties were applied 
to the smooths of flow-based variables which penalize departures from a 
flat function to help constrain extrapolations for high flow measurements.

Left-censored data were not uncommon in this dataset. Several methods are
available to account for censored data. We transformed left-censored nutrient 
concentrations to one-half the detection limit. Although this simple approach 
can introduce bias [@hornungEstimationAverageConcentration1990],  we considered 
it acceptable because high concentrations and loadings are associated with 
high-flow events and low-flow/low-concentration events will account for a small 
proportion of total loadings [@mcdowell_implications_2021].

Daily loads were estimated as the predicted concentration multiplied by the 
daily streamflow. For the Navidad River (USGS-08164525) site, daily loads at the 
dam were calculated 
from the discrete daily concentration at the discharge point of the lake and 
corresponding reported daily discharge from the dam.
Flow-normalized loads were estimated similar to WRTDS by setting flow-based 
covariates on each day of the year equal to each of the historical values for
that day of the year over the study period [@hirschWeightedRegressionsTime2010]. The 
flow-normalized estimate was calculated as the mean of all the predictions for 
each day considering all possible flow values.
Standard deviations and 90% credible intervals were obtained by 
drawing samples from the multivariate normal posterior distribution of the 
fitted GAM [@woodConfidenceIntervalsGeneralized2006; 
@marraCoveragePropertiesConfidence2012; @mcdowell_implications_2021]. 
GAM performance was evaluated 
using repeated 5-fold cross validation  [@burmanComparativeStudyOrdinary1989] 
and average Nash-Sutcliffe Efficiency (NSE), Pearson sample correlation (*r*) 
and percent bias (PBIAS) metrics across folds were calculated for each model.

## Linking Estuary Water Quality to Hydrology and Nutrient Loads

To test if changes in freshwater inflow and nutrient loading had explanatory
effect on changes in estuary water quality a series of GAM models were fit at
each site relating parameter concentration to temporal trends (Eq.\ \ref{eq5}), 
temporal trends and inflow (Eq.\ \ref{eq6}), and temporal trends, inflow, and
nutrient loads (Eq.\ \ref{eq7}):
\begin{align}
g(\mu) &= \alpha + f_1(ddate) + f_2(yday), \label{eq5} \\
g(\mu) &= \alpha + f_1(ddate) + f_2(yday) + f_3(Q), \label{eq6} \\
g(\mu) &= \alpha + f_1(ddate) + f_2(yday) + f_3(Q) + f_4(Load) \label{eq7}
\end{align}

where $\mu$ is the conditional expected response (nutrient concentration), $g()$ 
is the log link, and response variable was modeled as Gamma distributed with 
mean $\mu$ and scale $\lambda$. $f_1(ddate)$ is decimal date smoothed with a 
thin-plate regression spline, $f_2(yday)$ is the numeric day of year smoothed 
with a cyclic cubic regression spline, $f_3(Q)$ is mean daily inflow (the 
combined measurements from Lavaca River and Navidad River) and $f_4(Load)$ is 
the total NO~3~ or TP watershed load. The set of models specified for each water
quality response are in Table\ \ref{tab:table3}.

```{r table3}
#| warning: false
df <- tibble(
  `Parameter` = c(rep("TP", 3), rep("NO\\textsubscript{x}", 3),
               rep("Chlorophyll-\\textit{a}", 3), rep("Dissolved Oxygen", 3),
               rep("TKN", 2)),
  Model = c("Temporal", "Flow", "Flow+Load",
            "Temporal", "Flow", "Flow+Load",
            "Temporal", "Flow", "Flow+Load",
            "Temporal", "Flow", "Flow+Load",
            "Temporal", "Flow"),
  `Model Terms` = c("s(ddate) + s(yday)",
                    "s(ddate) + s(yday) + s(Q)",
                    "s(ddate) + s(yday) + s(Q) + s(TP Load)",
                    "s(ddate) + s(yday)",
                    "s(ddate) + s(yday) + s(Q)",
                    "s(ddate) + s(yday) + s(Q) + s(NO\\textsubscript{3} Load)",
                    "s(ddate) + s(yday)",
                    "s(ddate) + s(yday) + s(Q)",
                    "s(ddate) + s(yday) + s(Q) + s(TP Load) + s(NO\\textsubscript{3} Load)",
                    "s(ddate) + s(yday)",
                    "s(ddate) + s(yday) + s(Q)",
                    "s(ddate) + s(yday) + s(Q) + s(TP Load) + s(NO\\textsubscript{3}  Load)",
                    "s(ddate) + s(yday)",
                    "s(ddate) + s(yday) + s(Q)")
)

kbl(df, format = "latex",
    caption = "Set of GAM models specified for each water quality parameter response.",
    booktabs = TRUE,
    escape = FALSE) |> 
  collapse_rows(columns = 1, latex_hline = "major")

# flextable::flextable(df) |> 
#   set_caption(caption = "Set of GAM models specified for each water quality parameter response.",
#               autonum = run_autonum(seq_id = "tab", bkm = "table3")) |> 
#   style(pr_t = fp_text_default(font.size = 9), part = "body") |>
#   style(pr_t = fp_text_default(font.size = 9, bold = TRUE), part = "header") |>
#   set_table_properties(layout = "autofit") |>
#   padding(padding.top = 0, padding.bottom = 0, part = "body")

```

Because streamflow and nutrient loads are tightly correlated, freshwater inflow
can mask signals from nutrient loads alone. Following the 
methodology implemented by @murphyNutrientImprovementsChesapeake2022,
both freshwater inflow and nutrient loads were prepossessed to account for season
and streamflow respectively. Raw freshwater inflow values were replaced by 
seasonally adjusted log transformed inflow obtained from the residuals of a GAM model 
fit between season(day of year) and log transformed daily freshwater inflow. Raw nutrient
loads were replaced with flow-adjusted values obtained from the residuals of a
GAM model relating log transformed NO~3~ or TP loads to log transformed daily
inflow. Response residuals from the respective GAM models were 
used as *Q* and *Load* in Eq.\ \ref{eq6} and Eq.\ \ref{eq7}.

This study used an information theoretic approach to evaluate 
evidence of model covariate effects on Lavaca Bay water quality.
Model probabilities were calculated 
and compared using the AIC~c~ scores between each group of temporal, inflow, and 
inflow+load models [@burnhamAICModelSelection2011]. Improvements in model 
probabilities provide evidence that the terms explain additional variation in 
the response variable.

# Results {#sec3}

## Watershed Nutrient Loads

Predictive performance of nutrient loads ranged from "satisfactory" to "very good"
based on standardized evaluation metrics of NSE, *r*, and PBIAS 
[@moriasiHydrologicWaterQuality2015] calculated using 5-fold cross validation.
Median goodness-of-fit metrics for NO~3~ models in the Lavaca 
River were 0.34 NSE, 0.70 *r*, and 2.00 PBIAS. Navidad River NO~3~ models 
appeared to perform slightly better with 0.48 NSE and 0.87 *r* but with higher 
bias at 10.90 PBIAS. Generally, TP models performed better than NO~3~ models. 
Median goodness-of-fit metrics for TP in the Lavaca River were 0.81 NSE, 
0.93 *r*, and -7.20 PBIAS. Navidad River TP models has similar performance 
with 0.91 NSE, 0.99 *r*, and -3.30 PBIAS. Density plots of
metrics show similar distribution of values between sites for the 
same parameter, with the exception *r* values for NO~3~ loads where 
Lavaca River had a much larger variance in values compared to the Navidad River
(Figure\ \ref{fig:fig2}). TP GAMS had higher average NSE and *r* values and 
lower variance in metric values compared to NO~3~.

```{r fig2, warning=FALSE, message=FALSE, fig.width=6.85, fig.asp=0.65, fig.cap="Density plots of goodness-of-fit metrics (NSE, \\textit{r}, and PBIAS) from repeated 5-fold cross validation between predicted nutrient loads from GAM models and measured nutrient loads. Color indicates the tail probability calculated from the empirical cumulative distribution of the goodness-of-fit metrics."}

## some functions to overide scales in facets via dewey dunnington
#https://dewey.dunnington.ca/post/2018/modifying-facet-scales-in-ggplot2/
scale_override <- function(which, scale) {
  if(!is.numeric(which) || (length(which) != 1) || (which %% 1 != 0)) {
    stop("which must be an integer of length 1")
  }

  if(is.null(scale$aesthetics) || !any(c("x", "y") %in% scale$aesthetics)) {
    stop("scale must be an x or y position scale")
  }

  structure(list(which = which, scale = scale), class = "scale_override")
}

CustomFacetWrap <- ggproto(
  "CustomFacetWrap", FacetWrap,
  init_scales = function(self, layout, x_scale = NULL, y_scale = NULL, params) {
    # make the initial x, y scales list
    scales <- ggproto_parent(FacetWrap, self)$init_scales(layout, x_scale, y_scale, params)

    if(is.null(params$scale_overrides)) return(scales)

    max_scale_x <- length(scales$x)
    max_scale_y <- length(scales$y)

    # ... do some modification of the scales$x and scales$y here based on params$scale_overrides
    for(scale_override in params$scale_overrides) {
      which <- scale_override$which
      scale <- scale_override$scale

      if("x" %in% scale$aesthetics) {
        if(!is.null(scales$x)) {
          if(which < 0 || which > max_scale_x) stop("Invalid index of x scale: ", which)
          scales$x[[which]] <- scale$clone()
        }
      } else if("y" %in% scale$aesthetics) {
        if(!is.null(scales$y)) {
          if(which < 0 || which > max_scale_y) stop("Invalid index of y scale: ", which)
          scales$y[[which]] <- scale$clone()
        }
      } else {
        stop("Invalid scale")
      }
    }

    # return scales
    scales
  }
)

facet_wrap_custom <- function(..., scale_overrides = NULL) {
  # take advantage of the sanitizing that happens in facet_wrap
  facet_super <- facet_wrap(...)

  # sanitize scale overrides
  if(inherits(scale_overrides, "scale_override")) {
    scale_overrides <- list(scale_overrides)
  } else if(!is.list(scale_overrides) || 
            !all(vapply(scale_overrides, inherits, "scale_override", FUN.VALUE = logical(1)))) {
    stop("scale_overrides must be a scale_override object or a list of scale_override objects")
  }

  facet_super$params$scale_overrides <- scale_overrides

  ggproto(NULL, CustomFacetWrap,
          shrink = facet_super$shrink,
          params = facet_super$params
  )
}

df <- tar_read(cv_no3_08164000, store = store) |> 
  mutate(site = "USGS-08164000",
         parameter = "NO <sub>3</sub>") |> 
  bind_rows(
    tar_read(cv_tp_08164000, store = store) |> 
      mutate(site = "USGS-08164000",
             parameter = "TP")
  ) |> 
  bind_rows(
    tar_read(cv_no3_texana, store = store) |> 
      mutate(site = "USGS-08164525",
             parameter = "NO <sub>3</sub>")
  ) |> 
  bind_rows(
    tar_read(cv_tp_texana, store = store) |> 
      mutate(site = "USGS-08164525",
             parameter = "TP")
  ) |> 
  mutate(site = case_when(
    site == "USGS-08164525" ~ "Navidad",
    site == "USGS-08164000" ~ "Lavaca"
  ))

df_long <- df |> 
  ungroup() |> 
  select(NSE, r2, pbias, site, parameter) |> 
  pivot_longer(cols = c(NSE, r2, pbias),
               names_to = "metric")

## plots the density estimates of the repeated 5-fold cross-validation goodness-of-fit metric results,
## color indicates the tail probability calculated from the empirical cumulitive distribution of the goodness-of-fit metric values 
p1 <- ggplot(df, aes(y = site, x = NSE,
               fill = 0.5 - abs(0.5 - after_stat(ecdf)))) +
  stat_density_ridges(geom = "density_ridges_gradient",
                      calc_ecdf = TRUE,
                      n = 10000) +
  scale_fill_scico("Tail Probability", palette = "hawaii", direction = -1,
                   limit = c(0, 0.5), breaks = c(0, 0.1, 0.2, 0.3, 0.4)) +
  guides(fill = guide_colorbar(barwidth = unit(100L, "pt"),
                               title.position = "top")) +
  facet_wrap_custom(~parameter, scales = "free_x",
                    ncol = 2,
                    scale_overrides = list(
                      scale_override(1, scale_x_continuous(limits = c(-1,1),
                                                           expand = c(0,0),
                                                           breaks = c(-1,0,1))),
                      scale_override(2, scale_x_continuous(limits = c(0,1),
                                                           expand = c(0,0),
                                                           breaks = c(0,0.5,1)))
                    )) +
  labs(x = "NSE", y = "") +
  theme_TWRI_print(base_family = "Arial") +
  theme(axis.text = element_text(size = 8),
        axis.title.x = element_markdown(size = 8),
        strip.text.x = element_markdown(size = 8),
        strip.background = element_rect(fill = "white"),
        panel.spacing.x = unit(15L, "pt"),
        legend.direction = "horizontal",
        legend.text = element_text(size = 8),
        legend.title = element_text(size = 8),
        plot.margin = unit(c(1L,0L,1L,0L), "pt"))

p2 <- ggplot(df, aes(y = site, x = r2,
               fill = 0.5 - abs(0.5 - after_stat(ecdf)))) +
  stat_density_ridges(geom = "density_ridges_gradient",
                      calc_ecdf = TRUE,
                      n = 10000) +
  scale_fill_scico("Tail Probability", palette = "hawaii", direction = -1,
                   limit = c(0, 0.5), breaks = c(0, 0.1, 0.2, 0.3, 0.4)) +
  guides(fill = guide_colorbar(barwidth = unit(100L, "pt"),
                               title.position = "top")) +
  facet_wrap_custom(~parameter, scales = "free_x",
                    ncol = 2,
                    scale_overrides = list(
                      scale_override(1, scale_x_continuous(limits = c(0,1),
                                                           expand = c(0,0),
                                                           breaks = c(0,0.5,1))),
                      scale_override(2, scale_x_continuous(limits = c(0,1),
                                                           expand = c(0,0),
                                                           breaks = c(0,0.5,1)))
                    )) +
  labs(x = "*r*", y = "") +
  theme_TWRI_print(base_family = "Arial") +
  theme(axis.text = element_text(size = 8),
        axis.title.x = element_markdown(size = 8),
        strip.text.x = element_markdown(size = 8),
        strip.background = element_rect(fill = "white"),
        panel.spacing.x = unit(15L, "pt"),
        legend.direction = "horizontal",
        legend.text = element_text(size = 8),
        legend.title = element_text(size = 8),
        plot.margin = unit(c(0L,0L,1L,0L), "pt"))

p3 <- ggplot(df, aes(y = site, x = pbias,
               fill = 0.5 - abs(0.5 - after_stat(ecdf)))) +
  stat_density_ridges(geom = "density_ridges_gradient",
                      calc_ecdf = TRUE,
                      n = 10000) +
  scale_fill_scico("Tail Probability", palette = "hawaii", direction = -1,
                   limit = c(0, 0.5), breaks = c(0, 0.1, 0.2, 0.3, 0.4)) +
  guides(fill = guide_colorbar(barwidth = unit(100L, "pt"),
                               title.position = "top")) +
  facet_wrap_custom(~parameter, scales = "free_x",
                    ncol = 2,
                    scale_overrides = list(
                      scale_override(1, scale_x_continuous(limits = c(-150,200),
                                                           expand = c(0,0))),
                      scale_override(2, scale_x_continuous(limits = c(-75,75),
                                                           expand = c(0,0)))
                    )) +
  labs(x = "PBIAS", y = "") +
  theme_TWRI_print(base_family = "Arial") +
  theme(axis.text = element_text(size = 8),
        axis.title.x = element_markdown(size = 8),
        strip.text.x = element_markdown(size = 8),
        strip.background = element_rect(fill = "white"),
        panel.spacing.x = unit(15L, "pt"),
        legend.direction = "horizontal",
        legend.text = element_text(size = 8),
        legend.title = element_text(size = 8),
        axis.text.y = element_blank(),
        plot.margin = unit(c(0L,0.25,1L,0L), "pt"))

layout <- "
AAAA####
AAAA#DD#
AAAA####
BBBBCCCC
BBBBCCCC
BBBBCCCC
"

p4 <- p1 / p2 / p3 / guide_area() + plot_layout(guides = 'collect',
                                          design = layout)

p4

```

```{r echo=FALSE}
## calculates min and max load values and years
load_lav_no3 <- tar_read(daily_no3_08164000, store = store)
load_nav_no3 <- tar_read(daily_no3_texana, store = store)

load_lav_no3$annually |> 
  mutate(site = "Lavaca River") -> lavaca_no3_annually

load_nav_no3$annually |> 
  mutate(site = "Navidad River") -> navidad_no3_annually

mean_no3 <- scales::label_comma()(mean(lavaca_no3_annually$NO3_Estimate + navidad_no3_annually$NO3_Estimate ))
mean_no3_low <- scales::label_comma()(mean(lavaca_no3_annually$NO3_Lower + navidad_no3_annually$NO3_Lower ))
mean_no3_upper <- scales::label_comma()(mean(lavaca_no3_annually$NO3_Upper + navidad_no3_annually$NO3_Upper ))

no3 <- data.frame(lavaca = lavaca_no3_annually$NO3_Estimate,
                  navidad = navidad_no3_annually$NO3_Estimate,
                  year = lavaca_no3_annually$year)
no3 <- no3 |> 
  mutate(total = lavaca+navidad)

min_no3 <- min(no3$total)
min_no3_year <- no3[no3$total == min_no3,]$year
min_no3 <- scales::label_comma()(min_no3)

max_no3 <- max(no3$total)
max_no3_year <- no3[no3$total == max_no3,]$year
max_no3 <- scales::label_comma()(max_no3)

no3_prop <- no3 |> 
  mutate(prop = round(navidad/total, 2))

no3_mean_prop <- scales::label_percent()(mean(no3_prop$prop))

min_no3_prop <- scales::label_percent()(min(no3_prop$prop))

load_lav_tp <- tar_read(daily_tp_08164000, store = store)
load_nav_tp <- tar_read(daily_tp_texana, store = store)

load_lav_tp$annually |> 
  mutate(site = "Lavaca River") |> 
  filter(year >= 2005)-> lavaca_tp_annually

load_nav_tp$annually |> 
  mutate(site = "Navidad River") |> 
  filter(year >= 2005) -> navidad_tp_annually

mean_tp <- scales::label_comma()(mean(lavaca_tp_annually$TP_Estimate + navidad_tp_annually$TP_Estimate ))
mean_tp_low <- scales::label_comma()(mean(lavaca_tp_annually$TP_Lower + navidad_tp_annually$TP_Lower ))
mean_tp_upper <- scales::label_comma()(mean(lavaca_tp_annually$TP_Upper + navidad_tp_annually$TP_Upper ))

tp <- data.frame(lavaca = lavaca_tp_annually$TP_Estimate,
                  navidad = navidad_tp_annually$TP_Estimate,
                  year = lavaca_tp_annually$year)
tp <- tp |> 
  mutate(total = lavaca+navidad)

min_tp <- min(tp$total)
min_tp_year <- tp[tp$total == min_tp,]$year
min_tp <- scales::label_comma()(min_tp)

max_tp <- max(tp$total)
max_tp_year <- tp[tp$total == max_tp,]$year
max_tp <- scales::label_comma()(max_tp)

tp_prop <- tp |> 
  mutate(prop = round(navidad/total, 2))

tp_mean_prop <- scales::label_percent()(mean(tp_prop$prop))

min_tp_prop <- scales::label_percent()(min(tp_prop$prop))
```

Annual NO~3~ and TP loads show considerable variation, generally following
patterns in discharge (Figures\ \ref{fig:fig3}, \ref{fig:fig4}). 
Flow-normalized TP loads at both sites and flow-normalized NO~3~ loads in the Lavaca River
indicated watershed based loads did not change much over time
when accounting for variation driven by streamflow (Figure\ \ref{fig:fig3}). 
Flow-normalized loads in the Lavaca River showed small variation over time with 
some decreases in NO~3~ loads since 2013.

```{r fig3, error=FALSE, warning=FALSE, message=FALSE, fig.width=6.85, fig.asp=0.7, fig.cap="Aggregated estimated annual and flow-normalized annual NO\\textsubscript{3} and TP loads for USGS-08164000 and USGS-08164525."}

load <- tar_read(daily_no3_08164000, store = store)
fn_load <- tar_read(daily_no3_08164000_fn, store = store)

load_lav_tp <- tar_read(daily_tp_08164000, store = store)
load_nav_tp <- tar_read(daily_tp_texana, store = store)

a <- ggplot() +
  geom_point(data = load$annually, aes(year, NO3_Estimate,
                                   color = "Total Annual Load + 90% CI",
                                   shape = "Total Annual Load + 90% CI")) +
  geom_line(data = load$annually, aes(x = year, y = NO3_Estimate,
                                  color = "Total Annual Load + 90% CI",
                                  linetype = "Total Annual Load + 90% CI"),
            alpha = 0.5) +
  geom_linerange(data = load$annually, aes(x = year, ymin = NO3_Lower, ymax = NO3_Upper,
                                       color = "Total Annual Load + 90% CI")) +
  geom_point(data = fn_load$annually, aes(year, NO3_Estimate,
                                 color = "Flow-Normalized Annual Load + 90% CI",
                                 shape = "Flow-Normalized Annual Load + 90% CI")) +
  geom_line(data = fn_load$annually,
            aes(x = year, y = NO3_Estimate,
                color = "Flow-Normalized Annual Load + 90% CI",
                linetype = "Flow-Normalized Annual Load + 90% CI"),
            alpha = 0.5) +
  geom_linerange(data = fn_load$annually,
                 aes(x = year, ymin = NO3_Lower, ymax = NO3_Upper,
                     color = "Flow-Normalized Annual Load + 90% CI")) +
  labs(x = "", 
       y = expression(Annual~NO[3]~Load~"[kg]"), 
       subtitle = "Lavaca River") +
  scale_shape_manual(name = "values",
                     values = c(21, 19)) +
  scale_color_manual(name = "values",
                     values = c("#7E1900", "#1A3399")) +
  scale_linetype_manual(name = "values",
                        values = c(1, 2)) +
  scale_y_continuous(trans = "pseudo_log",
                     breaks = c(0,1E1, 1E2, 1E3, 1E4, 1E5, 1E6, 1E7),
                     labels = scales::label_log()) +
  coord_cartesian(ylim = c(1000, 1E7)) +
  theme_TWRI_print(base_family = "sans") +
  theme(axis.title.y = element_text(size = 8),
        axis.text.x = element_blank(),
        plot.subtitle = element_text(size = 8, hjust = 0.5, face = "bold"),
        panel.grid.major.x = element_line(color = "#d9d9d9",
                                          linetype = "dotted"),
        legend.title = element_blank(),
        legend.text = element_text(size = 8),
        plot.margin = unit(c(0L,3L,0L,0L), "pt"))

load <- tar_read(daily_no3_texana, store = store)
fn_load <- tar_read(daily_no3_texana_fn, store = store)

b <- ggplot() +
  geom_point(data = load$annually, aes(year, NO3_Estimate,
                                   color = "Total Annual Load + 90% CI",
                                   shape = "Total Annual Load + 90% CI")) +
  geom_line(data = load$annually, aes(x = year, y = NO3_Estimate,
                                  color = "Total Annual Load + 90% CI",
                                  linetype = "Total Annual Load + 90% CI"),
            alpha = 0.5) +
  geom_linerange(data = load$annually, aes(x = year, ymin = NO3_Lower, ymax = NO3_Upper,
                                       color = "Total Annual Load + 90% CI")) +
  geom_point(data = fn_load$annually, aes(year, NO3_Estimate,
                                 color = "Flow-Normalized Annual Load + 90% CI",
                                 shape = "Flow-Normalized Annual Load + 90% CI")) +
  geom_line(data = fn_load$annually,
            aes(x = year, y = NO3_Estimate,
                color = "Flow-Normalized Annual Load + 90% CI",
                linetype = "Flow-Normalized Annual Load + 90% CI"),
            alpha = 0.5) +
  geom_linerange(data = fn_load$annually,
                 aes(x = year, ymin = NO3_Lower, ymax = NO3_Upper,
                     color = "Flow-Normalized Annual Load + 90% CI")) +
  labs(x = "", 
       y = expression(Annual~NO[3]~Load~"[kg]"), 
       subtitle = "Navidad River") +
  scale_shape_manual(name = "values",
                     values = c(21, 19)) +
  scale_color_manual(name = "values",
                     values = c("#7E1900", "#1A3399")) +
  scale_linetype_manual(name = "values",
                        values = c(1, 2)) +
  scale_y_continuous(trans = "pseudo_log",
                     breaks = c(0,1E1, 1E2, 1E3, 1E4, 1E5, 1E6, 1E7),
                     labels = scales::label_log()) +
  coord_cartesian(ylim = c(1000, 1E7)) +
  theme_TWRI_print(base_family = "sans") +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.text.x = element_blank(),
        plot.subtitle = element_text(size = 8, hjust = 0.5, face = "bold"),
        panel.grid.major.x = element_line(color = "#d9d9d9",
                                          linetype = "dotted"),
        legend.title = element_blank(),
        legend.text = element_text(size = 8),
        plot.margin = unit(c(0L,0L,0L,3L), "pt"))

load <- tar_read(daily_tp_08164000, store = store)
fn_load <- tar_read(daily_tp_08164000_fn, store = store)

load$annually <- load$annually |> 
  filter(year >= 2005)

fn_load$annually <- fn_load$annually |> 
  filter(year >= 2005)

c <- ggplot() +
  geom_point(data = load$annually, aes(year, TP_Estimate,
                                   color = "Total Annual Load + 90% CI",
                                   shape = "Total Annual Load + 90% CI")) +
  geom_line(data = load$annually, aes(x = year, y = TP_Estimate,
                                  color = "Total Annual Load + 90% CI",
                                  linetype = "Total Annual Load + 90% CI"),
            alpha = 0.5) +
  geom_linerange(data = load$annually, aes(x = year, ymin = TP_Lower, ymax = TP_Upper,
                                       color = "Total Annual Load + 90% CI")) +
  geom_point(data = fn_load$annually, aes(year, TP_Estimate,
                                 color = "Flow-Normalized Annual Load + 90% CI",
                                 shape = "Flow-Normalized Annual Load + 90% CI")) +
  geom_line(data = fn_load$annually,
            aes(x = year, y = TP_Estimate,
                color = "Flow-Normalized Annual Load + 90% CI",
                linetype = "Flow-Normalized Annual Load + 90% CI"),
            alpha = 0.5) +
  geom_linerange(data = fn_load$annually,
                 aes(x = year, ymin = TP_Lower, ymax = TP_Upper,
                     color = "Flow-Normalized Annual Load + 90% CI")) +
  labs(x = "", y = "Annual TP Load [kg]", subtitle = "Lavaca River") +
  scale_shape_manual(name = "values",
                     values = c(21, 19)) +
  scale_color_manual(name = "values",
                     values = c("#7E1900", "#1A3399")) +
  scale_linetype_manual(name = "values",
                        values = c(1, 2)) +
  scale_y_continuous(trans = "pseudo_log",
                     breaks = c(0,1E1, 1E2, 1E3, 1E4, 1E5, 1E6, 1E7),
                     labels = scales::label_log()) +
  coord_cartesian(ylim = c(1000, 1E6)) +
  theme_TWRI_print(base_family = "sans") +
  theme(axis.title.y = element_text(size = 8),
        panel.grid.major.x = element_line(color = "#d9d9d9",
                                          linetype = "dotted"),
        legend.title = element_blank(),
        legend.text = element_text(size = 8),
        plot.margin = unit(c(0L,3L,0L,0L), "pt"),
        plot.subtitle = element_blank())

load <- tar_read(daily_tp_texana, store = store) 
fn_load <- tar_read(daily_tp_texana_fn, store = store)

load$annually <- load$annually |> 
  filter(year >= 2005)

fn_load$annually <- fn_load$annually |> 
  filter(year >= 2005)

d <- ggplot() +
  geom_point(data = load$annually, aes(year, TP_Estimate,
                                   color = "Total Annual Load + 90% CI",
                                   shape = "Total Annual Load + 90% CI")) +
  geom_line(data = load$annually, aes(x = year, y = TP_Estimate,
                                  color = "Total Annual Load + 90% CI",
                                  linetype = "Total Annual Load + 90% CI"),
            alpha = 0.5) +
  geom_linerange(data = load$annually, aes(x = year, ymin = TP_Lower, ymax = TP_Upper,
                                       color = "Total Annual Load + 90% CI")) +
  geom_point(data = fn_load$annually, aes(year, TP_Estimate,
                                 color = "Flow-Normalized Annual Load + 90% CI",
                                 shape = "Flow-Normalized Annual Load + 90% CI")) +
  geom_line(data = fn_load$annually,
            aes(x = year, y = TP_Estimate,
                color = "Flow-Normalized Annual Load + 90% CI",
                linetype = "Flow-Normalized Annual Load + 90% CI"),
            alpha = 0.5) +
  geom_linerange(data = fn_load$annually,
                 aes(x = year, ymin = TP_Lower, ymax = TP_Upper,
                     color = "Flow-Normalized Annual Load + 90% CI")) +
  labs(x = "", y = "Annual TP Load [kg]", subtitle = "Navidad River") +
  scale_shape_manual(name = "values",
                     values = c(21, 19)) +
  scale_color_manual(name = "values",
                     values = c("#7E1900", "#1A3399")) +
  scale_linetype_manual(name = "values",
                        values = c(1, 2)) +
  scale_y_continuous(trans = "pseudo_log",
                     breaks = c(0,1E1, 1E2, 1E3, 1E4, 1E5, 1E6, 1E7),
                     labels = scales::label_log()) +
  coord_cartesian(ylim = c(1000, 1E6)) +
  theme_TWRI_print(base_family = "sans") +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        panel.grid.major.x = element_line(color = "#d9d9d9",
                                          linetype = "dotted"),
        legend.title = element_blank(),
        legend.text = element_text(size = 8),
        plot.margin = unit(c(0L,1L,0L,3L), "pt"),
        plot.subtitle = element_blank())

layout <- "
11112222
11112222
11112222
11112222
33334444
33334444
33334444
33334444
55555555
"

a + b + c + d + guide_area() + plot_layout(guides = 'collect', design = layout)
```

Aggregated across both sites, the mean 
annual NO~3~ load 2005 through 2020 was `r mean_no3` kg (`r mean_no3_low` kg - 
`r mean_no3_upper` kg, 90% CI). Annual NO~3~ loads ranged from `r min_no3` kg 
in `r min_no3_year` to `r max_no3` kg in `r max_no3_year`. Total annual TP loads 
ranged from `r min_tp` kg in `r min_tp_year` to 
`r max_tp` kg  in `r max_tp_year`. Mean annual TP loading from 2005 through 2020
was `r mean_tp` kg (`r mean_tp_low` kg - `r mean_tp_upper` kg, 90% CI).
On average, the Navidad River accounted for `r no3_mean_prop` of NO~3~ loads and 
`r tp_mean_prop` of TP loads from 2005 through 2020. However, during periods of extreme 
drought the Lavaca River became the primary source of 
nutrient loading in the watershed with the Navidad River only accounting for 
`r min_no3_prop` and `r min_tp_prop` of NO~3~ and TP loads in 2011 
(Figure\ \ref{fig:fig4}).

```{r fig4, error=FALSE, message=FALSE, warning=FALSE, fig.width=3.3, fig.asp=2, fig.cap="Comparison of delivered annual loads and annual discharge at the Lavaca (USGS-08164000) and Navidad (USGS-08164525) Rivers.", out.width="0.5\\linewidth"}

load_lav_no3 <- tar_read(daily_no3_08164000, store = store)
load_nav_no3 <- tar_read(daily_no3_texana, store = store)

load_lav_no3$annually |> 
  mutate(site = "Lavaca River") -> lavaca_no3_annually

load_nav_no3$annually |> 
  mutate(site = "Navidad River") -> navidad_no3_annually

no3_annual <- bind_rows(lavaca_no3_annually, navidad_no3_annually) |> 
  mutate(x = as.Date(paste0(year, "-01-01"), "%Y-%m-%d"))

# b <- ggplot(no3_annual) +
#   geom_col(aes(year, NO3_Estimate, fill = site), width = 0.8) +
#   scale_x_continuous(expand = expansion(mult = c(0.05, 0.05)), 
#                      breaks = c(2005, 2010, 2015, 2020)) +
#   scale_y_continuous(labels = scales::comma) +
#   scale_fill_carto_d(palette = "Vivid", direction = -1) +
#   labs(x = "", 
#        y = "Predicted Annual<br>NO <sub>3</sub> Load [kg]") +
#   theme_TWRI_print(base_family = "Arial") +
#   theme(axis.title.y = element_text(size = 8),
#         axis.text = element_text(size = 8),
#         panel.grid.major.x = element_line(color = "#d9d9d9",
#                                           linetype = "dotted"),
#         legend.title = element_blank(),
#         legend.text = element_text(size = 8))

prop <- no3_annual |> 
  mutate(proportion = NO3_Estimate/sum(NO3_Estimate))

ylab1 <- expression(Annual~NO[3]~Load)
ylab2 <- expression("[Proportion]")
c <- ggplot() +
  geom_col(data = prop, aes(year, proportion, fill  = site)) +
  scale_fill_carto_d(palette = "Vivid", direction = -1) +
  scale_x_continuous(breaks = c(2005, 2010, 2015, 2020),
                     expand = expansion(mult = c(0.05, 0.05))) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = "",
       y = "")  +
  coord_cartesian(clip = "off") +
  annotation_custom(grid::textGrob(ylab1,
                                   x = unit(-0.23, "npc"),
                                   rot = 90,
                                   gp = grid::gpar(fontfamily = "sans",
                                             fontsize = 8))) +
  annotation_custom(grid::textGrob(ylab2,
                                   x = unit(-0.18, "npc"),
                                   rot = 90,
                                   gp = grid::gpar(fontfamily = "sans",
                                                   fontsize = 8))) +
  theme_TWRI_print(base_family = "sans") +
  theme(axis.title.y = element_text(size = 8),
        axis.text = element_text(size = 8),
        panel.grid.major.x = element_line(color = "#d9d9d9",
                                          linetype = "dotted"),
        legend.title = element_blank(),
        legend.text = element_text(size = 8))

ylab1 <- expression("Annual Discharge")
ylab2 <- expression("[Million Gallons]")
d <- tar_read(qdata, store = store) |>
  filter(site_no %in% c("8164000", "lktexana_g")) |>
  filter(Date >= as.Date("2005-01-01")) |>
  mutate(site = case_when(
    site_no == "8164000" ~ "Lavaca River",
    site_no == "lktexana_g" ~ "Navidad River"
  )) |>
  mutate(year = year(Date)) |>

  mutate(Flow = as_units(Flow, "ft^3/s")) |>
  mutate(Flow = set_units(Flow, "ft^3/day")) |>
  mutate(Flow = set_units(Flow, "1E6gallons/day")) |>
  group_by(year, site) |>
  summarise(Flow = sum(Flow, na.rm = TRUE)) |>
  mutate(Flow = drop_units(Flow)) |>
  ggplot() +
  geom_col(aes(year, Flow, fill = site)) +
  scale_fill_carto_d(palette = "Vivid", direction = -1) +
  scale_x_continuous(breaks = c(2005, 2010, 2015, 2020),
                     expand = expansion(mult = c(0.05, 0.05))) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = "",
       y = "") +
  coord_cartesian(clip = "off") +
  annotation_custom(grid::textGrob(ylab1,
                                   x = unit(-0.25, "npc"),
                                   rot = 90,
                                   gp = grid::gpar(fontfamily = "sans",
                                             fontsize = 8))) +
  annotation_custom(grid::textGrob(ylab2,
                                   x = unit(-0.20, "npc"),
                                   rot = 90,
                                   gp = grid::gpar(fontfamily = "sans",
                                                   fontsize = 8))) +
  theme_TWRI_print(base_family = "Arial") +
  theme(axis.title.y = element_text(size = 8),
        axis.text = element_text(size = 8),
        panel.grid.major.x = element_line(color = "#d9d9d9",
                                          linetype = "dotted"),
        legend.title = element_blank(),
        legend.text = element_text(size = 8),
        legend.position = "none",
        plot.margin = unit(c(1L,1L,1L,2L), "pt"))

load_lav_tp$annually |>
  mutate(site = "Lavaca River") -> lavaca_tp_annually

load_nav_tp$annually |>
  mutate(site = "Navidad River") -> navidad_tp_annually

tp_annual <- bind_rows(lavaca_tp_annually, navidad_tp_annually) |>
  filter(year >= 2005) |>
  mutate(x = as.Date(paste0(year, "-01-01"), "%Y-%m-%d"))
# 
# e <- ggplot(tp_annual) +
#   geom_col(aes(year, TP_Estimate, fill = site), width = 0.8) +
#   scale_x_continuous(expand = expansion(mult = c(0.05, 0.05)), 
#                      breaks = c(2005, 2010, 2015, 2020)) +
#   scale_y_continuous(labels = scales::comma) +
#   scale_fill_carto_d(palette = "Vivid", direction = -1) +
#   labs(x = "", y = "Predicted Annual<br>TP Load [kg]") +
#   theme_TWRI_print(base_family = "Arial") +
#   theme(axis.title.y = element_text(size = 8),
#         axis.text = element_text(size = 8),
#         panel.grid.major.x = element_line(color = "#d9d9d9",
#                                           linetype = "dotted"),
#         legend.title = element_blank(),
#         legend.text = element_text(size = 8))
# 
# 
prop <- tp_annual |>
  mutate(proportion = TP_Estimate/sum(TP_Estimate))

ylab1 <- expression("Annual TP Load")
ylab2 <- expression("[Proportion]")
f <- ggplot() +
  geom_col(data = prop, aes(year, proportion, fill  = site)) +
  scale_fill_carto_d(palette = "Vivid", direction = -1) +
  scale_x_continuous(breaks = c(2005, 2010, 2015, 2020),
                     expand = expansion(mult = c(0.05, 0.05))) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = "",
       y = "") +
  coord_cartesian(clip = "off") +
  annotation_custom(grid::textGrob(ylab1,
                                   x = unit(-0.23, "npc"),
                                   rot = 90,
                                   gp = grid::gpar(fontfamily = "sans",
                                             fontsize = 8))) +
  annotation_custom(grid::textGrob(ylab2,
                                   x = unit(-0.18, "npc"),
                                   rot = 90,
                                   gp = grid::gpar(fontfamily = "sans",
                                                   fontsize = 8))) +
  theme_TWRI_print(base_family = "Arial") +
  theme(axis.title.y = element_text(size = 8),
        axis.text = element_text(size = 8),
        panel.grid.major.x = element_line(color = "#d9d9d9",
                                          linetype = "dotted"),
        legend.title = element_blank(),
        legend.text = element_text(size = 8),
        legend.position = "none")

layout <- "
CC
CC
CC
CC
FF
FF
FF
FF
DD
DD
DD
DD
GG
"

c + d + f + guide_area() +
  plot_layout(guides = 'collect', design  = layout)

```

## Linkages Between Water Quality and Watershed Flows and Loads

There is no evidence of long-term changes in TP or DO concentrations at any
Lavaca Bay site (Figure\ \ref{fig:fig5}). The upper-bay site, TCEQ-13563, shows
evidence of a long-term linear increase in NO~x~ while chlorophyll-*a*
decreased from 2005 through 2014 (Figure\ \ref{fig:fig5}). NO~x~ concentration
at the mid-bay site, TCEQ-13383, displayed an unusual periodic pattern that is
indicative of a strong influence from inflow or precipitation. The temporal GAMs
did not provide evidence of long-term trends in any of the water quality constituents
at the lower-bay TCEQ-13384 site.

```{r fig5, message=FALSE, warning=FALSE, fig.width=6.85, fig.asp=0.7, fig.cap="Fitted splines (shaded regions indicate 90\\% confidence intervals) from the temporal estuary GAM display the marginal smoothed effect of date on TP (A), NO\\textit{\\textsubscript{x}} (B), chlorophyll-\\textit{a} (C), TKN (D), and DO (E) concentrations at each site in Lavaca Bay."}

draw_smoothed_years <- function(model1, model2, model3,
                                ylab, subtitle) {

  `TCEQ-13563` <- model1
  `TCEQ-13383` <- model2
  `TCEQ-13384` <- model3

  newdata <- data_slice(model1, ddate = evenly(2005:2020, n = 200))

  comp <- gratia::compare_smooths(
    `TCEQ-13563`,
    `TCEQ-13383`,
    `TCEQ-13384`,
    smooths = "s(ddate)",
    data = newdata,
    overall_uncertainty = FALSE)

  crit <- gratia:::coverage_normal(0.90)

  comp |> 
    unnest(data) |> 
    mutate(lower_ci = est + (crit * se),
           upper_ci = est - (crit * se),
           label = if_else(ddate == max(ddate),
                           as.character(model),
                           NA_character_)) |> 
    ggplot(aes(x = ddate, y = est, group = model)) +
    geom_line(aes(color = model, linetype = model), alpha = 0.75) +
    geom_ribbon(aes(ymin = lower_ci,
                    ymax = upper_ci,
                    fill = model),
                alpha = 0.25) +
    scale_x_continuous(breaks = c(2005, 2010, 2015, 2020),
                       expand = expansion(mult = c(0.05,0.05))) +
    scale_fill_manual(name = "", values = unname(palette.colors(palette = "Okabe-Ito")[2:4])) +
    scale_color_manual(name = "", values = unname(palette.colors(palette = "Okabe-Ito")[2:4])) +
    scale_linetype_manual(name = "", values = c("solid","twodash","dashed")) +
    labs(x = "Year", y = ylab,
         subtitle = subtitle) +
    theme_TWRI_print(base_family = "Arial") +
    theme(axis.title.y = element_markdown(size = 8),
          axis.title.x = element_text(size = 8),
          axis.text.x = element_text(size = 8),
          axis.text.y = element_text(size = 8),
          panel.grid = element_blank(),
          plot.subtitle = element_text(size = 8),
          legend.direction = "vertical",
          legend.text = element_text(size = 8),
          legend.key.width = unit(25, "points"))

}

p_tp <- draw_smoothed_years(tar_read(tp_lavaca_13563_temporal, store = store),
                            tar_read(tp_lavaca_13383_temporal, store = store),
                            tar_read(tp_lavaca_13384_temporal, store = store),
                            ylab = "TP [mg/L]", 
                            subtitle = "A")

p_no3 <- draw_smoothed_years(tar_read(no3_lavaca_13563_temporal, store = store),
                             tar_read(no3_lavaca_13383_temporal, store = store),
                             tar_read(no3_lavaca_13384_temporal, store = store),
                             ylab = "NO<sub>*x*</sub> [mg/L]", 
                             subtitle = "B")

p_tkn <- draw_smoothed_years(tar_read(tkn_lavaca_13563_temporal, store = store),
                             tar_read(tkn_lavaca_13383_temporal, store = store),
                             tar_read(tkn_lavaca_13384_temporal, store = store),
                             ylab = "TKN [mg/L]", 
                             subtitle = "D")

p_chla <- draw_smoothed_years(tar_read(chla_lavaca_13563_temporal, store = store),
                             tar_read(chla_lavaca_13383_temporal, store = store),
                             tar_read(chla_lavaca_13384_temporal, store = store),
                             ylab = "Chlorophyll-*a* [μg/L]", 
                             subtitle = "C")

p_do <- draw_smoothed_years(tar_read(do_lavaca_13563_temporal, store = store),
                             tar_read(do_lavaca_13383_temporal, store = store),
                             tar_read(do_lavaca_13384_temporal, store = store),
                             ylab = "DO [mg/L]", 
                            subtitle = "E")

design <- "
112266
334455
"

#(p_tp + p_no3) / (p_chla + p_tkn) / (p_do)
p_tp + p_no3 + p_chla + p_tkn + p_do + guide_area() + plot_layout(guides = "collect",
                                                   design = design)
```

Freshwater inflow provided additional explanation for changes in TP and 
NO~*x*~ concentration at all of the Lavaca Bay sites according to AIC~c~ and 
model probability values (Table\ \ref{tab:table4}).
TCEQ-13563, the site closest to the river outlet, was the only site that
had improvements in the explanations of DO and TKN concentration with the inclusion 
of inflow. Both TCEQ-13563 and TCEQ-13383, the mid-bay site, saw improvements 
in explanations for variations in chlorophyll-*a* with the inclusion of freshwater
inflow. The addition of nutrient loads (both TP and NO~3~) terms did not provide
additional explanation for changes in chlorophyll-*a* or DO concentrations.
Inclusion of TP loads provided additional explanation of TP concentrations at
the upper- and mid-bay sites, TCEQ-13563 and TCEQ-13383. Inclusion of NO~3~
loads only provided marginal improvements in the explanation of NO~*X*~ 
concentration at the lower-bay TCEQ-13384 site.

```{r table4, message=FALSE, warning=FALSE}

dplyr::bind_rows(
    tar_read(tp_13563, store = store),
    tar_read(tp_13383, store = store),
    tar_read(tp_13384, store = store),
    tar_read(no3_13563, store = store),
    tar_read(no3_13383, store = store),
    tar_read(no3_13384, store = store),
    tar_read(chla_13563, store = store),
    tar_read(chla_13383, store = store),
    tar_read(chla_13384, store = store),
    tar_read(do_13563, store = store),
    tar_read(do_13383, store = store),
    tar_read(do_13384, store = store),
    tar_read(tkn_13563, store = store),
    tar_read(tkn_13383, store = store),
    tar_read(tkn_13384, store = store)
    ) |>
  mutate(Parameter = case_when(
    Parameter == "Nitrite+Nitrate" ~ "NO\\textsubscript{x}",
    Parameter == "Chlorophyll-a" ~ "Chlorophyll-\\textit{a}",
    TRUE ~ as.character(Parameter)),
    Model = case_when(
      Model == "Temporal" ~ "Temporal",
      Model == "Flow" ~ "Inflow",
      Model == "Flow + Load" ~ "Inflow + Load",
      TRUE ~ as.character(Model)
    )) |>
  mutate(Parameter = fct_relevel(Parameter, "TP", "NO\\textsubscript{x}", "Chlorophyll-\\textit{a}","TKN", "DO"),
         Model = fct_relevel(Model, "Temporal", "Inflow", "Inflow + Load")) |>
  tidyr::complete(Site, Model, nesting(Parameter)) |>
  mutate(model_prob = paste0(round(AICc,1), " (", round(weight,2), ")")) |>
  mutate(model_prob = case_when(
    model_prob == "NA (NA)" ~ "-",
    model_prob != "NA (NA)" ~ model_prob
  )) |>
  select(-c(AICc, weight)) |>
  pivot_wider(names_from = Model, values_from = model_prob) |>
  select(Parameter, Site, Temporal, Inflow, `Inflow + Load`) |>
  arrange(Parameter, Site) |>
  kbl(format = "latex",
      caption = "Estuary GAM AIC\\textsubscript{c} values and associated model probabilities (in parenthesis). Models with the highest probability for each site and water quality parameter combination are bolded and italicized for emphasis.",
      booktabs = TRUE, escape = FALSE) |> 
  column_spec(column = 3, 
              bold = c(F,F,F,F,F,F,F,T,F,T,T,F,T,T,F),
              italic = c(F,F,F,F,F,F,F,T,F,T,T,F,T,T,F)) |> 
  column_spec(column = 4, 
              bold = c(F,T,F,T,F,T,T,F,T,F,F,T,F,F,T),
              italic = c(F,T,F,T,F,T,T,F,T,F,F,T,F,F,T)) |> 
  column_spec(column = 5,
              bold = c(T,F,T,F,T,F,F,F,F,F,F,F,F,F,F),
              italic = c(T,F,T,F,T,F,F,F,F,F,F,F,F,F,F)) |> 
  collapse_rows(columns = 1, latex_hline = "major")

```

GAMs showed that increases in freshwater inflow resulted in nearly linear increases 
in TP and NO~*x*~ concentration at all three sites (Figure\ \ref{fig:fig6}). 
At the upper-bay TCEQ-13563 site, GAMs showed that increases in freshwater inflow initially 
increased chlorophyll-*a* and DO concentration but concentrations leveled and
potentially decreased at higher flows. The mid-bay TCEQ-13383 site showed a
nearly linear increased in chlorophyll-*a* concentration in response to increases
freshwater inflow. Freshwater flow did not have significant
effects on chlorophyll-*a*, TKN, or DO at the lower-bay TCEQ-13384 site.

```{r fig6, message=FALSE, warning=FALSE, fig.width=6.85, fig.asp=0.7, fig.cap="Fitted splines from estuary GAMs display the marginal smoothed effect of freshwater inflow (controlled for season) on TP (A), NO\\textit{\\textsubscript{x}} (B), chlorophyll-\\textit{a} (C), TKN (D), and DO (E) concentrations at each site in Lavaca Bay."}

draw_smoothed_surface <- function(comp,
                                  x = flw_res,
                                  xlab,
                                  ylab, 
                                  subtitle) {

  crit <- gratia:::coverage_normal(0.90)

  comp |>
    unnest(data) |>
    mutate(lower_ci = est + (crit * se),
           upper_ci = est - (crit * se)) |>
    group_by(model) |> 
    mutate(label = if_else(
      {{ x }} == max( {{ x }} ),
      as.character(model),
      NA_character_)) |> 
    ggplot(aes(x = {{ x }}, y = est, group = model)) +
    geom_line(aes(color = model, linetype = model)) +
    geom_ribbon(aes(ymin = lower_ci,
                    ymax = upper_ci,
                    fill = model),
                alpha = 0.25) +
    scale_x_continuous(expand = expansion(mult = c(0.05,0.05))) +
    scale_fill_manual(name = "", values = unname(palette.colors(palette = "Okabe-Ito")[2:4])) +
    scale_color_manual(name = "", values = unname(palette.colors(palette = "Okabe-Ito")[2:4])) +
    scale_linetype_manual(name = "", values = c("solid","twodash","dashed")) +
    labs(x = xlab, y = ylab,
         subtitle = subtitle) +
    theme_TWRI_print(base_family = "Arial") +
    theme(axis.title.y = element_markdown(size = 8),
          axis.title.x = element_markdown(size = 8),
          axis.text.x = element_text(size = 8),
          axis.text.y = element_text(size = 8),
          panel.grid = element_blank(),
          plot.subtitle = element_text(size = 8),
          legend.direction = "vertical",
          legend.text = element_text(size = 8),
          legend.key.width = unit(25, "points"))

}

`TCEQ-13383` <- tar_read(tp_lavaca_13383_flow, store = store)

comp <- compare_smooths(model = `TCEQ-13383`,
                      `TCEQ-13384` = tar_read(tp_lavaca_13384_flow, store = store),
                      `TCEQ-13563` = tar_read(tp_lavaca_13563_flow, store = store),
                      smooths = c("s(flw_res)"),
                      overall_uncertainty = FALSE)
p_tp <- draw_smoothed_surface(comp,
                      x = flw_res,
                      xlab = "Residuals log(Inflow) [cfs]",
                      ylab =  "TP Smooth<br>Estimate [mg/L]",
                      subtitle = "A")

`TCEQ-13383` <- tar_read(no3_lavaca_13383_flow, store = store)

comp <- compare_smooths(model = `TCEQ-13383`,
                      `TCEQ-13384` = tar_read(no3_lavaca_13384_flow, store = store),
                      `TCEQ-13563` = tar_read(no3_lavaca_13563_flow, store = store),
                      smooths = c("s(flw_res)"),
                      overall_uncertainty = FALSE)
p_no3 <- draw_smoothed_surface(comp,
                      x = flw_res,
                      xlab = "Residuals log(Inflow) [cfs]",
                      ylab = "NO<sub>*x*</sub> Smooth<br>Estimate [mg/L]",
                      subtitle = "B")

`TCEQ-13383` <- tar_read(chla_lavaca_13383_flow, store = store)

comp <- compare_smooths(model = `TCEQ-13383`,
                      `TCEQ-13384` = tar_read(chla_lavaca_13384_flow, store = store),
                      `TCEQ-13563` = tar_read(chla_lavaca_13563_flow, store = store),
                      smooths = c("s(flw_res)"),
                      overall_uncertainty = FALSE)
p_chla <- draw_smoothed_surface(comp,
                      x = flw_res,
                      xlab = "Residuals log(Inflow) [cfs]",
                      ylab = "Chlorophyll-*a* Smooth<br>Estimate [µg/L]",
                      subtitle = "C")

`TCEQ-13383` <- tar_read(tkn_lavaca_13383_flow, store = store)

comp <- compare_smooths(model = `TCEQ-13383`,
                      `TCEQ-13384` = tar_read(tkn_lavaca_13384_flow, store = store),
                      `TCEQ-13563` = tar_read(tkn_lavaca_13563_flow, store = store),
                      smooths = c("s(flw_res)"),
                      overall_uncertainty = FALSE)
p_tkn <- draw_smoothed_surface(comp,
                      x = flw_res,
                      xlab = "Residuals log(Inflow) [cfs]",
                      ylab = "TKN Smooth<br>Estimate [mg/L]",
                      subtitle = "D")

`TCEQ-13383` <- tar_read(do_lavaca_13383_flow, store = store)

comp <- compare_smooths(model = `TCEQ-13383`,
                      `TCEQ-13384` = tar_read(do_lavaca_13384_flow, store = store),
                      `TCEQ-13563` = tar_read(do_lavaca_13563_flow, store = store),
                      smooths = c("s(flw_res)"),
                      overall_uncertainty = FALSE)
p_do <- draw_smoothed_surface(comp,
                      x = flw_res,
                      xlab = "Residuals log(Inflow) [cfs]",
                      ylab = "DO Smooth<br>Estimate [mg/L]",
                      subtitle = "E")

design <- "
112266
334455
"

p_tp + p_no3 + p_chla + p_tkn + p_do + guide_area() + plot_layout(guides = "collect",
                                                   design = design)

```

Increased TP loads resulted in nearly linear increases of TP concentration at
the upper- and mid-bay sites, TCEQ-13563 and TCEQ-13383 respectively 
(Figure\ \ref{fig:fig7}).
The relative effect size appeared to much smaller than the effect of freshwater
inflow alone. Increased NO~3~ loads only showed an effect at the 
lower-bay TCEQ-13384 site. The effect was quite small compared to streamflow and
provided only small improvements to the model (Table\ \ref{tab:table4}). As noted
above, nutrient loadings did not provide any explanation in changes in the
remaining assessed water quality parameters.

```{r fig7, message=FALSE, warning=FALSE, fig.width=6.85, fig.asp=0.4, fig.cap="Fitted splines from the nutrient loading GAMs display the marginal smoothed effect of TP and NO\\textsubscript{3} loads on TP (A) and NO\\textsubscript{\\textit{x}} (B) concentrations at each site in Lavaca Bay."}

`TCEQ-13383` <- tar_read(tp_lavaca_13383_full, store = store)

comp <- compare_smooths(model = `TCEQ-13383`,
                      `TCEQ-13384` = tar_read(tp_lavaca_13384_full, store = store),
                      `TCEQ-13563` = tar_read(tp_lavaca_13563_full, store = store),
                      smooths = c("s(TP_resid)"),
                      overall_uncertainty = FALSE)
p_tp <- draw_smoothed_surface(comp,
                      x = TP_resid,
                      xlab = "Residuals log(TP) Load [kg]",
                      ylab = "TP [mg/L]",
                      subtitle = "A")

`TCEQ-13383` <- tar_read(no3_lavaca_13383_full, store = store)

comp <- compare_smooths(model = `TCEQ-13383`,
                      `TCEQ-13384` = tar_read(no3_lavaca_13384_full, store = store),
                      `TCEQ-13563` = tar_read(no3_lavaca_13563_full, store = store),
                      smooths = c("s(NO3_resid)"),
                      overall_uncertainty = FALSE)
p_no3 <- draw_smoothed_surface(comp,
                      x = NO3_resid,
                      xlab = "Residuals log(NO<sub>3</sub>) Load [kg]",
                      ylab = "NO<sub>*x*</sub> [mg/L]",
                      subtitle = "B")

design <- "
11223
"

p_tp + p_no3 + guide_area() + plot_layout(guides = "collect",
                                                   design = design)

```

# Discussion {#sec4}

**resume here**

## Watershed Nutrient Loads

TP and NO~3~ loadings from the Lavaca Bay watershed showed high inter-annual
variability driven primarily by fluctuations in discharge. Notably, there were
no indications of trends in flow-normalized NO~3~ and TP loads in the Navidad River.
Freshwater discharges in the Navidad River are regulated by the Palmetto Dam forming
Lake Texana at the lower extent of the river. While the dominant agricultural land uses
differ between the Lavaca (primarily grazed pasture and rangeland) and Navidad 
(mix of pasture and row crops) catchments, we did not have a reason to expect 
different flow normalized trends between the two systems from land use alone.
Lentic nitrogen uptake and cycling may have regulating effects that mask changes
in upstream nitrogen loadings. Additional nutrient data collection in the contributing 
tributaries of Lake Texana is needed to fully assess the role of Lake Texana
in regulating nutrient delivery to the Lavaca Bay system. However, these results
suggest that there have been no changes in the NO~3~ or TP loading from Lake Texana
when accounting for variations in year to year discharge.

There is some evidence of recent decreases in flow-normalized NO~3~ loads in 
the Lavaca River. Although there is no work directly correlating water quality 
planning and implementation efforts in the watershed to water quality outcomes, 
efforts to increase agricultural producer participation in the watershed have 
been ongoing since 2016 [@schrammLavacaRiverWatershed2018; 
@bertholdDirectMailingEducation2021]. The decrease in flow-normalized NO~3~ 
loads could be a reflection of those collective efforts but the lack of evidence
for similar changes in flow-normalized TP loads provide contrary support. The
inconsistent flow-normalized trends may also reflect some of the weakness of 
the water quality dataset that is primarily composed of ambient water quality
measurements.

Cross-validation of the nutrient loading models highlights that predictions are
prone to high bias, owing to the lack of targeted storm or flow biased measurements.
The high biases are indicative that subsets of values were unable to capture the
functional relationships with the flow based dependent variables. It was beyond the scope
of the current study to evaluate the subsets of cross-validation data and scores.
However, the cross-validation procedure is indicative that more robust sampling
would be beneficial for reducing prediction variance. Supplementary flow-biased 
monitoring targeting storm- or high-flow conditions is critical to
improve model performance and strength of evidence produced by these models [@horowitzEvaluationSedimentRating2003; @snelderEstimationCatchmentNutrient2017].
There is no absolute guidance for the number of samples needed...
Would be useful to develop guidance for apply GAMs to estimate loads and estimate trends of different effect sizes.

Converted to average annual yield, the estimates of annual TP loads for the
Lavaca River are within the ranges in previous published studies
[Table\ \ref{tab:table5}, @dunnTrendsNutrientInflows1996; @rebichSourcesDeliveryNutrients2011; @omaniEstimationSedimentNutrient2014; @wise_spatially_2019]. It isn't obvious
why TP estimates in @dunnTrendsNutrientInflows1996 were notably lower. Given that none
of the studies identify substantially sized trends in TP, it is is possible that 
the period used in @dunnTrendsNutrientInflows1996 was drier on average than the
other studies. The SPARROW models used in @rebichSourcesDeliveryNutrients2011 and 
@wise_spatially_2019 utilize a version of LOADEST in the underlying load
estimation procedure, so a difference due to methodology alone is unlikely.

```{r table5, warning=FALSE, message=FALSE}

load_lav_tp <- tar_read(daily_tp_08164000, store = store)
load_nav_tp <- tar_read(daily_tp_texana, store = store)

dat <- load_lav_tp$annually |>
  ungroup() |>
  filter(year >= 2005) |>
  summarise(annual = round(mean(TP_Estimate)/2116,1),
            upper = round(mean(TP_Upper)/2116,1),
            lower = round(mean(TP_Lower)/2116,1)) ## 2116 sq km

df <- tibble(
 # Parameter = rep("TP", 5),
  "Reported Yield (kg$\\cdot$km\\textsuperscript{2}$\\cdot$year\\textsuperscript{-1})" =
    c(paste0(dat[[1,1]]," (", dat[[1,3]], ", ", dat[[1,2]], ")\\textsuperscript{a}"),
      "45.2", "42", "20.81-91.58\\textsuperscript{b}", "28.9"),
  Approach = c("GAM", "SPARROW", "SWAT", "SPARROW", "LOADEST"),
  `Time Period` = c("2005-2020", "2000-2014", "1977-2005", "1980-2002", "1972-1993"),
  Reference = c("This work", "\\cite{wise_spatially_2019}",
                "\\cite{omaniEstimationSedimentNutrient2014}",
                "\\cite{rebichSourcesDeliveryNutrients2011}",
                "\\cite{dunnTrendsNutrientInflows1996}")
)

kbl(df, format = "latex", booktabs = TRUE, escape = FALSE,
    caption = "Comparisons of previously published estimates of mean annual TP yield at the Lavaca River site.") |>
  footnote(alphabet = c("Mean of the annual point estimates and the lower and upper 95% credible intervals.",
                        "Represents a binned value range from a choropleth map."))

```

## Estuary Water Quality Responses

The non-linear temporal water quality trends identified using GAMs differed
slightly from previously identified trends [@bugica_water_2020]. This is
not unexpected due to the different time periods, different methodology, and
generally small slopes identified for most of the significant water quality
parameters in prior work. The trend in DO and 
cholorophyll-*a* concentrations are stable in comparison to other Texas
estuaries that are facing larger demands for freshwater diversions, higher
population growth, and more intense agricultural production 
[@wetzWaterQualityDynamics2016; @bugica_water_2020]. The trend of 
increasing NO~*x*~ concentration at the upper-bay TCEQ-13563 site and recent 
increases in TKN concentration at the mid-bay TCEQ-13383
site are concerning due to the nitrogen limitation identified in many Texas 
estuaries [@gardnerNitrogenFixationDissimilatory2006; 
@houTransformationFateNitrate2012; @doradoUnderstandingInteractionsFreshwater2015; 
@paudelRelationshipSuspendedSolids2019; @wetz_exceptionally_2017] 
and the relatively low ambient concentrations observed in Lavaca Bay.

The strong positive effect of freshwater inflow on NO~*x*~, TKN, and TP are
suggestive of nonpoint watershed sources, consistent with watershed uses and
with other studies relating freshwater inflow with nutrient 
concentrations in Lavaca Bay and other estuaries [@russell_effect_2006; 
@caffreyHighNutrientPulses2007; @peierlsNonmonotonicResponsesPhytoplankton2012; 
@palmerImpactsDroughtsLow2015; @ciraPhytoplanktonDynamicsLowinflow2021]. Inflow 
had a non-linear relationship with TKN at the two upstream sites, with TKN 
increasing as freshwater inflow transitioned from low to moderate levels. 
At higher freshwater inflows, the effect was attenuated, possibly indicating a 
flushing effect at higher freshwater inflow. No relationship between TKN and 
freshwater inflow were observed at TCEQ-13384 located in the lower reach of 
Lavaca Bay. Tidal flushing from Matagorda Bay could be responsible for 
diluting TKN and acting as a control on the effects of freshwater inflow in 
lower reaches of Lavaca Bay. Previous work suggests the processing of 
organic loads in the upper portions of Lavaca Bay might reduce the transport of 
nutrients into the lower reaches of the Bay [@russell_effect_2006].

Freshwater inflow had a strong positive effect on chlorophyll-*a* at 
the upper- and mid-bay sites. The upper-bay site, TCEQ-13563, showed decreases
in chlorophyll-*a* at the highest freshwater inflow volumes. Freshwater flushing
or increases in turbidity are associated with decreases in chlorophyll-*a* in
other estuaries [@peierlsNonmonotonicResponsesPhytoplankton2012; 
@cloernPhytoplanktonPrimaryProduction2014]. No relationships between inorganic
nitrogen or TP loadings with chlorophyll-*a* were observed. Due to the lack
of TKN loading information, no assessment between organic nitrogen loads and
chlorophyll-*a* were possible.

Although other studies have identified complex relationships between estuary 
nutrient concentrations, nutrient loading and chlorophyll-*a* concentrations in 
Texas estuaries [@ornolfsdottirNutrientPulsingRegulator2004; 
@doradoUnderstandingInteractionsFreshwater2015; @ciraPhytoplanktonDynamicsLowinflow2021; @tominackVariabilityPhytoplanktonBiomass2022], 
this study specifically used flow-adjusted freshwater derived nutrient loads to 
parse out contributions from changes in nutrient loadings while accounting for 
variations in load due to flow. Loading GAMs indicated no evidence of changes in 
flow-normalized TP loads in either river, and no changes in flow-normalized NO~3~
loads in the Navidad River. The small changes in flow-normalized NO~3~ loads in
the Lavaca River are probably masked under most conditions by discharge from the
Navidad River. Given the relatively small variation in flow-normalized loads,
it can be expected that they would contribute little to the variance in downstream
water quality.

GAMs did not identify responses in DO concentration to inflows or nutrient loads.
The seasonality term in the temporal GAM models explained a substantial amount of DO
variation at all of the sites. Responses of estuary metabolic processes and 
resulting DO concentrations can be quite complicated and often locally specific [@caffreyFactorsControllingNet2004]. While the lack of total nitrogen or TKN 
loading data hinders interpretation, the large seasonal effect on DO suggests 
physical factors play an important role and should be included in future models.
Prior work suggests that Lavaca Bay may not be limited by nutrients alone, 
with high turbidity or nutrient processing in upper portions of the Bay or 
intertidal river limiting production [@russell_effect_2006]. Finally, it
is reasonable to assume that fluctuations in DO may not occur immediately in 
response to nutrient pulses or freshwater inflow. Work has has shown that various
water quality parameters may have lagged effects lasting days or even months following
storms and large discharge events [@mooneyWatershedExportEvents2012a; @wetzExtremeFutureEstuaries2013; @bukaveckasInfluenceStormEvents2020; @walkerTimescalesMagnitudeWater2021]. 
However, our work only evaluates responses to
loading and inflows occuring the day of water quality observations.

# Conclusion

GAM models appear to provide reliable estimates of nutrient loads in the Lavaca
Bay watershed. However, additional flow-biased data collection efforts would
decrease the prediction variance and improve accuracy at critical 
high flow events. Ongoing projects will fill data gaps for total nitrogen and 
TKN loading. This study, consistent with others along the Texas coast, found
strong effects of freshwater flow on nutrient and chlorophyll-*a* 
concentrations. DO concentrations, dominated by seasonal effects, did not show
strong direct responses to freshwater flow. Small variance in flow-adjusted
nutrient loads indicates that (1) there have been limited changes in non-point
sources of nutrients and (2) that there isn't strong evidence that those small 
changes have had effects on chlorophyll-*a* or dissolved oxygen in Lavaca Bay. 
Although the study did not identify strong responses to changes in nutrient 
loading, this does provide a baseline assessment for future water quality 
management activities in the watershed.

\backmatter

\bmhead{Acknowledgments}
The author extends thanks to Dr. Mike Wetz (Harte Research Institute, 
Texas A&M Corpus Christi), Chad Kinsfather and Partick Brzozowski 
(Lavaca-Navidad River Authority), Brian Koch (Texas State Soil and Water 
Conservation Board), Bill Balboa (Matagorda Bay Foundation), Jason Pinchbeck 
(Texas General Land Office) and the Lavaca Bay Foundation for supporting 
development of this project and providing valuable feedback.

# Declarations {.unnumbered}

\bmhead{Funding}
This project was funded in part by a Texas Coastal Management Program grant 
approved by the Texas Land Commissioner, providing financial assistance under 
the Coastal Zone Management Act of 1972, as amended, awarded by the National 
Oceanic and Atmospheric Administration (NOAA), Office for Coastal Management,
pursuant to NOAA Award No. NA21NOS4190136. The views expressed herein are those 
of the author(s) and do not necessarily reflect the views of NOAA, the U.S. 
Department of Commerce, or any of their subagencies.

\bmhead{Data Availability}
Reproducible code and datasets generated during this study are available in
the Zenodo repository, https://doi.org/10.5281/zenodo.733075.