### 2023-08-24 README.Rmd
accepted version

---
title: "Lavaca Bay nutrient loading manuscript files"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This repo contains the rmarkdown manuscript files used to generate the following accepted manuscript:

Schramm, M. forthcoming. Linking watershed nutrient loading to estuary water quality with Generalized Additive Models. PeerJ

The rmarkdown files require LaTeX and the [lavaca-nutrients](https://github.com/TxWRI/lavaca-nutrients) repo for targets to work properly.

```{r}
renv::diagnostics()
```